### 2013-10-28 centc.tex
version 4, first latex commit

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[usenames,dvipsnames]{color}
\usepackage{natbib}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
\newcommand{\jri}[1]{{\small \textcolor{blue}{\bf #1}}}
\newcommand{\pb}[1]{{\small \textcolor{Tan}{\bf #1}}}

%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Diversity and evolution of centromere repeats in the maize genome%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Paul Bilinski \and Kevin Distor \and Jose Gutierez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and Kelly Dawe \and Jeffrey Ross-Ibarra}

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
\jri{Please add addresses etc. Gaby, Jose, and Kevin all need to be listed as plant sciences as they were here when they did the work. we can ask them if they would like their other addresses ALSO listed (fine by me) but they need to have plant sciences too.  Put Junghua as Dept. of Plant Bio UGA.  Then make a separate one for Dept. Genetics UGA, and put Kelly as both.}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}

Centromere repeats are found in most eukaryotes and play a critical role in kinetochore formation.  Though they exhibit considerable diversity both within and among species, little is understood about the mechanisms that drive centromere repeat evolution.  Here, we use maize as a model to investigate how a complex history involving polyploidy, fractionation, and recent domestication has impacted the diversity of the maize CentC repeat.  We first validate the existence of long tandem arrays of repeats in maize and other taxa in Zea.  We show that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.  In spite of this, we see little evidence that CentC variants form distinct genetic groups, instead finding that homoplasious mutations have likely homogenized CentC diversity.  We identify Cent4, a repeat initially thought to be specific to the chromosome 4 centromere, as an LTR retrotransposon that has increased drastically in frequency through domestication.  Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that a vast majority of all CentC repeats derive from only one of the parental genomes.  Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC has decreased through domestication.

\jri{I deleted the phrase talking about rates because we don’t really do/say anything about rate do we?}

\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section*{Introduction}
\label{intro}
In spite of the rapid growth of sequenced plant genomes, plant centromeres remain poorly understood and relatively cryptic, due largely to their highly repetitive content.  Centromere repeats are highly diverse across taxa and their turnover appears to be very rapid \citep{Melters2012}. However, little is known about the genetic mechanisms that produce centromere repeat diversity.  
Domesticated maize (\emph{Zea mays} ssp. \emph{mays}) has a high quality genome assembly \citep{Schnable2009} including complete sequence of two centromeres (\jri{cite wolfgruber here?? or is schnable enough?}), and the breadth of research into maize centromeres makes it one of the best systems in which to investigate the processes governing centromere repeat evolution.
\jri{The line about old nonfunctional repeats is a good point, but we don't look for these or really say much about function.}

Maize centromeres are comprised primarily of the 156bp satellite repeat CentC and a family of retrotransposons (CRM), both of which interact with  kinetochore proteins such as CENH3 (\jri{please do citations as shown above. just put the bibtex version of the citation from bookends in the refcentc.bib file.} Wolfgruber et al 2009, Zhong et al 2002), and both repeats show considerable variation in local abundance across taxa (Albert et al 2010).  But while there is considerable effort in investigating the molecular function of maize centromere repeats (Presting Chapter and references therein? \jri{here. citing “and references therein” is sort of weak scholarship. yes, i have been guilty of it myself.}), we know comparatively little about the evolution of the sequences themselves. CRM elements are better understood, including the age and insertion preferences of different CRM families (Wolfgruber et al 2009) \jri{shouldn't we cite Anupma's PNAS paper here too?}.  In contrast,  studies \jri{what studies? reference?} have only examined the flanking sequences to CentC islands despite their known association with functional centromeres.  To date, there is no in-depth characterization of the genetic diversity of centromere repeats in the maize genome.

In this paper, we describe the patterns of diversity of centromere repeats across the maize genome.  We investigate the genetic diversity of these repeats, including whether the differential ancestry of maize centromeres (Wang and Bennetzen 2012) has led to chromosome-specific repeats as found in other species (Heslop-Harrison 1999, Hirsch et al 2009) and how genetic relatedness among individual repeats varies spatially along chromosomes and across the genome.  We then compare centromere abundance across a number of maize lines, including to its wild relatives the teosintes.  We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity.  Instead, we show higher genetic similarity within clusters, potely indicating the predominance of tandem duplications in the formation of new CentC copies.  Lastly, we use low coverage sequencing and cytological analyses datshow that domesticated maize has less CentC and reli its wild relative  teosinte line

\section*{Methods}
\label{methods}

\subsection*{CentC Repeat Identification and Diversity}

We downloaded 218 previously annotated CentC sequences (Annaniev et al 1998, Nagaki et al 2003) from Genbank.  We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST (McGinnis and Madden 2004) using the 218 annotated CentCs as a reference.  We kept hits with a length of over 140bp and a minimum bit score of 100.  After meeting the bit score threshold, the longest hit was retained.  We defined CentC’s as being in tandem if the CentC’s start location was within 1000bp of the start location of another CentC.

All 12,162 CentC sequences were aligned using 7 iterations of Muscle (Edgar 2004) with default parameters.  A Jukes-Cantor distance matrix of all sequences was calculated with PHYLIP (Felsenstein 2005, \url{http://evolution.genetics.washington.edu/phylip.html}), and an unrooted neighbor joining tree was built based on the distance matrix.

We used principle coordinate analysis (PCoA) to cluster CENTC variants based on their genetic distances. Eigenvalues from the PCoA were used to determine the number of statistically significant clusters using the Tracy-Widom distribution (c.f. Patterson et al 2006).

We employed the software SpaGeDi (\url{http://ebe.ulb.ac.be/ebe/Software.html}, Hardy and Vekemans 2002) to estimate the spatial autocorrelation of sequence similarity of CENTC repeats in the completely sequenced centromeres 2 and 5.  We calculated Moran’s I statistic using Jukes-Cantor genetic distance and measures of physical distance between CENTC repeats in base pairs.  Confidence intervals for the values of I were estimated by 20,000 random permutations of the physical distances.

Statistical analyses were performed in R with the packages ape (Paradis, Claude, and Strimmer 2004) and RMTstat (Perry et al 2009).  We compared clusters to chromosome of origin and syntenic maps of maize ancient tetraploidy (Schnable et al 2011) to determine if the genetic history of maize left a footprint on CentC similarity.

\subsection*{Read Mapping and Genome Size Correction}

We mapped Illumina reads from a broad panel of Zea species \citep{Chia2012,Tenaillon2011}  to a reference consisting of the full complement of 12,162 CentC variants identified in the B73 genome.  We also used previously published whole genome chromatin immunoprecipitation (ChIP) (Wolfgruber et al 2009, Wang et al 2013) with CenH3.  Reads were mapped using  Mosaik v1.0 (\url{https://code.google.com/p/mosaik-aligner/}). We first optimized mapping parameters by relaxing mapping stringency and evaluating the number of successfully mapped reads with each combination.  Consistent with parameters from previous studies mapping reads to repetitive elements \citep{Tenaillon2011}, we required homology to remain at a minimum of 80\%.  For other non-default parameters, we permuted over many values of hash size, alignment candidate threshold,  percent of read aligning, and maximum number of hash positions per seed to find a combination that produced believable alignments.  We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Supplementary Figure X \jri{please start filling these in}).   The parameter boundary at which percentage of reads mapping increased disproportionately was identified, and parameters just below this threshold were selected.  Our final set of parameters for tandem repeats used an initial hash size of 8, an alignment candidate threshold of 15 bases, 20\% percent of mismatching bases, a minimum of 30\% overlap to the reference, and stored the top 100 hits for alignment.  After reads were mapped, we calculated the percentage of total reads hitting the given reference and multiplied this value by the relative genome size of each accession as reported in \citet{Chia2012} and \citet{Tenaillon2011}. \jri{note difference here in effect of citep and citet commands} The total number of reads mapping did not change drastically when using one random copy of CentC versus the full AGPv2 reference, suggesting that our parameters are sufficiently broad to capture genome-wide CentC.  Because library preparation has an effect on estimates of repeat abundance (see results), we only used individuals from maize HapMap v2 \citep{Chia2012} with libraries prepared using identical methods.

We used a different set of mapping parameters for long repeats such as transposable elements.  Previous studies (Schnable et al 2009) estimated that approximately 85\% of maize genome derives from transposable elements.  Using the short read libraries from Tenaillon et al (2011), we selected parameters so that approximately 85\% of the library mapped to the maize transposable element database (\url{maizetedb.org}) with a minimum homology of 80\%.  The final parameters for TEs were a hash size of 10, alignment candidate threshold of 11, 80\% homology excluding non-aligned portions of the read, and a 30\% minimum overlap.

\subsection*{CentC simulations}

\jri{needs some basic setup. we start with alignment (?) and then do what to it? what was the code written in? is the code available? if so cite github url}. Our simulation assumed that CentC had been evolving since the divergence of maize and \emph{Tripsacum} (10 million years, Ross-Ibarra et al 2009)\jri{that's not the date i cite in my paper! should be 1 million.}, a closely related genus whose centromere repeat shares a large degree of homology \citep{Melters2012}.  We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$, and one generation per year.

\subsection*{Sequencing}

Library preparation and sequencing was performed according to the methods cited in Melters et al 2012.  Using those protocols, we sequenced one individual from \emph{mays}, \emph{mexicana}, \emph{parviglumis}, and \emph{Z. luxurians} with Pacific Biosciences (\jri{cite}) technology. Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Supplementary Table \ref{supp.pacbio} \jri{check out how tables are done and referenced with this example}).  CentC copies were considered in tandem if the read had 4 CentC copies within 300bp of each other.

\subsection*{FISH}

Flourescent in situ hybridization (FISH) was used to compare abundance of CENTC repeats in chromosomes of a maize x parviglumis F1 hybrid and a maize x Z. luxurians F1 hybrid. FISH protocols closely followed those of X and Y.
\jri{Need Kelly to fill in}

\section*{Results}
\label{results}
%i imagine results go 1) repeats in the genome. 2) subgenome stuff 3) realtedness (PcoA and spagedi and mutations 4) abundance.

\subsection*{Centromere repeats in the maize genome}

We found a total of 12,162 CentC variants in the maize reference genome and the unassembled BACs.  Of these 12,162 copies, 9,896 were unique over their full length. No CentC occurred more than 10 times in the genome, and the vast majority ($>$X\% \jri{if it’s easy, this number would be nice.}) of non-unique CentC variants occurred only twice.  Of the 2,266 non-unique CentC sequences, only 3 were tandem duplicates.  Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.  Cluster size varied from single CENTC copies to 84KB, with a mean of ~7kb (approximately 45 CentC copies). Chromosomes varied greatly in copies of CentC.   Chromosome 7, with 3,200 copies, had the most CentC, while chromosome 6 had the fewest with 32 copies.  The presence of 1,523 copies of CentC on the unassembled BACs and so few copies on chromosome 6 suggest that Centromeric regions still require further assembly.

We used long-read Pacific Biosciences sequencing to verify that most CentC is in tandem arrays. We sequenced whole genome ($\approx 0.3$X \jri{I made up 0.3. can you estimate what this is? Just total Mb of data / genome size.}) libraries from 4 Zea species.  In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Supplementary Table \jri{add}).  In one 6.7kb read from the maize reference line B73, for example, we identified approximately 40 independent CentC copies in tandem, and similar arrays were seen in all four Zea species analyzed.  These results show that overall structure of the repeats has been maintained for the approximately 140,000 years since the \emph{luxurians}-\emph{mays} divergence (Hanson et al. 1996; Ross-Ibarra et al. 2009) and that a majority of CentC is found in tandem arrays (\jri{add table ref}). \jri{this last sentence is a nice addition}

\jri{seems like in this paragraph we should cite the circos figure, no?} We then identified how many large clusters of CentC were retained from each of the two parental genomes that comprise the extant maize genome.  Previous work identified the parental genome for individual chromosomal segments (Schnable et al 2011) and centromeres (Wang and Bennetzen 2012).  Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig Histogram \jri{add}).  While Schnable et al. (2010) do not assign subgenome status to repetitive regions, we assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.  Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable); of the 17 clusters $>$20KB, 16 were unambiguously assigned to subgenome 1.  Even correcting for the genome-wide overrepresentation of subgenome 1 ( 62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters ( binomial test, $p<0.005$ for both 10kb and 20kb clusters). \jri{isn't this a fisher's exact?}

\begin{figure}
% For two-column wide figures use
% Use the relevant command to insert your figure file.
%  \includegraphics[width=0.75\textwidth]{circos.png}
  \includegraphics{circos.png}
\caption{CentC repeat location in relation to the maize subgenomes.  High confidence regions are colored with darker colors while low confidence regions are colored with lighter colors.  Breakpoints between the subgenomes remain uncolored to indicate uncertainty.}
\label{circos}    
\end{figure}

In addition to CentC, previous studies have described a second high copy number centromere repeat specific to maize chromosome 4 (Page et al 2001).   BLAST analyses of Cent4 sequences, however, revealed that all 15 had high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs (Tenaillon et al 2011, Chia et al 2012), suggesting that Cent4 is actually a pericentromeric retrotransposon.

Whole-genome cenH3 chromatin immuno-precipitation data from Wolfgruber (et al 2009) shows no significant over-representation of cent4 compared to five known non-centromeric TE’s, suggesting the cent4 repeat is not involved in kinetichore formation on chromosome 4.  \jri{the jin result is very cool and i wasn't aware of. i moved to discussion though. }

\subsection*{Relatedeness of CentC in the maize genome.}

CentC copies in the maize genome exhibit tremendous diversity: the overall pairwise identity in our alignment was only 65\%, and ~98\% of sites in the alignment had at least 2 variants .  Such diversity led us to ask whether genetic groups of CentC variants could be distinguished. We performed principle coordinate analyses from a genetic distance matrix estimated from our alignment and assigned individual repeats to  genetic clusters following the approach of Patterson et al (2006).  We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Supplemental Table X \jri{add}).

The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  This predicts that local clusters of CentC should be more closely related than CentC from different clusters.  Analysis of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10 KB (Supplementary Figure SPAGEDI \jri{add}). \jri{need to standardize usage of kb. also i thought autocorrelation was significant up to larger distances?}

\begin{figure}
% For two-column wide figures use
% Use the relevant command to insert your figure file.
\includegraphics[width=1\textwidth]{heatmap.png}
\caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.  On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.  Scale bar is in Mb.  Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top) other copies within its island of tandem repeats delineated by dotted lines and (bottom) all other copies on the chromosome.  Darker colors indicate higher relatedness.  The total number of CentC in each cluster is shown below the map.  
}
\label{heatmap}    
\end{figure}

The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats. We see no evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by PCoA (Supplemental Table X \jri{add}).  A comparison of all pairs of CentC reveals a likely explanation: of the $\approx 74$ million possible pairs,  approximately 6 million \jri{this said 10, isn't it 6 observed 10 simulated??} share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.  \jri{have you verified any of these? that is, take some centcs in a single genetic group, find one that isn’t from the same physical cluster, and show that it’s in that genetic group because it shares some non-concensus mutations that are different from those repeats in its own physical cluster?} A simple forward simulation (see \ref{methods}) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.  By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\approx 10$ million pairs of repeats sharing $\geq 2$ mutations.

\subsection*{Variation in CentC abundance in Zea}

Shotgun sequence data from the maize HapMap v2 \ref{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize ((p$<0.01$; Fig. boxplot). Teosintes have more CentC than inbred maize.  Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z. luxurians} (Tenaillon et al 2011).  Analysis of these data find nearly twice as much CentC in \emph{Z. luxurians} as the maize inbred B73.  To corroborate these results, we performed fluorescent in-situ hybridization of F1 crosses between inbred maize and teosinte to determine if cytological observations agreed with our sequencing findings.  Cytology needed to be performed on an F1 cross so that we could compare relative probe fluorescence of the chromosomes within a single individual.  FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z. luxurians} have more CentC than inbred maize (Figure 3) \jri{add in this figure, and better quality images of first two}.

\section*{Discussion}
\label{discussion}

%% STUFF JEFF MAY ADD TO DISCUSSION
%Phylogenetic analysis of centromere repeats from a large number of eukaryotes found that repeat abundance evolves more rapidly than expected under a simple model of brownian motion \ref{Melters2012}.

%While genome size (Poggio et al 1998), transposable elements (TEs) (Chia et al 2012), and heterochromatic knob content (Poggio et al 1998) are known to have changed greatly through domestication, little information exists as to how domestication has impacted centromeric repeats.  Previous cytogenetic work has identified considerable variation in centromere repeat content  between domesticated maize and its wild relatives Z. mays ssp. parviglumis,), Z. mays ssp. mexicana,, and Z. luxurians (hereafter parviglumis, mexicana, and luxurians) (Albert et al 2010).

% It is likely that our empirical value is lower than our simulated value because we assume that all CentC copies have existed since the divergence event, while it is much more likely that age is high variable across copies.  Provided with such a large percentage of copies sharing mutations through homoplasy, we posit that it is possible for homoplasy, in the form of repeated mutations in different CentC copies, to drive our PCoA groupings.

Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.  Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression of its genes than maize subgenome 2.  Regulatory microRNA’s are known to correspond to centromeric repeats (Reinhart and Bartel 2002).  Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be very common across plants (A. thaliana Cheng et al 2012 and Gossypium Kapp and Wendel unpublished).  \pb{Two comments about this: first, its a terrible sentence so help make this better, because i think the point is really nice.  second, the best citations i found for cotton was an unpublished ref that is still not published from what i can gather.  advice? e.g. \url{http://www.plosone.org/article/info\%3Adoi\%2F10.1371\%2Fjournal.pone.0036442} also check out papers that cite Schabs: \url{http://scholar.google.com/scholar?cites=5313714612108479532&as_sdt=2005&sciodt=0,5&hl=en} } There are several possible explanations for the differential contribution of subgenome 1 to the diversity of CentC in the extant maize  genome. We put forth several hypotheses that may explain the strong bias for CentC to be located within a block from subgenome 1. Subgenome 2 may have had severely reduced quantities of CentC and therefore never contributed equally.  The CentC from subgenome 2 may have decayed rapidly and thus less of it remains.  Lastly, the CentC from subgenome 1 may have expanded in copy number since the allopolyploidy event.  Though we are unable to conclusively identify the reason, the lack of a large number of identical tandem duplicates  and lack of close proximity highest genetic relatedness pairs suggests that any large scale expansion of CentC has not been recent.  Also, CentC’s in subgenome 2 cluster do not appear to have accumulated excess mutation, suggesting that they are not decaying more rapidly than their subgenome 1 counterparts.  We therefore believe that it is most likely that the subgenomes had an unequal contribution of CentC at the formation of the allopolyploid, though more research about the ancient parents is required.

We also explored how genetic relatedness between CentC’s correlated with location along the physical map.  According to our PCoA analyses, CentC copies across the genome do not form distinct genetic groups correlating with past origin.  The lack of subgenome or chromosome grouping in modern CentC may be a result of homogenization of the repeats since the polyploidy event or a lack of differentiation between the repeats of the ancestors.  \jri{the fact that repeats in Trip are the same suggests no differentiation in ancestors right? if they had homogenized (read: changed) since ancestors, why would trip == maize?} \pb{unless trip and maize both went with the CentC most like one of the 2 ancestors} Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.  We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.  \pb{This part is ugly, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity ( Wolfgruber et al 2009).  Furthermore, the role of the repeat arrays on chromosomes 2 and 5 appear very different, as the largest array on chromosome 2 interacts with the kinetichore, while no array on chromosome 5 does \jri{cite}.  We find that repeats on chromosomes 2 and 5 are most highly genetically related to neighboring copies (Figure Heatmap).  This relationship was recapitulated in analyses using SpaGeDi, where CentC’s within approximately 10Kb of one another were more genetically similar (Supplemental Data).  Though we did not investigate the relationship between CentC location and genetic similarity on the other chromosomes due to incomplete sequencing of centromeres, we observe that many CentC’s within an array fall into the same  significant grouping in our Tracy-Widom clustering analysis, suggesting that high local similarity of CentC’s is a genome-wide phenomenon.  This local relatedness suggests that most CentC is evolving through tandem duplication and not a long distance mechanism such as transposition that had been previously suggested ( Shi et al 2010).  Concerning chromosomes 2 and 5, the higher local relatedness of all CentC clusters regardless of presence within the kinetochore suggests that CentC interaction with kinetochore proteins does not have a detectable change on their rates of evolution.

When investigating why a pair of chromosomes on two different chromosomes are each other closest genetic relative, we revealed that homoplasy in mutations is common across CentC variants.  We suggest that copies of CentC are sufficiently old within the genome that  homoplasious mutations cause physically distant CentC’s to be highly genetically related.  Importantly, we also do not see PCoA group capturing full clusters across chromosomes in a way that would be consistent with  retrotransposition. \pb{So that was a cool point, but I dont know how/if we can test it statisticall? kinda a cool observation.} We speculate that a vast majority of the CentC’s exist as a result of very old duplications meaning that mutations have had a long time to accumulate, as only 3 instances of identical tandem copies exist. Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, on observation we would expect if CentC’s on a chromosome share ancestry.  However, only 14\% of closest genetic pairs are found within 10KB of each other, suggesting that their tandem duplication is old and that most CentC’s have persisted within the genome for a long time. \pb{This is an important sentence, and I can’t seem to get it to communicate my point properly.  perhaps because im making too many points}

When studying CentC changes through domestication, our findings show that the repeat has experienced a decrease in copy number over time without a noticeable change in the structure of the repeat arrays. Using PacBio long read sequencing, we confirm genome-wide the observation we see on chromosomes 2 and 5, that most copies of CentC exist in large tandem arrays in all Zea taxa.  \jri{i think the discussion should largely mirror the results. this doesn’t go in the domestication part really.}  \pb{the pacbio is no where in the results, just methods.  Ok here in discussion?}From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure3).  Both teosintes within the species mays and its sister species luxurians have elevated levels of CentC.  Lower levels of centromere repeats in inbred maize contrasts previous studies that characterized the changing abundance of other common repeats.  For example, the abundance of most transposable element families increased after domestication \citep{Chia2012}.  Knowing that TE abundance increases in domesticated maize, we might have expected CentC content to increase as well if centromere size had to expand alongside most other repetitive content.  Alternatively, due to their structural role in kinetichore formation, we might have also expected centromere repeats to be largely excluded from the genome wide fluctuations in repetitive content assuming that selection exists to maintain centromere size.  Instead, the decrease in CentC content may indicate an active process of removing CentC.  We hypothesize that the removal of CentC content may correlate with genome size, as the sharp decrease in knob content through domestication actually led to an overall smaller genome size in inbred maize. The correlation between  functional  centromere size and genome size has been observed in grass species (Zhang and Dawe 2012). \jri{the problem here is kelly shows functional cent. size, not repeat array size.   } \pb{i dont think this changes the point... CentC, whether in or out the kinetichore, seems to maintain the same relatedness/diversity} Further investigation would be required to discover whether only those copies of CentC outside of the active kinetichore \jri{dude you spell kinetochore wrong throughout whole ms!} are being deleted.

\pb{I dont like wrapping up with this... but it makes the most sense.  It also does have some cool findings grounded in other works.} We also identify the cent4 repeat as the LTR retrotransposon RLX\_sela.  Characterization of the TE consensus sequences shows that it contains a 232bp LTR on both ends but lacks any of the protein sequences for autonomous transposition, such as GAG and POL complexes.  Furthermore, RLX\_sela has [three] \jri{why brackets? is this uncerstain?} regions showing $>60$bp of homology to the 180bp knob repeat, one of which is approximately 200bp in length \jri{how does a 200bp piece show homology to a 180bp repeat? (I guess it could have indels or repeats) Is this is correct?}.  The lack of binding between CenH3 and RLX\_sela seen in the ChIP data further suggests that this repeat is likely pericentromeric rather and not involved in centromere function, a result echoed by  Jin et al (2004), who showed that Cent4 probes lagging behind CentC probes in cell division. \jri{great find. I didn’t know this part.} Previous work in rice \jri{Jimiung's paper is only rice correct?} has documented the frequent presence of nonautonomous LTR retrotransposons  in or near \jri{ correct to say this?} the centromere (Jiang et al 2002).  Given that the centromere on chromosome 4 shows signs of strong selection during domestication (Hufford et al. 2012) \jri{doublecheck i'm not misremembering this please}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.

In conclusion, our study pairs bioinformatics and cytology to validate observations of decreasing centromere repeat content through maize domestication.  We show that patterns of CentC similarity are consistent with tandem duplications and maize’s allopolyploid history has [left no significant differentiation between CentC’s from the different maize subgenomes]. We also identify the chromosome 4 specific repeat, cent4, as the TE RLX\_sela and show that it has increased through domestication, unlike a majority of the other TEs \citep{Chia2012}.

\begin{acknowledgements}
We wish to thank Pacific Biosciences for sequencing resources.  We thank Lauren Sagara, Gernot Presting,  NSF summer exchange program interns Cesar, Aurelio, and Siddharth Bhadra-Lobo for helpful discussion.

\jri{needs last names}

US–NSF grant IOS-0922703

\end{acknowledgements}

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{refcentc}   % name your BibTeX data base

\newpage

\section*{Supplemental Material}

% For tables use
\begin{table}[h!]
\caption{PacBio Read Counts and Tandem CentC}
\label{supp.pacbio}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{llll}
\hline\noalign{\smallskip}
Maize Line & Reads over 600bp & Reads with $\geq 4$ CentC & \% Reads Showing Tandem CentC \\
\noalign{\smallskip}\hline\noalign{\smallskip}				
B73	& 237995	& 72	& 0.030252736	\\
\emph{luxurians}		&156964	&	79	&	0.050330012	\\
\emph{mexicana}	 	&141939 	&	150	&	0.1056792	\\
\emph{parviglumis}	&227050	& 89		&	0.039198414	\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

\end{document}
% end of file template.tex

### 2013-11-10 centc.tex
minor changes to text, new bibtex -PB
Addresses Added -PB

@@
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
+\institute{Paul Bilinski : Kevin Distor : Jose Gutierez-Lopez : Gabriela Mendoza Mendoza : Jeffrey Ross-Ibarra\at
+              Department of Plant Sciences, UC Davis\\
+              Davis, California, 95616
+              \and
+              Jeffrey Ross-Ibarra\at
+              The Genome Center and Center for Population Biology, University of California\\
+              Davis, California, 95616\\
+              \email{rossibarra@ucdavis.edu}           %  \\
 %             \emph{Present address:} of F. Author  %  if needed
            \and
-           S. Author \at
-              second address
-\jri{Please add addresses etc.
-Gaby, Jose, and Kevin all need to be listed as plant sciences as they were here when they did the work.
-we can ask them if they would like their other addresses ALSO listed (fine by me) but they need to have plant sciences too.
-Put Junghua as Dept.
-of Plant Bio UGA.
-Then make a separate one for Dept.
-Genetics UGA, and put Kelly as both.}
+           Jinghua Shi : R. Kelly Dawe\at
+           Department of Plant Biology, University of Georgia\\
+	  Athens, Georgia, 30602
+           \and
+           R. Kelly Dawe\at
+           Department of Genetics, University of Georgia\\
+           Athens, Georgia, 30602\\
 }
@@
 % The correct dates will be entered by the editor
-
 \maketitle
@@
 We used long-read Pacific Biosciences sequencing to verify that most CentC is in tandem arrays.
-We sequenced whole genome ($\approx 0.3$X \jri{I made up 0.3.
-can you estimate what this is?
-Just total Mb of data / genome size.}) libraries from 4 Zea species.
+We sequenced whole genome ($\approx 0.1$X ) libraries from 4 Zea species.
 In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Supplementary Table \jri{add}).

### 2014-04-19 centc.tex
DA PAPER
Almost ready to submit. will make slight changes, but have all of the
accessory figures.

To do: update pacbio table with the new numbers i have, push vincesims
to github, decide about wiping seal.

@@
 %
+\usepackage[utf8]{inputenc}
+\usepackage{longtable}
 \usepackage{graphicx}
@@
 \usepackage{natbib}
+\usepackage{lscape}
 
@@
 Here, we use maize as a model to investigate how a complex history involving polyploidy, fractionation, and recent domestication has impacted the diversity of the maize CentC repeat.
-We first validate the existence of long tandem arrays of repeats in maize and other taxa in Zea.
+We first validate the existence of long tandem arrays of repeats in maize and other taxa in \emph{Zea}.
 We show that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
@@
 
-\jri{I deleted the phrase talking about rates because we don’t really do/say anything about rate do we?}
-
-\keywords{First keyword \and Second keyword \and More}
+\keywords{Centromere \and Evolution \and Subgenome \and sweetscience}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 Domesticated maize (\emph{Zea mays} ssp.
-\emph{mays}) has a high quality genome assembly \citep{Schnable2009} including complete sequence of two centromeres (\jri{cite wolfgruber here??
-or is schnable enough?}), and the breadth of research into maize centromeres makes it one of the best systems in which to investigate the processes governing centromere repeat evolution.
-\jri{The line about old nonfunctional repeats is a good point, but we don't look for these or really say much about function.}
+\emph{mays}) has a high quality genome assembly \citep{Schnable2009} including complete sequence of two centromeres \citep{Wolfgruber2009}, and the breadth of research into maize centromeres makes it one of the best systems in which to investigate the processes governing centromere repeat evolution.
 
-Maize centromeres are comprised primarily of the 156bp satellite repeat CentC and a family of retrotransposons (CRM), both of which interact with  kinetochore proteins such as CENH3 (\jri{please do citations as shown above.
-just put the bibtex version of the citation from bookends in the refcentc.bib file.} Wolfgruber et al 2009, Zhong et al 2002), and both repeats show considerable variation in local abundance across taxa (Albert et al 2010).
-But while there is considerable effort in investigating the molecular function of maize centromere repeats (Presting Chapter and references therein?
-\jri{here.
-citing “and references therein” is sort of weak scholarship.
-yes, i have been guilty of it myself.}), we know comparatively little about the evolution of the sequences themselves.
-CRM elements are better understood, including the age and insertion preferences of different CRM families (Wolfgruber et al 2009) \jri{shouldn't we cite Anupma's PNAS paper here too?}.
-In contrast,  studies \jri{what studies?
-reference?} have only examined the flanking sequences to CentC islands despite their known association with functional centromeres.
+Maize centromeres are comprised primarily of the 156bp satellite repeat CentC and a family of retrotransposons (CRM), both of which interact with  kinetochore proteins such as CENH3 \citep{Wolfgruber2009, Zhong2002}, and both repeats show considerable variation in local abundance across taxa \citep{Albert2010}.
+But while there is considerable effort in investigating the molecular function of maize centromere repeats \citep{Ananiev1998B, Nagaki2003, Wolfgruber2009}, we know comparatively little about the evolution of the sequences themselves.
+CRM elements are better understood, including the age and insertion preferences of different CRM families \citep{Wolfgruber2009, Sharma2008}.
+In contrast,  studies \citep{Wang2012} have only examined the flanking sequences to CentC islands despite their known association with functional centromeres.
 To date, there is no in-depth characterization of the genetic diversity of centromere repeats in the maize genome.
@@
 In this paper, we describe the patterns of diversity of centromere repeats across the maize genome.
-We investigate the genetic diversity of these repeats, including whether the differential ancestry of maize centromeres (Wang and Bennetzen 2012) has led to chromosome-specific repeats as found in other species (Heslop-Harrison 1999, Hirsch et al 2009) and how genetic relatedness among individual repeats varies spatially along chromosomes and across the genome.
-We then compare centromere abundance across a number of maize lines, including to its wild relatives the teosintes.
+We investigate the genetic diversity of these repeats, including whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation as found in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual repeats varies spatially along chromosomes and across the genome.
+We then compare centromere abundance across a number of maize lines, including to its wild relatives, the teosintes.
 We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity.
-Instead, we show higher genetic similarity within clusters, potely indicating the predominance of tandem duplications in the formation of new CentC copies.
-Lastly, we use low coverage sequencing and cytological analyses datshow that domesticated maize has less CentC and reli its wild relative  teosinte line
+Instead, we show higher genetic similarity within clusters, potentially indicating the predominance of tandem duplications in the formation of new CentC copies.
+Lastly, we use low coverage sequencing and cytological analyses data to show that domesticated maize has less CentC than its wild relatives.
 
@@
 
-We downloaded 218 previously annotated CentC sequences (Annaniev et al 1998, Nagaki et al 2003) from Genbank.
-We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST (McGinnis and Madden 2004) using the 218 annotated CentCs as a reference.
+We downloaded 218 previously annotated CentC sequences \citep{Ananiev1998B, Nagaki2003} from Genbank.
+We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference.
 We kept hits with a length of over 140bp and a minimum bit score of 100.
@@
 
-All 12,162 CentC sequences were aligned using 7 iterations of Muscle (Edgar 2004) with default parameters.
-A Jukes-Cantor distance matrix of all sequences was calculated with PHYLIP (Felsenstein 2005, \url{http://evolution.genetics.washington.edu/phylip.html}), and an unrooted neighbor joining tree was built based on the distance matrix.
+All 12,162 CentC sequences were aligned using 7 iterations of Muscle \citep{Edgar2004} with default parameters.
+A Jukes-Cantor distance matrix of all sequences was calculated with PHYLIP (\citep{Felsenstein1989} \url{http://evolution.genetics.washington.edu/phylip.html}), and an unrooted neighbor joining tree was built based on the distance matrix.
 
 We used principle coordinate analysis (PCoA) to cluster CENTC variants based on their genetic distances.
-Eigenvalues from the PCoA were used to determine the number of statistically significant clusters using the Tracy-Widom distribution (c.f.
-Patterson et al 2006).
+Eigenvalues from the PCoA were used to determine the number of statistically significant clusters using the Tracy-Widom distribution \citep{Patterson2006}.
 
-We employed the software SpaGeDi (\url{http://ebe.ulb.ac.be/ebe/Software.html}, Hardy and Vekemans 2002) to estimate the spatial autocorrelation of sequence similarity of CENTC repeats in the completely sequenced centromeres 2 and 5.
+We employed the software SpaGeDi (\citep{Hardy2002}\url{http://ebe.ulb.ac.be/ebe/Software.html}) to estimate the spatial autocorrelation of sequence similarity of CENTC repeats in the completely sequenced centromeres 2 and 5.
 We calculated Moran’s I statistic using Jukes-Cantor genetic distance and measures of physical distance between CENTC repeats in base pairs.
@@
 
-Statistical analyses were performed in R with the packages ape (Paradis, Claude, and Strimmer 2004) and RMTstat (Perry et al 2009).
-We compared clusters to chromosome of origin and syntenic maps of maize ancient tetraploidy (Schnable et al 2011) to determine if the genetic history of maize left a footprint on CentC similarity.
+Statistical analyses were performed in R with the packages ape \citep{Paradis2004} and RMTstat \citep{Perry2009}.
+We compared clusters to chromosome of origin and syntenic maps of maize ancient tetraploidy \citep{Schnable2011} to determine if the genetic history of maize left a footprint on CentC similarity.
 
@@
 
-We mapped Illumina reads from a broad panel of Zea species \citep{Chia2012,Tenaillon2011}  to a reference consisting of the full complement of 12,162 CentC variants identified in the B73 genome.
-We also used previously published whole genome chromatin immunoprecipitation (ChIP) (Wolfgruber et al 2009, Wang et al 2013) with CenH3.
+We mapped Illumina reads from a broad panel of \emph{Zea} species \citep{Chia2012,Tenaillon2011}  to a reference consisting of the full complement of 12,162 CentC variants identified in the B73 genome.
+We also used previously published whole genome chromatin immunoprecipitation (ChIP) \citep{Wolfgruber2009, Wang2013} with CenH3.
 Reads were mapped using  Mosaik v1.0 (\url{https://code.google.com/p/mosaik-aligner/}).
@@
 For other non-default parameters, we permuted over many values of hash size, alignment candidate threshold,  percent of read aligning, and maximum number of hash positions per seed to find a combination that produced believable alignments.
-We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Supplementary Figure X \jri{please start filling these in}).
+We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Supplementary Figure \ref{Supp_MPS}).
 The parameter boundary at which percentage of reads mapping increased disproportionately was identified, and parameters just below this threshold were selected.
@@
 After reads were mapped, we calculated the percentage of total reads hitting the given reference and multiplied this value by the relative genome size of each accession as reported in \citet{Chia2012} and \citet{Tenaillon2011}.
-\jri{note difference here in effect of citep and citet commands} The total number of reads mapping did not change drastically when using one random copy of CentC versus the full AGPv2 reference, suggesting that our parameters are sufficiently broad to capture genome-wide CentC.
+The total number of reads mapping did not change drastically when using one random copy of CentC versus the full AGPv2 reference, suggesting that our parameters are sufficiently broad to capture genome-wide CentC.
 Because library preparation has an effect on estimates of repeat abundance (see results), we only used individuals from maize HapMap v2 \citep{Chia2012} with libraries prepared using identical methods.
@@
 We used a different set of mapping parameters for long repeats such as transposable elements.
-Previous studies (Schnable et al 2009) estimated that approximately 85\% of maize genome derives from transposable elements.
-Using the short read libraries from Tenaillon et al (2011), we selected parameters so that approximately 85\% of the library mapped to the maize transposable element database (\url{maizetedb.org}) with a minimum homology of 80\%.
+Previous studies \citep{Schnable2009} estimated that approximately 85\% of maize genome derives from transposable elements.
+Using the short read libraries from \citet{Tenaillon2011}, we selected parameters so that approximately 85\% of the library mapped to the maize transposable element database (\url{maizetedb.org}) with a minimum homology of 80\%.
 The final parameters for TEs were a hash size of 10, alignment candidate threshold of 11, 80\% homology excluding non-aligned portions of the read, and a 30\% minimum overlap.
 
-\subsection*{CentC simulations}
+Furthermore, we wanted to know the accuracy of our pipeline in measuring CentC content and therefore designed a simulation to vary CentC content in an artificial genome (code available at: \url{https://github.com/kddistor/dnasims}).
+In short, our simulations altered the copy number of CentC repeats over a given genome size, therefore changing the percentage of the genome deriving from the repeat.
+Illumina reads were simulated from each of the DNA strings and mapped using our pipeline.
+Relative differences in abundance were captured well, though our pipeline underestimated overall quantity of CentC.
+Altering the increments of CentC change, we found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that differences greater than 0.05\% are likely to be biologically real (Supplementary Figure \ref{Supp_Accuracy}).
 
-\jri{needs some basic setup.
-we start with alignment (?) and then do what to it?
-what was the code written in?
-is the code available?
-if so cite github url}.
-Our simulation assumed that CentC had been evolving since the divergence of maize and \emph{Tripsacum} (10 million years, Ross-Ibarra et al 2009)\jri{that's not the date i cite in my paper!
-should be 1 million.}, a closely related genus whose centromere repeat shares a large degree of homology \citep{Melters2012}.
+\subsection*{CentC homoplasious mutations simulation}
+
+\jri{needs some basic setup. we start with alignment (?) and then do what to it? what was the code written in? is the code available? if so cite github url}.
+
+We wanted to better understand the diversity present in all CentC copies across the genome.
+Looking at our alignment, we found that the sequences had a pairwise percent identify of 65.8 with only 2.2\% of sites identical, suggesting large sequence diversity.
+However, these statistics do not inform us whether the amount of diversity observed would be expected due to cumulative mutation over time.
+Therefore, we used simulations in python to count the number of sites with shared diversity across copies of CentC (code available on \url{GITHUB} \pb{we should talk about this.}).
+Our simulation assumed that CentC has been evolving for the 1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
 We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$, and one generation per year.
@@
 Using those protocols, we sequenced one individual from \emph{mays}, \emph{mexicana}, \emph{parviglumis}, and \emph{Z.
-luxurians} with Pacific Biosciences (\jri{cite}) technology.
-Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Supplementary Table \ref{supp.pacbio} \jri{check out how tables are done and referenced with this example}).
+luxurians} with Pacific Biosciences (Pacific Biosciences, Menlo Park, CA) technology.
+Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Supplementary Table \ref{supp.pacbio}).
 CentC copies were considered in tandem if the read had 4 CentC copies within 300bp of each other.
@@
 
-Flourescent in situ hybridization (FISH) was used to compare abundance of CENTC repeats in chromosomes of a maize x parviglumis F1 hybrid and a maize x Z.
-luxurians F1 hybrid.
+Fluorescent in situ hybridization (FISH) was used to compare abundance of CentC repeats in chromosomes of a maize x \emph{parviglumis} F1 hybrid and a maize x \emph{Z.
+luxurians} F1 hybrid.
 FISH protocols closely followed those of X and Y.
@@
 We found a total of 12,162 CentC variants in the maize reference genome and the unassembled BACs.
-Of these 12,162 copies, 9,896 were unique over their full length.
-No CentC occurred more than 10 times in the genome, and the vast majority ($>$X\% \jri{if it’s easy, this number would be nice.}) of non-unique CentC variants occurred only twice.
-Of the 2,266 non-unique CentC sequences, only 3 were tandem duplicates.
+Of these 12,162 copies, 8,259 were unique over their full length.
+No CentC occurred more than 10 times in the genome, and the vast majority ($>$75\%, Supplemental Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.
+Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.
 Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.
-Cluster size varied from single CENTC copies to 84KB, with a mean of ~7kb (approximately 45 CentC copies).
-Chromosomes varied greatly in copies of CentC.
-Chromosome 7, with 3,200 copies, had the most CentC, while chromosome 6 had the fewest with 32 copies.
-The presence of 1,523 copies of CentC on the unassembled BACs and so few copies on chromosome 6 suggest that Centromeric regions still require further assembly.
+Cluster size varied from single CENTC copies to 84KB, with a mean of ~7KB (approximately 45 CentC copies).
+Chromosomes varied greatly in copies of CentC, though we know that centromere assemblies for all of the chromosomes are not complete.
+For example, in mapping reads from ChIP with CENH3 to an oat plant with a single maize chromosome (OMA) (\pb{whats the ref here?
+which paper did jiming send us reads from?}), we find reads from many chromosomes have hits to the unassembled BACs (Supplemental table \ref{gaby}).
+In particular, chromosome 6 had many more reads aligning to the BACs than it did to its own centromere, suggesting a particularly lacking assembly.
+Examining total repeat number, Chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.
 
 We used long-read Pacific Biosciences sequencing to verify that most CentC is in tandem arrays.
-We sequenced whole genome ($\approx 0.1$X ) libraries from 4 Zea species.
-In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Supplementary Table \jri{add}).
-In one 6.7kb read from the maize reference line B73, for example, we identified approximately 40 independent CentC copies in tandem, and similar arrays were seen in all four Zea species analyzed.
-These results show that overall structure of the repeats has been maintained for the approximately 140,000 years since the \emph{luxurians}-\emph{mays} divergence (Hanson et al.
-1996; Ross-Ibarra et al.
-2009) and that a majority of CentC is found in tandem arrays (\jri{add table ref}).
-\jri{this last sentence is a nice addition}
+We sequenced whole genome ($\approx 0.1$X) libraries from 4 \emph{Zea} species.
+In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Supplementary Table \ref{supp.pacbio}).
+In one 6.7KB read from the maize reference line B73, for example, we identified approximately 40 independent CentC copies in tandem, and similar arrays were seen in all four \emph{Zea} species analyzed.
+These results show that overall structure of the repeats has been maintained for the approximately 140,000 years since the \emph{luxurians}-\emph{mays} divergence \citep{Hanson1996,  Ross-Ibarra2009} and that a majority of CentC is found in tandem arrays (Supplementary Table \ref{supp.pacbio}).
 
-\jri{seems like in this paragraph we should cite the circos figure, no?} We then identified how many large clusters of CentC were retained from each of the two parental genomes that comprise the extant maize genome.
-Previous work identified the parental genome for individual chromosomal segments (Schnable et al 2011) and centromeres (Wang and Bennetzen 2012).
+We then identified how many large clusters of CentC were retained from each of the two parental genomes that comprise the extant maize genome, referred to here as subgenome 1 and subgenome 2 (Figure \ref{circos}).
+Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.
 Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig Histogram \jri{add}).
 While Schnable et al.
-(2010) do not assign subgenome status to repetitive regions, we assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
+\citet{Schnable2010} do not assign subgenome status to repetitive regions, we assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
 Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable); of the 17 clusters $>$20KB, 16 were unambiguously assigned to subgenome 1.
-Even correcting for the genome-wide overrepresentation of subgenome 1 ( 62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters ( binomial test, $p<0.005$ for both 10kb and 20kb clusters).
-\jri{isn't this a fisher's exact?}
+Even correcting for the genome-wide overrepresentation of subgenome 1 (62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10KB and 20KB clusters).
 
@@
 
-In addition to CentC, previous studies have described a second high copy number centromere repeat specific to maize chromosome 4 (Page et al 2001).
-BLAST analyses of Cent4 sequences, however, revealed that all 15 had high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs (Tenaillon et al 2011, Chia et al 2012), suggesting that Cent4 is actually a pericentromeric retrotransposon.
+In addition to CentC, previous studies have described a second high copy number centromere repeat specific to maize chromosome 4 \citep{Page2001}.
+BLAST analyses of Cent4 sequences, however, revealed that all 15 had high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}, suggesting that Cent4 is actually a pericentromeric retrotransposon.
+Whole-genome cenH3 chromatin immuno-precipitation data from  \citet{Wolfgruber2009} shows no significant over-representation of cent4 compared to five known non-centromeric TE’s, suggesting the cent4 repeat is not involved in kinetochore formation on chromosome 4.
 
-Whole-genome cenH3 chromatin immuno-precipitation data from Wolfgruber (et al 2009) shows no significant over-representation of cent4 compared to five known non-centromeric TE’s, suggesting the cent4 repeat is not involved in kinetichore formation on chromosome 4.
-\jri{the jin result is very cool and i wasn't aware of.
-i moved to discussion though.
-}
-
-\subsection*{Relatedeness of CentC in the maize genome.}
+\subsection*{Relatedness of CentC in the maize genome.}
 
@@
 Such diversity led us to ask whether genetic groups of CentC variants could be distinguished.
-We performed principle coordinate analyses from a genetic distance matrix estimated from our alignment and assigned individual repeats to  genetic clusters following the approach of Patterson et al (2006).
-We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Supplemental Table X \jri{add}).
+We performed principle coordinate analyses from a genetic distance matrix estimated from our alignment and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
+We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Supplemental Table \ref{long}).
 
@@
 This predicts that local clusters of CentC should be more closely related than CentC from different clusters.
-Analysis of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10 KB (Supplementary Figure SPAGEDI \jri{add}).
-\jri{need to standardize usage of kb.
-also i thought autocorrelation was significant up to larger distances?}
+Analysis of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Supplementary Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
 
@@
 On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.
-Scale bar is in Mb.
+Scale bar is in MB.
 Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top) other copies within its island of tandem repeats delineated by dotted lines and (bottom) all other copies on the chromosome.
@@
 The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats.
-We see no evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by PCoA (Supplemental Table X \jri{add}).
-A comparison of all pairs of CentC reveals a likely explanation: of the $\approx 74$ million possible pairs,  approximately 6 million \jri{this said 10, isn't it 6 observed 10 simulated??} share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.
-\jri{have you verified any of these?
-that is, take some centcs in a single genetic group, find one that isn’t from the same physical cluster, and show that it’s in that genetic group because it shares some non-concensus mutations that are different from those repeats in its own physical cluster?} A simple forward simulation (see \ref{methods}) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.
+We see no evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by PCoA (Figure \ref{pcoa}).
+A comparison of all pairs of CentC reveals a likely explanation: of the $\approx 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.
+Looking at several triplets at random from our alignment confirms that two sequences in one PCoA assignment share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.
+A simple forward simulation (see \ref{methods} \pb{guessing you want link to github here, or link to section in methods?}) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.
 By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\approx 10$ million pairs of repeats sharing $\geq 2$ mutations.
 
-\subsection*{Variation in CentC abundance in Zea}
+\begin{figure}
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{TW_CentCgraph.jpeg}
+\caption{Presence of CentC in each of the heirarchical groups.
+The 58 clusters found to be statistically significant in forming genetic groups are represented on the x-axis and chromosome of origin for CentC's on the y-axis.
+The intersection shows the number of CentC's in the group found on that chromosome, scaled logarithmically.
+CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown.
+}
+\label{pcoa}    
+\end{figure}
 
-Shotgun sequence data from the maize HapMap v2 \ref{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize ((p$<0.01$; Fig.
-boxplot).
+\begin{figure}
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{Figure3_CentC_Abundance.png}
+\caption{(a) Mb of CentC in genomic libraries of maize and teosinte.
+Box plots show data from \citet{Chia2012}.
+Points show data for maize inbred B73 and the teosinte \emph{Z.
+luxurians} from \citet{Tenaillon2011}.
+For comparison, the data point of maize inbred B73 in \citet{Chia2012} is shown with a tick mark on the box plot.
+(b) FISH images from an F1 hybrids between maize and the teosinte \emph{parviglumis}, and maize and the teosinte \emph{Z.
+luxurians}.
+Upper row shows the karyotype for the teosinte and the lower row shows the maize karyotype.
+Red=CentC, Green=CRM.
+}
+\label{abundance}    
+\end{figure}
+
+\subsection*{Variation in CentC abundance in \emph{Zea}}
+
+Shotgun sequence data from the maize HapMap v2 \citep{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize (p$<0.01$; Figure \ref{abundance}).
 Teosintes have more CentC than inbred maize.
 Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z.
-luxurians} (Tenaillon et al 2011).
+luxurians} \citep{Tenaillon2011}.
 Analysis of these data find nearly twice as much CentC in \emph{Z.
@@
 FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z.
-luxurians} have more CentC than inbred maize (Figure 3) \jri{add in this figure, and better quality images of first two}.
+luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
+\jri{add in this figure, and better quality images of first two} \pb{of the FISH?
+thats the best quality i have insert slantyface}.
 
@@
 Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression of its genes than maize subgenome 2.
-Regulatory microRNA’s are known to correspond to centromeric repeats (Reinhart and Bartel 2002).
-Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be very common across plants (A.
-thaliana Cheng et al 2012 and Gossypium Kapp and Wendel unpublished).
+Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
+Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be very common across plants (\emph{A.
+thaliana} \cite{Cheng2012} and \emph{Gossypium} Kapp and Wendel unpublished).
 \pb{Two comments about this: first, its a terrible sentence so help make this better, because i think the point is really nice.
@@
 We also explored how genetic relatedness between CentC’s correlated with location along the physical map.
-According to our PCoA analyses, CentC copies across the genome do not form distinct genetic groups correlating with past origin.
-The lack of subgenome or chromosome grouping in modern CentC may be a result of homogenization of the repeats since the polyploidy event or a lack of differentiation between the repeats of the ancestors.
+According to our PCoA analyses, CentC copies across the genome do not form distinct genetic groups correlating with past origin (Figure \ref{pcoa}).
+The lack of subgenome or chromosome grouping in modern CentC suggests a little differentiation between the repeats since the ancestral divergence.
 \jri{the fact that repeats in Trip are the same suggests no differentiation in ancestors right?
-if they had homogenized (read: changed) since ancestors, why would trip == maize?} \pb{unless trip and maize both went with the CentC most like one of the 2 ancestors} Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.
+if they had homogenized (read: changed) since ancestors, why would trip == maize?} \pb{unless trip and maize both went with the CentC most like one of the 2 ancestors, we can't know without knowing the ancestral repeat.
+I mean I would venture to say that the repeat slated to "win" the fight in maize is also the one I would place my bets on in trip} Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.
 We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.
-\pb{This part is ugly, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity ( Wolfgruber et al 2009).
-Furthermore, the role of the repeat arrays on chromosomes 2 and 5 appear very different, as the largest array on chromosome 2 interacts with the kinetichore, while no array on chromosome 5 does \jri{cite}.
-We find that repeats on chromosomes 2 and 5 are most highly genetically related to neighboring copies (Figure Heatmap).
-This relationship was recapitulated in analyses using SpaGeDi, where CentC’s within approximately 10Kb of one another were more genetically similar (Supplemental Data).
+\pb{This part is ugly, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
+Furthermore, the role of the repeat arrays on chromosomes 2 and 5 appear very different, as the largest array on chromosome 2 interacts with the kinetochore, while no array on chromosome 5 does \citep{Wolfgruber2009}.
+We find that repeats on chromosomes 2 and 5 are most highly genetically related to neighboring copies (Figure \ref{heatmap}).
+This relationship was recapitulated in analyses using SpaGeDi, where CentC’s within approximately 10-50KB of one another were more genetically similar (Supplemental Data \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
 Though we did not investigate the relationship between CentC location and genetic similarity on the other chromosomes due to incomplete sequencing of centromeres, we observe that many CentC’s within an array fall into the same  significant grouping in our Tracy-Widom clustering analysis, suggesting that high local similarity of CentC’s is a genome-wide phenomenon.
-This local relatedness suggests that most CentC is evolving through tandem duplication and not a long distance mechanism such as transposition that had been previously suggested ( Shi et al 2010).
+This local relatedness suggests that most CentC is evolving through tandem duplication and not a long distance mechanism such as transposition that had been previously suggested \citep{Shi2010}.
 Concerning chromosomes 2 and 5, the higher local relatedness of all CentC clusters regardless of presence within the kinetochore suggests that CentC interaction with kinetochore proteins does not have a detectable change on their rates of evolution.
@@
 Importantly, we also do not see PCoA group capturing full clusters across chromosomes in a way that would be consistent with retrotransposition.
-\pb{So that was a cool point, but I dont know how/if we can test it statisticall?
-kinda a cool observation.} We speculate that a vast majority of the CentC’s exist as a result of very old duplications meaning that mutations have had a long time to accumulate, as only 3 instances of identical tandem copies exist.
-Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, on observation we would expect if CentC’s on a chromosome share ancestry.
+We speculate that a vast majority of the CentC’s exist as a result of very old duplications meaning that mutations have had a long time to accumulate.
+Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, an observation we would expect if CentC’s on a chromosome share ancestry.
 However, only 14\% of closest genetic pairs are found within 10KB of each other, suggesting that their tandem duplication is old and that most CentC’s have persisted within the genome for a long time.
@@
 When studying CentC changes through domestication, our findings show that the repeat has experienced a decrease in copy number over time without a noticeable change in the structure of the repeat arrays.
-Using PacBio long read sequencing, we confirm genome-wide the observation we see on chromosomes 2 and 5, that most copies of CentC exist in large tandem arrays in all Zea taxa.
+Using PacBio long read sequencing, we confirm genome-wide the observation we see on chromosomes 2 and 5, that most copies of CentC exist in large tandem arrays in all \emph{Zea} taxa.
 \jri{i think the discussion should largely mirror the results.
 this doesn’t go in the domestication part really.}  \pb{the pacbio is no where in the results, just methods.
-Ok here in discussion?}From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure3).
-Both teosintes within the species mays and its sister species luxurians have elevated levels of CentC.
+Ok here in discussion?}From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure \ref{abundance}).
+Both teosintes within the species \emph{mays} and its sister species \emph{luxurians} have elevated levels of CentC.
 Lower levels of centromere repeats in inbred maize contrasts previous studies that characterized the changing abundance of other common repeats.
@@
 We hypothesize that the removal of CentC content may correlate with genome size, as the sharp decrease in knob content through domestication actually led to an overall smaller genome size in inbred maize.
-The correlation between  functional  centromere size and genome size has been observed in grass species (Zhang and Dawe 2012).
+The correlation between  functional  centromere size and genome size has been observed in grass species \citep{Zhang2012}.
 \jri{the problem here is kelly shows functional cent.
 size, not repeat array size.
-} \pb{i dont think this changes the point...
-CentC, whether in or out the kinetichore, seems to maintain the same relatedness/diversity} Further investigation would be required to discover whether only those copies of CentC outside of the active kinetichore \jri{dude you spell kinetochore wrong throughout whole ms!} are being deleted.
+} \pb{i dont think this changes the point, CentC, whether in or out the kinetichore, seems to maintain the same relatedness/diversity, not sure this is addressed so left comments in here} Further investigation would be required to discover whether only those copies of CentC outside of the active kinetochore are being deleted.
 
-\pb{I dont like wrapping up with this...
-but it makes the most sense.
-It also does have some cool findings grounded in other works.} We also identify the cent4 repeat as the LTR retrotransposon RLX\_sela.
+\pb{Michelle is pretty adamant that this isnt a TE.
+As i found, no GAG or POL, but still a repeat at ends.
+Strike the whole thing, or worthwhile mentioning something about it?} We also identify the cent4 repeat as the LTR retrotransposon RLX\_sela.
 Characterization of the TE consensus sequences shows that it contains a 232bp LTR on both ends but lacks any of the protein sequences for autonomous transposition, such as GAG and POL complexes.
-Furthermore, RLX\_sela has [three] \jri{why brackets?
-is this uncerstain?} regions showing $>60$bp of homology to the 180bp knob repeat, one of which is approximately 200bp in length \jri{how does a 200bp piece show homology to a 180bp repeat?
-(I guess it could have indels or repeats) Is this is correct?}.
-The lack of binding between CenH3 and RLX\_sela seen in the ChIP data further suggests that this repeat is likely pericentromeric rather and not involved in centromere function, a result echoed by  Jin et al (2004), who showed that Cent4 probes lagging behind CentC probes in cell division.
-\jri{great find.
-I didn’t know this part.} Previous work in rice \jri{Jimiung's paper is only rice correct?} has documented the frequent presence of nonautonomous LTR retrotransposons  in or near \jri{ correct to say this?} the centromere (Jiang et al 2002).
-Given that the centromere on chromosome 4 shows signs of strong selection during domestication (Hufford et al.
-2012) \jri{doublecheck i'm not misremembering this please}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
+Furthermore, RLX\_sela has several \jri{why brackets?
+is this uncerstain?}\pb{had brackets because its unclear how I want to group these and set lengths of them, so i settle for several.} regions showing $>60$bp of homology to the 180bp knob repeat.
+The lack of binding between CenH3 and RLX\_sela seen in the ChIP data further suggests that this repeat is likely pericentromeric rather and not involved in centromere function, a result echoed by  \citet{Jin2004}, who showed that Cent4 probes lagging behind CentC probes in cell division.
+Previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}.
+Given that the centromere on chromosome 4 shows signs of strong selection during domestication \citep{Hufford2012} \jri{doublecheck i'm not misremembering this please} \pb{Super high peak in landracevsimproved just below centromere, and reasonable peak around centromere in teovslandrace.
+However, I am pretty sure there are enough peaks to say sure, one of those is this}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
 
 In conclusion, our study pairs bioinformatics and cytology to validate observations of decreasing centromere repeat content through maize domestication.
-We show that patterns of CentC similarity are consistent with tandem duplications and maize’s allopolyploid history has [left no significant differentiation between CentC’s from the different maize subgenomes].
+We show that patterns of CentC similarity are consistent with tandem duplications and maize’s allopolyploid history has left no significant differentiation between CentC’s from the different maize subgenomes.
 We also identify the chromosome 4 specific repeat, cent4, as the TE RLX\_sela and show that it has increased through domestication, unlike a majority of the other TEs \citep{Chia2012}.
@@
 We wish to thank Pacific Biosciences for sequencing resources.
-We thank Lauren Sagara, Gernot Presting,  NSF summer exchange program interns Cesar, Aurelio, and Siddharth Bhadra-Lobo for helpful discussion.
-
-\jri{needs last names}
+We thank Lauren Sagara, Gernot Presting,  NSF summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista, and Siddharth Bhadra-Lobo for helpful discussion.
 
@@
 
+\begin{figure}[h!]
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{Suppl_ACT_Seln.jpeg}
+\setcounter{figure}{0}
+\caption{Parameter selection for Alignment Candidate Threshold (ACT) for Mosaik.
+All other parameters were kept constant while ACT was changed.
+ACT was the parameter for which a non-linear pattern was observed.
+We selected to use an ACT of 15, the value for which we observed the greatest relative decrease between total percent mapping values.
+The sharp change suggests that, at a lower ACT, we may be mapping a non-CentC element to our reference.}
+\label{Supp_MPS}    
+\end{figure}
+
+\begin{figure}[h!]
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{Suppl_Chr02_Rspagedi.jpeg}
+\caption{Measure of Moran's I for Chromosome 2.  Gray areas show the confidence interval, calculated using permutations of genetic distance.}
+\label{Supp_SpagediChr02}    
+\end{figure}
+
+\begin{figure}[h!]
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{Suppl_Chr05_Rspagedi.jpeg}
+\caption{Measure of Moran's I for Chromosome 5.  Gray areas show the confidence interval, calculated using permutations of genetic distance.}
+\label{Supp_SpagediChr05}    
+\end{figure}
+
+\begin{figure}[h!]
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{AccuracyTest_2graphs.jpeg}
+\caption{Graphs showing our ability to capture changes in CentC repeat abundance under constant genome size.
+We simulated 10MB of DNA with varying CentC content and  simulated Illumina reads from the DNA.
+Reads were mapped with our Mosaik pipeiline, and several simulations at each percentage of genomic content were performed.}
+\label{Supp_Accuracy}    
+\end{figure}
+
+\begin{table}[h!]
+\caption{CentC Occurence Count In Maize RefGenv2}
+\label{supp.occurence}       % Give a unique label
+% For LaTeX tables use
+\begin{tabular}{ll}
+\hline\noalign{\smallskip}
+Occurrences &  Number of CentC's\\
+\noalign{\smallskip}\hline\noalign{\smallskip}				
+1	&8259\\
+2	&1233\\
+3	&263\\
+4	&89\\
+5	&31\\
+6	&13\\
+7	&7\\
+8	&0\\
+9	&0\\
+10	&1\\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
 % For tables use
@@
 \noalign{\smallskip}\hline\noalign{\smallskip}				
-B73	& 237995	& 72	& 0.030252736	\\
+B73	& 237995	& 30	& 0.030252736	\\
 \emph{luxurians}		&156964	&	79	&	0.050330012	\\
@@
 
+% For tables use
+\begin{table}[h!]
+\caption{ChIP Reads mapping to Unassembled from different Oat-Maize Addition (OMA) Lines}
+\label{supp.gaby}       % Give a unique label
+% For LaTeX tables use
+\begin{tabular}{llll}
+\hline\noalign{\smallskip}
+File Key & Maize Chr & Percent Reads Aligning to Unassembled BACs\\
+\noalign{\smallskip}\hline\noalign{\smallskip}				
+JJ1BU (OMA 6.34) & 6 & 21.35\\
+JJ1BR (OMA 1.36) & 1 & 15.56\\
+JJ1CF (OMA 9.41) & 9 & 6.16\\
+JJ1CH (OMA 8.05) & 8 & 6.69\\
+JJ1CG (OMA 10.26) & 10 & 9.21\\
+JJ1CI (OMA 8.05) & 8 & 8.5\\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\pagebreak
+ \begin{longtable}[c]{|c|c|c|c|c|c|c|c|c|c|c|}
+ \caption{Heirarchical clustering group assignment for copies of CentC, sorted by chromosome.  The number of CentC's from each chromosome is represented in the table.\label{long}}\\ 
+\hline
+
+\hline
+Group & Chr01 & Chr02 & Chr03 & Chr04 & Chr05 & Chr06 & Chr07 & Chr08 & Chr09 & Chr10\\
+ \hline
+ \endfirsthead
+
+\hline
+\multicolumn{1}{|c|}{Table \ref{long} Continued}\\
+\hline
+Group & Chr01 & Chr02 & Chr03 & Chr04 & Chr05 & Chr06 & Chr07 & Chr08 & Chr09 & Chr10\\ 
+\hline
+ \endhead %Jeff, I absolutely cannot get this stupid repeated header thing to work.  Any guesses as to why its not?
+
+ \hline
+ \endhead
+
+ \hline
+ \endfoot
+
+ \hline
+ \endlastfoot
+
+1 & 59 & 10 & 19 & 15 & 28 & 0 & 101 & 31 & 23 & 51\\
+2 & 98 & 0 & 0 & 0 & 1 & 0 & 1 & 0 & 1 & 0\\
+3 & 149 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 1 & 1\\
+4 & 125 & 45 & 105 & 64 & 72 & 4 & 494 & 117 & 155 & 303\\
+5 & 6 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+6 & 1 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+7 & 234 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+8 & 0 & 0 & 0 & 1 & 0 & 0 & 427 & 0 & 0 & 4\\
+9 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 28 & 81\\
+10 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 152 & 38\\
+11 & 0 & 0 & 0 & 0 & 0 & 0 & 91 & 0 & 0 & 0\\
+12 & 242 & 33 & 115 & 0 & 0 & 0 & 8 & 2 & 0 & 1\\
+13 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+14 & 20 & 2 & 17 & 0 & 7 & 0 & 340 & 2 & 3 & 0\\
+15 & 28 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+16 & 65 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+17 & 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+18 & 113 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+19 & 225 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+20 & 135 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+21 & 200 & 62 & 104 & 0 & 0 & 0 & 1 & 0 & 1 & 0\\
+22 & 0 & 0 & 0 & 0 & 8 & 0 & 122 & 0 & 0 & 0\\
+23 & 7 & 0 & 0 & 3 & 0 & 0 & 98 & 205 & 123 & 3\\
+24 & 17 & 0 & 89 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+25 & 0 & 1 & 0 & 0 & 0 & 0 & 198 & 0 & 0 & 0\\
+26 & 11 & 64 & 22 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+27 & 6 & 0 & 140 & 47 & 0 & 0 & 0 & 0 & 0 & 0\\
+28 & 0 & 0 & 169 & 47 & 0 & 0 & 0 & 0 & 0 & 0\\
+29 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 151\\
+30 & 0 & 0 & 0 & 66 & 283 & 0 & 1 & 0 & 0 & 0\\
+31 & 0 & 0 & 0 & 87 & 202 & 0 & 0 & 0 & 0 & 0\\
+32 & 0 & 0 & 0 & 0 & 25 & 0 & 0 & 0 & 0 & 0\\
+33 & 0 & 0 & 0 & 0 & 0 & 12 & 55 & 0 & 0 & 0\\
+34 & 0 & 0 & 0 & 0 & 2 & 16 & 213 & 0 & 0 & 0\\
+35 & 0 & 0 & 0 & 0 & 0 & 0 & 56 & 0 & 0 & 0\\
+36 & 0 & 0 & 0 & 0 & 0 & 0 & 217 & 0 & 0 & 0\\
+37 & 0 & 0 & 0 & 0 & 0 & 0 & 25 & 0 & 0 & 0\\
+38 & 0 & 0 & 0 & 0 & 3 & 0 & 248 & 1 & 19 & 1\\
+39 & 0 & 0 & 0 & 0 & 0 & 0 & 60 & 80 & 0 & 0\\
+40 & 0 & 0 & 0 & 0 & 0 & 0 & 182 & 0 & 0 & 0\\
+41 & 0 & 0 & 0 & 0 & 0 & 0 & 90 & 0 & 0 & 0\\
+42 & 0 & 0 & 0 & 0 & 0 & 0 & 58 & 0 & 0 & 0\\
+43 & 0 & 0 & 0 & 0 & 0 & 0 & 87 & 0 & 0 & 0\\
+44 & 0 & 0 & 0 & 0 & 0 & 0 & 22 & 135 & 1 & 0\\
+45 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 85 & 98 & 0\\
+46 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 150 & 1\\
+47 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 254\\
+48 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 70 & 0\\
+49 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 193 & 1\\
+50 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 93 & 0\\
+51 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 62\\
+52 & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 0 & 0 & 311\\
+53 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 168\\
+54 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 112\\
+55 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 42\\
+56 & 0 & 0 & 0 & 0 & 0 & 0 & 2 & 0 & 0 & 226\\
+57 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 66\\
+58 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 56\\
+
+ \end{longtable}
+
 \end{document}

### 2014-04-19 centc.tex
minor edits

@@
 
-\subsection*{CentC homoplasious mutations simulation}
+\subsection*{Simulation of homoplasious mutations}
 
@@
 High confidence regions are colored with darker colors while low confidence regions are colored with lighter colors.
-Breakpoints between the subgenomes remain uncolored to indicate uncertainty.}
+Breakpoints between the subgenomes remain uncolored to indicate uncertainty.
+\jri{fix legend.
+what is middle outer inner ring?}}
 \label{circos}    
@@
 Looking at several triplets at random from our alignment confirms that two sequences in one PCoA assignment share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.
-A simple forward simulation (see \ref{methods} \pb{guessing you want link to github here, or link to section in methods?}) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.
+A simple forward simulation (see Methods) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.
 By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\approx 10$ million pairs of repeats sharing $\geq 2$ mutations.
@@
 CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown.
+\jri{low quality fig.
+use pdf not png or jpeg.}
 }
@@
 Red=CentC, Green=CRM.
+\jri{center these figs}
 }
@@
 luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
-\jri{add in this figure, and better quality images of first two} \pb{of the FISH?
-thats the best quality i have insert slantyface}.
 
@@
 Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression of its genes than maize subgenome 2.
-Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
+\jri{cite needed} Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
 Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be very common across plants (\emph{A.
@@
 The lack of subgenome or chromosome grouping in modern CentC suggests a little differentiation between the repeats since the ancestral divergence.
-\jri{the fact that repeats in Trip are the same suggests no differentiation in ancestors right?
-if they had homogenized (read: changed) since ancestors, why would trip == maize?} \pb{unless trip and maize both went with the CentC most like one of the 2 ancestors, we can't know without knowing the ancestral repeat.
-I mean I would venture to say that the repeat slated to "win" the fight in maize is also the one I would place my bets on in trip} Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.
+Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.
 We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.
@@
 this doesn’t go in the domestication part really.}  \pb{the pacbio is no where in the results, just methods.
-Ok here in discussion?}From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure \ref{abundance}).
+Ok here in discussion?} \jri{nope, need pacbio in results, even if briefly} From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure \ref{abundance}).
 Both teosintes within the species \emph{mays} and its sister species \emph{luxurians} have elevated levels of CentC.
@@
 As i found, no GAG or POL, but still a repeat at ends.
-Strike the whole thing, or worthwhile mentioning something about it?} We also identify the cent4 repeat as the LTR retrotransposon RLX\_sela.
+Strike the whole thing, or worthwhile mentioning something about it?
+} \jri{i say we keep in, but with a more nuanced explanation.
+basically we can say it is tandem.
+i say some sleuthing may be in order -- simple blast hits find it in oryza and vitis telmore repeats...}We also identify the cent4 repeat as the LTR retrotransposon RLX\_sela.
 Characterization of the TE consensus sequences shows that it contains a 232bp LTR on both ends but lacks any of the protein sequences for autonomous transposition, such as GAG and POL complexes.
@@
 \begin{acknowledgements}
+\jri{add vince, anne lorant?}
 We wish to thank Pacific Biosciences for sequencing resources.

### 2014-04-22 centc.tex
Ready for Kelly to comment

@@
 \usepackage{lscape}
+\usepackage{float}
 
@@
 
-\jri{needs some basic setup. we start with alignment (?) and then do what to it? what was the code written in? is the code available? if so cite github url}.
-
 We wanted to better understand the diversity present in all CentC copies across the genome.
@@
 However, these statistics do not inform us whether the amount of diversity observed would be expected due to cumulative mutation over time.
-Therefore, we used simulations in python to count the number of sites with shared diversity across copies of CentC (code available on \url{GITHUB} \pb{we should talk about this.}).
+Therefore, we used simulations in python to count the number of sites with shared diversity across copies of CentC (code available on \url{GITHUB}).
 Our simulation assumed that CentC has been evolving for the 1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
@@
 FISH protocols closely followed those of X and Y.
-\jri{Need Kelly to fill in}
+\pb{Kelly, would you mind helping fill this in?}
 
@@
 For example, in mapping reads from ChIP with CENH3 to an oat plant with a single maize chromosome (OMA) (\pb{whats the ref here?
-which paper did jiming send us reads from?}), we find reads from many chromosomes have hits to the unassembled BACs (Supplemental table \ref{gaby}).
+which paper did jiming send us reads from?
+Gaby?}), we find reads from many chromosomes have hits to the unassembled BACs (Supplemental table \ref{gaby}).
 In particular, chromosome 6 had many more reads aligning to the BACs than it did to its own centromere, suggesting a particularly lacking assembly.
@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.
-Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig Histogram \jri{add}).
+Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist).
 While Schnable et al.
-\citet{Schnable2010} do not assign subgenome status to repetitive regions, we assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
+\citet{Schnable2011} do not assign subgenome status to repetitive regions, we assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
 Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable); of the 17 clusters $>$20KB, 16 were unambiguously assigned to subgenome 1.
@@
 Breakpoints between the subgenomes remain uncolored to indicate uncertainty.
-\jri{fix legend.
-what is middle outer inner ring?}}
+The middle ring, shaded in blue, displays the locations of all CentCs across the genome.
+The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20KB in length.}
 \label{circos}    
@@
 
-\begin{figure}
+\begin{figure}[h]
+\centering
 % For two-column wide figures use
@@
 
-\begin{figure}
+\begin{figure}[h]
+\centering
 % For two-column wide figures use
 % Use the relevant command to insert your figure file.
-\includegraphics[width=1\textwidth]{TW_CentCgraph.jpeg}
+\includegraphics[width=1\textwidth]{Rplot02}
 \caption{Presence of CentC in each of the heirarchical groups.
@@
 CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown.
-\jri{low quality fig.
-use pdf not png or jpeg.}
 }
@@
 % Use the relevant command to insert your figure file.
-\includegraphics[width=1\textwidth]{Figure3_CentC_Abundance.png}
+\centering
+\includegraphics[width=1\textwidth]{Figure3_noedge}
 \caption{(a) Mb of CentC in genomic libraries of maize and teosinte.
@@
 Red=CentC, Green=CRM.
-\jri{center these figs}
 }
@@
 luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
+Using whole genome shotgun PacBio long reads, we further investigated the overall structure of repeats across the different \emph{Zea} species.
+Percentages of the libraries showing tandem repeats did not differ greatly across the taxa (Supplementary Figure \ref{supp.pacbio}).
 
@@
 Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.
-Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression of its genes than maize subgenome 2.
-\jri{cite needed} Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
+Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression of its genes than maize subgenome 2 \cite{Schnable2011}.
+Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
 Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be very common across plants (\emph{A.
@@
 Using PacBio long read sequencing, we confirm genome-wide the observation we see on chromosomes 2 and 5, that most copies of CentC exist in large tandem arrays in all \emph{Zea} taxa.
-\jri{i think the discussion should largely mirror the results.
-this doesn’t go in the domestication part really.}  \pb{the pacbio is no where in the results, just methods.
-Ok here in discussion?} \jri{nope, need pacbio in results, even if briefly} From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure \ref{abundance}).
+From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure \ref{abundance}).
 Both teosintes within the species \emph{mays} and its sister species \emph{luxurians} have elevated levels of CentC.
@@
 The correlation between  functional  centromere size and genome size has been observed in grass species \citep{Zhang2012}.
-\jri{the problem here is kelly shows functional cent.
-size, not repeat array size.
-} \pb{i dont think this changes the point, CentC, whether in or out the kinetichore, seems to maintain the same relatedness/diversity, not sure this is addressed so left comments in here} Further investigation would be required to discover whether only those copies of CentC outside of the active kinetochore are being deleted.
+Further investigation would be required to discover whether only those copies of CentC outside of the active kinetochore are being deleted.
 
-\pb{Michelle is pretty adamant that this isnt a TE.
-As i found, no GAG or POL, but still a repeat at ends.
-Strike the whole thing, or worthwhile mentioning something about it?
-} \jri{i say we keep in, but with a more nuanced explanation.
-basically we can say it is tandem.
-i say some sleuthing may be in order -- simple blast hits find it in oryza and vitis telmore repeats...}We also identify the cent4 repeat as the LTR retrotransposon RLX\_sela.
-Characterization of the TE consensus sequences shows that it contains a 232bp LTR on both ends but lacks any of the protein sequences for autonomous transposition, such as GAG and POL complexes.
-Furthermore, RLX\_sela has several \jri{why brackets?
-is this uncerstain?}\pb{had brackets because its unclear how I want to group these and set lengths of them, so i settle for several.} regions showing $>60$bp of homology to the 180bp knob repeat.
-The lack of binding between CenH3 and RLX\_sela seen in the ChIP data further suggests that this repeat is likely pericentromeric rather and not involved in centromere function, a result echoed by  \citet{Jin2004}, who showed that Cent4 probes lagging behind CentC probes in cell division.
+We also sought to further characterize the chromosome specific repeat cent4 \citep{Page2001}.
+Previous work showed that Cent4 probes lagged behind CentC probes in cell division \citep{Jiang2002}, suggesting that the repeat is not involved in the active kinetochore and rather located in the pericentromere.
+A lack of enrichment of cent4 in ChIP data agrees with a pericentromeric location.
+Using the published cent4 sequences, we identify the cent4 repeat as the LTR retrotransposon RLX\_sela in the transposable element database \citep{Baucom2009}.
+Characterization of the TE consensus sequence shows that it contains repetitive motifs with homology to both knob and telomere repeats.
+It lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.
 Previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}.
-Given that the centromere on chromosome 4 shows signs of strong selection during domestication \citep{Hufford2012} \jri{doublecheck i'm not misremembering this please} \pb{Super high peak in landracevsimproved just below centromere, and reasonable peak around centromere in teovslandrace.
-However, I am pretty sure there are enough peaks to say sure, one of those is this}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
+However, the RLX\_sela sequence also appears to be missing some of the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a tandem repeat, and further sequencing of the region will be required to uncover the role of cent4.
+Given that the centromere and pericentromere on chromosome 4 show signs of strong selection during domestication \citep{Hufford2012}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
 
@@
 \begin{acknowledgements}
-\jri{add vince, anne lorant?}
 We wish to thank Pacific Biosciences for sequencing resources.
-We thank Lauren Sagara, Gernot Presting,  NSF summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista, and Siddharth Bhadra-Lobo for helpful discussion.
+We thank Vince Buffalo, Lauren Sagara, Anne Lorant, Michelle Stitzer, Gernot Presting,  NSF summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista, and Siddharth Bhadra-Lobo for helpful discussion.
 
@@
 \section*{Supplemental Material}
+
+\begin{figure}[h!]
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{Supp_clusterhist}
+\setcounter{figure}{0}
+\caption{CentC cluster size across all chromosomes.}
+\label{Supp_clusterhist}    
+\end{figure}
 

### 2014-04-22 centc.tex
fixed missing bracket on line 174

@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.
-Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist).
+Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
 While Schnable et al.

### 2014-04-30 centc.tex
intro done

@@
 Here, we use maize as a model to investigate how a complex history involving polyploidy, fractionation, and recent domestication has impacted the diversity of the maize CentC repeat.
-We first validate the existence of long tandem arrays of repeats in maize and other taxa in \emph{Zea}.
-We show that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
+We first validate the existence of long tandem arrays of repeats in maize and other taxa in the genus \emph{Zea}.
+We find that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
 In spite of this, we see little evidence that CentC variants form distinct genetic groups, instead finding that homoplasious mutations have likely homogenized CentC diversity.
-We identify Cent4, a repeat initially thought to be specific to the chromosome 4 centromere, as an LTR retrotransposon that has increased drastically in frequency through domestication.
 Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that a vast majority of all CentC repeats derive from only one of the parental genomes.
-Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC has decreased through domestication.
+Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC, but not the pericentromeric repeat Cent4, has decreased through domestication.
 
-\keywords{Centromere \and Evolution \and Subgenome \and sweetscience}
+\keywords{Centromere \and Evolution \and Subgenome}
 % \PACS{PACS code1 \and PACS code2 \and more}
@@
 
-Maize centromeres are comprised primarily of the 156bp satellite repeat CentC and a family of retrotransposons (CRM), both of which interact with  kinetochore proteins such as CENH3 \citep{Wolfgruber2009, Zhong2002}, and both repeats show considerable variation in local abundance across taxa \citep{Albert2010}.
+Maize centromeres are comprised primarily of the 156bp satellite repeat CentC and the CRM family of retrotransposons.
+Both repeats interact with  kinetochore proteins such as CENH3 \citep{Wolfgruber2009, Zhong2002} and show  variation in  abundance across taxa \citep{Albert2010}.
 But while there is considerable effort in investigating the molecular function of maize centromere repeats \citep{Ananiev1998B, Nagaki2003, Wolfgruber2009}, we know comparatively little about the evolution of the sequences themselves.
 CRM elements are better understood, including the age and insertion preferences of different CRM families \citep{Wolfgruber2009, Sharma2008}.
-In contrast,  studies \citep{Wang2012} have only examined the flanking sequences to CentC islands despite their known association with functional centromeres.
-To date, there is no in-depth characterization of the genetic diversity of centromere repeats in the maize genome.
+In contrast, there is no in-depth characterization of the genetic diversity of centromere repeats in the maize genome.
+%  studies \citep{Wang2012} have only examined the flanking sequences to CentC islands despite their known association with functional centromeres.  To date,
 
 In this paper, we describe the patterns of diversity of centromere repeats across the maize genome.
-We investigate the genetic diversity of these repeats, including whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation as found in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual repeats varies spatially along chromosomes and across the genome.
-We then compare centromere abundance across a number of maize lines, including to its wild relatives, the teosintes.
-We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity.
-Instead, we show higher genetic similarity within clusters, potentially indicating the predominance of tandem duplications in the formation of new CentC copies.
+We investigate the genetic diversity of these repeats, including whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation similar to that seen in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual repeats varies spatially across the genome.
+We then compare centromere abundance across a number of samples of maize and its wild relative teosinte.
+We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, but instead that most of the large blocks of CentC appear to have been inherited from only one of the ancestral subgenomes of maize.
+We show higher genetic similarity of CentC repeats within clusters, indicating the predominance of tandem duplications in the formation of new CentC copies.
 Lastly, we use low coverage sequencing and cytological analyses data to show that domesticated maize has less CentC than its wild relatives.

### 2014-04-30 centc.tex
into results

@@
 We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Supplementary Figure \ref{Supp_MPS}).
-The parameter boundary at which percentage of reads mapping increased disproportionately was identified, and parameters just below this threshold were selected.
 Our final set of parameters for tandem repeats used an initial hash size of 8, an alignment candidate threshold of 15 bases, 20\% percent of mismatching bases, a minimum of 30\% overlap to the reference, and stored the top 100 hits for alignment.
 After reads were mapped, we calculated the percentage of total reads hitting the given reference and multiplied this value by the relative genome size of each accession as reported in \citet{Chia2012} and \citet{Tenaillon2011}.
-The total number of reads mapping did not change drastically when using one random copy of CentC versus the full AGPv2 reference, suggesting that our parameters are sufficiently broad to capture genome-wide CentC.
+The total number of reads mapping did not change drastically when using one random copy of CentC versus the full AGPv2 reference, suggesting that our parameters are sufficiently broad to capture genome-wide CentC abundances.
 Because library preparation has an effect on estimates of repeat abundance (see results), we only used individuals from maize HapMap v2 \citep{Chia2012} with libraries prepared using identical methods.
@@
 Previous studies \citep{Schnable2009} estimated that approximately 85\% of maize genome derives from transposable elements.
-Using the short read libraries from \citet{Tenaillon2011}, we selected parameters so that approximately 85\% of the library mapped to the maize transposable element database (\url{maizetedb.org}) with a minimum homology of 80\%.
+Using the short read libraries from \citet{Tenaillon2011}, we selected parameters so that approximately 85\% of the library mapped to the maize transposable element database (\url{www.maizetedb.org}) with a minimum homology of 80\%.
 The final parameters for TEs were a hash size of 10, alignment candidate threshold of 11, 80\% homology excluding non-aligned portions of the read, and a 30\% minimum overlap.
 
-Furthermore, we wanted to know the accuracy of our pipeline in measuring CentC content and therefore designed a simulation to vary CentC content in an artificial genome (code available at: \url{https://github.com/kddistor/dnasims}).
-In short, our simulations altered the copy number of CentC repeats over a given genome size, therefore changing the percentage of the genome deriving from the repeat.
+We designed a simulation to estimate the accuracy of our measurements of CentC content (code available at: \url{https://github.com/kddistor/dnasims}).
+In short, our simulations altered the copy number of CentC repeats over a region of fixed length (10Mb), changing the percentage of the genome deriving from the repeat.
 Illumina reads were simulated from each of the DNA strings and mapped using our pipeline.
-Relative differences in abundance were captured well, though our pipeline underestimated overall quantity of CentC.
-Altering the increments of CentC change, we found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that differences greater than 0.05\% are likely to be biologically real (Supplementary Figure \ref{Supp_Accuracy}).
+These simulations showed that our pipeline captured relative differences in abundance well, but underestimated total abundance of CentC.
+We found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that larger differences are likely to be biologically real (Supplementary Figure \ref{Supp_Accuracy}).
 
@@
 
-We wanted to better understand the diversity present in all CentC copies across the genome.
-Looking at our alignment, we found that the sequences had a pairwise percent identify of 65.8 with only 2.2\% of sites identical, suggesting large sequence diversity.
+%We wanted to better understand the diversity present in all CentC copies across the genome.
+Looking at our alignment, we found that the sequences had a pairwise percent identify of 65.8\% with only 2.2\% of sites identical, suggesting large sequence diversity.
 However, these statistics do not inform us whether the amount of diversity observed would be expected due to cumulative mutation over time.
-Therefore, we used simulations in python to count the number of sites with shared diversity across copies of CentC (code available on \url{GITHUB}).
-Our simulation assumed that CentC has been evolving for the 1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
-We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$, and one generation per year.
+In order to better understand patterns of diversity at CentC, we performed simulations to test the likelihood of homoplasious mutations (i.e.
+independent mutations occurring at the same position in two different CentC repeats).
+Our simulation (code available at: \jri{need this github URL!}) assumed that CentC has been evolving for  1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
+We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ (\jri{CITE Clark 2005 MBE}), and one generation per year.
 
-\subsection*{Sequencing}
+\subsection*{PacBio Sequencing}
 
-Library preparation and sequencing was performed according to the methods cited in Melters et al 2012.
+Library preparation and sequencing was performed according to the methods cited in \citep{Melters2012}.
 Using those protocols, we sequenced one individual from \emph{mays}, \emph{mexicana}, \emph{parviglumis}, and \emph{Z.
@@
 Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Supplementary Table \ref{supp.pacbio}).
-CentC copies were considered in tandem if the read had 4 CentC copies within 300bp of each other.
+CentC copies were considered tandem if the read had 4 CentC copies within 300bp of each other.
 
@@
 luxurians} F1 hybrid.
-FISH protocols closely followed those of X and Y.
-\pb{Kelly, would you mind helping fill this in?}
+\jri{Kellyto fill in}
 
@@
 
-We found a total of 12,162 CentC variants in the maize reference genome and the unassembled BACs.
-Of these 12,162 copies, 8,259 were unique over their full length.
-No CentC occurred more than 10 times in the genome, and the vast majority ($>$75\%, Supplemental Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.
+We found a total of 12,162 CentC variants in the maize reference genome and  unassembled BACs.
+Of these, 8,259 were unique over their full length.
+No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Supplemental Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.
 Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.
-Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.
-Cluster size varied from single CENTC copies to 84KB, with a mean of ~7KB (approximately 45 CentC copies).
-Chromosomes varied greatly in copies of CentC, though we know that centromere assemblies for all of the chromosomes are not complete.
-For example, in mapping reads from ChIP with CENH3 to an oat plant with a single maize chromosome (OMA) (\pb{whats the ref here?
-which paper did jiming send us reads from?
-Gaby?}), we find reads from many chromosomes have hits to the unassembled BACs (Supplemental table \ref{gaby}).
-In particular, chromosome 6 had many more reads aligning to the BACs than it did to its own centromere, suggesting a particularly lacking assembly.
+Genome-wide CentC locations also show that nearly all of the 10,639 \jri{why is this different from 8259 number?} CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.
+Clusters varied in width from single CentC copies to 84KB \jri{check is it KB or Kb?}, with a mean of $\sim$7KB ($\sim$45 CentC copies).
+Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.
+For example, CENH3-ChIP sequence from an oat-maize addition line including only one maize chromosome \jri{CITE} nonetheless has many reads that map to the unassembled BACs (Supplemental table \ref{gaby}).
+In particular, chromosome 6 had many more reads aligning to the unassembled BACs than it did to its own centromere, suggesting a particularly incomplete assembly.
 Examining total repeat number, Chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.
+\jri{consistency check: is Chromosome capitalize?}
 
 We used long-read Pacific Biosciences sequencing to verify that most CentC is in tandem arrays.
-We sequenced whole genome ($\approx 0.1$X) libraries from 4 \emph{Zea} species.
+We sequenced whole genome ($\sim 0.1$X) libraries from 4 \emph{Zea} species.
 In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Supplementary Table \ref{supp.pacbio}).
@@
 Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
-While Schnable et al.
-\citet{Schnable2011} do not assign subgenome status to repetitive regions, we assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
-Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable); of the 17 clusters $>$20KB, 16 were unambiguously assigned to subgenome 1.
+We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
+Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
+If we restrict the analysis to clusters  $>$20KB, 16 of 17 clusters were unambiguously assigned to subgenome 1.
 Even correcting for the genome-wide overrepresentation of subgenome 1 (62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10KB and 20KB clusters).
+\jri{I thought that the 1 large cluster is on chr3 that gernot said is wrong -- if so, can't we claim 17/17 or at least 16/16 assignable?}
 
@@
 \caption{CentC repeat location in relation to the maize subgenomes.
-High confidence regions are colored with darker colors while low confidence regions are colored with lighter colors.
+The outer ring depicts chromosomal assignment to the two subgenomes, with higher confidence regions are colored with darker colors.
 Breakpoints between the subgenomes remain uncolored to indicate uncertainty.
@@
 
-In addition to CentC, previous studies have described a second high copy number centromere repeat specific to maize chromosome 4 \citep{Page2001}.
-BLAST analyses of Cent4 sequences, however, revealed that all 15 had high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}, suggesting that Cent4 is actually a pericentromeric retrotransposon.
-Whole-genome cenH3 chromatin immuno-precipitation data from  \citet{Wolfgruber2009} shows no significant over-representation of cent4 compared to five known non-centromeric TE’s, suggesting the cent4 repeat is not involved in kinetochore formation on chromosome 4.
+In addition to CentC, previous studies have described Cent4, a second high copy number centromere repeat specific to maize chromosome 4 \citep{Page2001}.
+BLAST analyses of Cent4 sequences, however, revealed that all 15 had high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
+Whole-genome cenH3 chromatin immuno-precipitation data from  \citet{Wolfgruber2009} shows no significant over-representation of Cent4 compared to five known non-centromeric TE’s.
+Together, these data suggesting that Cent4 is actually a pericentromeric retrotransposon that is not involved in kinetochore formation on chromosome 4.
 
@@
 
-CentC copies in the maize genome exhibit tremendous diversity: the overall pairwise identity in our alignment was only 65\%, and ~98\% of sites in the alignment had at least 2 variants .
+CentC copies in the maize genome exhibit tremendous diversity: the overall pairwise identity in our alignment was only 65\%, and $\sim$98\% of sites in the alignment had at least 2 variants .
 Such diversity led us to ask whether genetic groups of CentC variants could be distinguished.
-We performed principle coordinate analyses from a genetic distance matrix estimated from our alignment and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
+We performed principle coordinate  \jri{principal coordinate or principal component?} analyses from a genetic distance matrix estimated from our alignment and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
 We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Supplemental Table \ref{long}).
@@
 This predicts that local clusters of CentC should be more closely related than CentC from different clusters.
-Analysis of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Supplementary Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
-
+Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Supplementary Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
+%! I'm here
 \begin{figure}[h]

### 2014-05-01 centc.tex
bogged down in the discussion

@@
 
-\subsection*{Relatedness of CentC in the maize genome.}
+\subsection*{Relatedness of CentC in the maize genome}
 
-CentC copies in the maize genome exhibit tremendous diversity: the overall pairwise identity in our alignment was only 65\%, and $\sim$98\% of sites in the alignment had at least 2 variants .
+CentC copies in the maize genome exhibit tremendous diversity: the overall pairwise identity in our alignment was only 65\%, and $\sim$98\% of sites in the alignment had at least 2 variants.
 Such diversity led us to ask whether genetic groups of CentC variants could be distinguished.
-We performed principle coordinate  \jri{principal coordinate or principal component?} analyses from a genetic distance matrix estimated from our alignment and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
-We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Supplemental Table \ref{long}).
+We performed principle coordinate  \jri{principal coordinate or principal component?} analyses (PCoA) from a genetic distance matrix estimated from our alignment, and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
+We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Figure \ref{pcoa}; Supplemental Table \ref{long}).
+
+\begin{figure}[h]
+\centering
+% For two-column wide figures use
+% Use the relevant command to insert your figure file.
+\includegraphics[width=1\textwidth]{Rplot02}
+\caption{Presence of CentC in each of the heirarchical groups.
+The 58 clusters found to be statistically significant in forming genetic groups are represented on the x-axis and chromosome of origin on the y-axis.
+The size of each point is proportional to the log number of sequences in that group on that chromosome.
+CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown.
+}
+\label{pcoa}    
+\end{figure}
 
@@
 Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Supplementary Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
-%! I'm here
+
 \begin{figure}[h]
@@
 \caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.
-On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.
+\jri{need a and b labels } On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.
 Scale bar is in MB.
-Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top) other copies within its island of tandem repeats delineated by dotted lines and (bottom) all other copies on the chromosome.
+Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top row) other copies within its island of tandem repeats delineated by dotted lines and (bottom row) all other copies on the chromosome.
 Darker colors indicate higher relatedness.
@@
 The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats.
-We see no evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by PCoA (Figure \ref{pcoa}).
-A comparison of all pairs of CentC reveals a likely explanation: of the $\approx 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.
-Looking at several triplets at random from our alignment confirms that two sequences in one PCoA assignment share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.
+We see little evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by principle component \jri{component or coordinate?} analysis (Figure \ref{pcoa}).
+A comparison of all pairs of CentC sequences reveals a likely explanation: of the $\sim 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.
+Looking at several triplets at random from our alignment confirms that two sequences in one PCoA assignment sometimes share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.
 A simple forward simulation (see Methods) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.
-By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\approx 10$ million pairs of repeats sharing $\geq 2$ mutations.
-
-\begin{figure}[h]
-\centering
-% For two-column wide figures use
-% Use the relevant command to insert your figure file.
-\includegraphics[width=1\textwidth]{Rplot02}
-\caption{Presence of CentC in each of the heirarchical groups.
-The 58 clusters found to be statistically significant in forming genetic groups are represented on the x-axis and chromosome of origin for CentC's on the y-axis.
-The intersection shows the number of CentC's in the group found on that chromosome, scaled logarithmically.
-CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown.
-}
-\label{pcoa}    
-\end{figure}
+By stochastically applying mutations to an initially homogeneous group of repeat sequences, we find that plausible parameter values produce $\sim 10$ million pairs of repeats sharing $\geq 2$ mutations.
 
@@
 Shotgun sequence data from the maize HapMap v2 \citep{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize (p$<0.01$; Figure \ref{abundance}).
-Teosintes have more CentC than inbred maize.
 Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z.
@@
 To corroborate these results, we performed fluorescent in-situ hybridization of F1 crosses between inbred maize and teosinte to determine if cytological observations agreed with our sequencing findings.
-Cytology needed to be performed on an F1 cross so that we could compare relative probe fluorescence of the chromosomes within a single individual.
 FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z.
@@
 Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.
-Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression of its genes than maize subgenome 2 \cite{Schnable2011}.
+Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression than maize subgenome 2 \cite{Schnable2011}.
 Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
@@
 e.g.
-\url{http://www.plosone.org/article/info\%3Adoi\%2F10.1371\%2Fjournal.pone.0036442} also check out papers that cite Schabs: \url{http://scholar.google.com/scholar?cites=5313714612108479532&as_sdt=2005&sciodt=0,5&hl=en} } There are several possible explanations for the differential contribution of subgenome 1 to the diversity of CentC in the extant maize  genome.
+\url{http://www.plosone.org/article/info\%3Adoi\%2F10.1371\%2Fjournal.pone.0036442} also check out papers that cite Schabs: \url{http://scholar.google.com/scholar?cites=5313714612108479532&as_sdt=2005&sciodt=0,5&hl=en} } \jri{see journal policies, but citing unpublished usually kinda lame.} There are several possible explanations for the differential contribution of subgenome 1 to the diversity of CentC in the extant maize  genome.
 We put forth several hypotheses that may explain the strong bias for CentC to be located within a block from subgenome 1.
@@
 
+\jri{quick scan of the references looks OK, but do double check format etc.}
+
 \begin{acknowledgements}
 We wish to thank Pacific Biosciences for sequencing resources.
-We thank Vince Buffalo, Lauren Sagara, Anne Lorant, Michelle Stitzer, Gernot Presting,  NSF summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista, and Siddharth Bhadra-Lobo for helpful discussion.
-
-US–NSF grant IOS-0922703
-
+We thank Siddharth Bhadra-Lobo, Vince Buffalo, Anne Lorant, Gernot Presting, Lauren Sagara, Michelle Stitzer, and NSF summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista for advice and helpful discussion.
+This project was funded by NSF grant IOS-0922703.
 \end{acknowledgements}

### 2014-05-02 centc.tex
supplement tweak

@@
 \usepackage{float}
+
+% supplement numbering here i come!
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+ }
 
@@
 For other non-default parameters, we permuted over many values of hash size, alignment candidate threshold,  percent of read aligning, and maximum number of hash positions per seed to find a combination that produced believable alignments.
-We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Supplementary Figure \ref{Supp_MPS}).
+We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Figure \ref{Supp_MPS}).
 Our final set of parameters for tandem repeats used an initial hash size of 8, an alignment candidate threshold of 15 bases, 20\% percent of mismatching bases, a minimum of 30\% overlap to the reference, and stored the top 100 hits for alignment.
@@
 These simulations showed that our pipeline captured relative differences in abundance well, but underestimated total abundance of CentC.
-We found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that larger differences are likely to be biologically real (Supplementary Figure \ref{Supp_Accuracy}).
+We found that our pipeline could accurately capture differences of 0.05\% change in CentC abundance, suggesting that larger differences are likely to be biologically real (Figure \ref{Supp_Accuracy}).
 
@@
 luxurians} with Pacific Biosciences (Pacific Biosciences, Menlo Park, CA) technology.
-Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Supplementary Table \ref{supp.pacbio}).
+Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Table \ref{supp.pacbio}).
 CentC copies were considered tandem if the read had 4 CentC copies within 300bp of each other.
@@
 We sequenced whole genome ($\sim 0.1$X) libraries from 4 \emph{Zea} species.
-In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Supplementary Table \ref{supp.pacbio}).
+In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Table \ref{supp.pacbio}).
 In one 6.7KB read from the maize reference line B73, for example, we identified approximately 40 independent CentC copies in tandem, and similar arrays were seen in all four \emph{Zea} species analyzed.
-These results show that overall structure of the repeats has been maintained for the approximately 140,000 years since the \emph{luxurians}-\emph{mays} divergence \citep{Hanson1996,  Ross-Ibarra2009} and that a majority of CentC is found in tandem arrays (Supplementary Table \ref{supp.pacbio}).
+These results show that overall structure of the repeats has been maintained for the approximately 140,000 years since the \emph{luxurians}-\emph{mays} divergence \citep{Hanson1996,  Ross-Ibarra2009} and that a majority of CentC is found in tandem arrays (Table \ref{supp.pacbio}).
 
@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.
-Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
+Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Fig \ref{Supp_clusterhist}).
 We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
@@
 This predicts that local clusters of CentC should be more closely related than CentC from different clusters.
-Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Supplementary Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
+Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
 
@@
 Using whole genome shotgun PacBio long reads, we further investigated the overall structure of repeats across the different \emph{Zea} species.
-Percentages of the libraries showing tandem repeats did not differ greatly across the taxa (Supplementary Figure \ref{supp.pacbio}).
+Percentages of the libraries showing tandem repeats did not differ greatly across the taxa (Figure \ref{supp.pacbio}).
 
@@
 
+\beginsupplement
 \section*{Supplemental Material}

### 2014-05-02 centc.tex
tweaked supp stuff

@@
 \usepackage{graphicx}
+\usepackage{booktabs}
 \usepackage{hyperref}
@@
 Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
-If we restrict the analysis to clusters  $>$20KB, 16 of 17 clusters were unambiguously assigned to subgenome 1.
+If we restrict the analysis to clusters  $>$20KB, all of the 16 unambiguously assignable clusters come from subgenome 1.
 Even correcting for the genome-wide overrepresentation of subgenome 1 (62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10KB and 20KB clusters).
-\jri{I thought that the 1 large cluster is on chr3 that gernot said is wrong -- if so, can't we claim 17/17 or at least 16/16 assignable?}
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[]
 % For two-column wide figures use
@@
 
-\begin{table}[h!]
+\begin{table}[]
 \caption{CentC Occurence Count In Maize RefGenv2}
@@
 % For tables use
-\begin{table}[h!]
+\begin{table}[]
 \caption{PacBio Read Counts and Tandem CentC}
@@
 % For tables use
-\begin{table}[h!]
+\begin{table}[]
 \caption{ChIP Reads mapping to Unassembled from different Oat-Maize Addition (OMA) Lines}
@@
 
-\pagebreak
+\clearpage
  \begin{longtable}[c]{|c|c|c|c|c|c|c|c|c|c|c|}
@@
 
-\hline
-\multicolumn{1}{|c|}{Table \ref{long} Continued}\\
+\multicolumn{1}{c}{Table \ref{long} Continued}\\
 \hline
@@
 \hline
- \endhead %Jeff, I absolutely cannot get this stupid repeated header thing to work.  Any guesses as to why its not?
-
- \hline
 \endhead
-
- \hline
- \endfoot
-
- \hline
- \endlastfoot
 
@@
 40 & 0 & 0 & 0 & 0 & 0 & 0 & 182 & 0 & 0 & 0\\ 
-41 & 0 & 0 & 0 & 0 & 0 & 0 & 90 & 0 & 0 & 0\\
+41 & 0 & 0 & 0 & 0 & 0 & 0 & 90 & 0 & 0 & 0\\  \hline
 42 & 0 & 0 & 0 & 0 & 0 & 0 & 58 & 0 & 0 & 0\\
@@
 58 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 56\\
-
+\bottomrule
 \end{longtable}

### 2014-05-02 centc.tex
merge test

@@
 
-We used principle coordinate analysis (PCoA) to cluster CENTC variants based on their genetic distances.
+We used principle coordinate analysis (PCoA) to cluster CentC variants based on their genetic distances.
 Eigenvalues from the PCoA were used to determine the number of statistically significant clusters using the Tracy-Widom distribution \citep{Patterson2006}.
 
-We employed the software SpaGeDi (\citep{Hardy2002}\url{http://ebe.ulb.ac.be/ebe/Software.html}) to estimate the spatial autocorrelation of sequence similarity of CENTC repeats in the completely sequenced centromeres 2 and 5.
-We calculated Moran’s I statistic using Jukes-Cantor genetic distance and measures of physical distance between CENTC repeats in base pairs.
+We employed the software SpaGeDi (\citep{Hardy2002}\url{http://ebe.ulb.ac.be/ebe/Software.html}) to estimate the spatial autocorrelation of sequence similarity of CentC repeats in the completely sequenced centromeres 2 and 5.
+We calculated Moran’s I statistic using Jukes-Cantor genetic distance and measures of physical distance between CentC repeats in base pairs.
 Confidence intervals for the values of I were estimated by 20,000 random permutations of the physical distances.
@@
 independent mutations occurring at the same position in two different CentC repeats).
-Our simulation (code available at: \jri{need this github URL!}) assumed that CentC has been evolving for  1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
-We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ (\jri{CITE Clark 2005 MBE}), and one generation per year.
+Our simulation (code available at: \url{https://github.com/paulbilinski/CentC_Analyses/tree/master/Diversity_sims}) assumed that CentC has been evolving for  1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
+We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ \citep{clark2005}, and one generation per year.
 
@@
 Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.
-Genome-wide CentC locations also show that nearly all of the 10,639 \jri{why is this different from 8259 number?} CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.
-Clusters varied in width from single CentC copies to 84KB \jri{check is it KB or Kb?}, with a mean of $\sim$7KB ($\sim$45 CentC copies).
+Excluding the 1,523 copies that occur on the unassembled BACs, genome-wide CentC locations also show that nearly all of the 10,639 CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.
+Clusters varied in width from single CentC copies to 84KB with a mean of $\sim$7KB ($\sim$45 CentC copies).
 Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.
-For example, CENH3-ChIP sequence from an oat-maize addition line including only one maize chromosome \jri{CITE} nonetheless has many reads that map to the unassembled BACs (Supplemental table \ref{gaby}).
+For example, CENH3-ChIP sequence from an oat-maize addition line including only one maize chromosome \citep{kynast2001} nonetheless has many reads that map to the unassembled BACs (Supplemental table \ref{supp.gaby}).
 In particular, chromosome 6 had many more reads aligning to the unassembled BACs than it did to its own centromere, suggesting a particularly incomplete assembly.
-Examining total repeat number, Chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.
-\jri{consistency check: is Chromosome capitalize?}
+Examining total repeat number, chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.
 
@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.
-Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Fig \ref{Supp_clusterhist}).
+Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
 We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
 Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
-If we restrict the analysis to clusters  $>$20KB, all of the 16 unambiguously assignable clusters come from subgenome 1.
+If we restrict the analysis to clusters  $>$20KB with clear assignment to one subgenome, all 16 clusters were found in subgenome 1 regions.
 Even correcting for the genome-wide overrepresentation of subgenome 1 (62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10KB and 20KB clusters).
+One cluster $>$20KB falls within an unassigned region on chromosome 3.
 
@@
 %  \includegraphics[width=0.75\textwidth]{circos.png}
-  \includegraphics{circos.png}
+\includegraphics{Fig1_circos.png}
 \caption{CentC repeat location in relation to the maize subgenomes.
@@
 The middle ring, shaded in blue, displays the locations of all CentCs across the genome.
-The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20KB in length.}
+The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20KB in length}
 \label{circos}    
@@
 Such diversity led us to ask whether genetic groups of CentC variants could be distinguished.
-We performed principle coordinate  \jri{principal coordinate or principal component?} analyses (PCoA) from a genetic distance matrix estimated from our alignment, and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
+We performed principle coordinate analyses (PCoA) from a genetic distance matrix estimated from our alignment, and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
 We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Figure \ref{pcoa}; Supplemental Table \ref{long}).
@@
 % Use the relevant command to insert your figure file.
-\includegraphics[width=1\textwidth]{Rplot02}
+\includegraphics[width=1\textwidth]{Fig2_TWgroups}
 \caption{Presence of CentC in each of the heirarchical groups.
@@
 The size of each point is proportional to the log number of sequences in that group on that chromosome.
-CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown.
+CentC counts from chromosomes whose centromeres were derived from subgenome 1 are colored green and those from subgenome 2 are colored brown
 }
@@
 % Use the relevant command to insert your figure file.
-\includegraphics[width=1\textwidth]{heatmap.png}
+\includegraphics[width=1\textwidth]{Fig3_Heatmap}
 \caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.
-\jri{need a and b labels } On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.
+On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.
 Scale bar is in MB.
@@
 Darker colors indicate higher relatedness.
-The total number of CentC in each cluster is shown below the map.
+The total number of CentC in each cluster is shown below the map
 }
@@
 The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats.
-We see little evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by principle component \jri{component or coordinate?} analysis (Figure \ref{pcoa}).
+We see little evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
 A comparison of all pairs of CentC sequences reveals a likely explanation: of the $\sim 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.
@@
 \centering
-\includegraphics[width=1\textwidth]{Figure3_noedge}
+\includegraphics[width=1\textwidth]{Figure4_noedge}
 \caption{(a) Mb of CentC in genomic libraries of maize and teosinte.
@@
 
-%While genome size (Poggio et al 1998), transposable elements (TEs) (Chia et al 2012), and heterochromatic knob content (Poggio et al 1998) are known to have changed greatly through domestication, little information exists as to how domestication has impacted centromeric repeats.
-Previous cytogenetic work has identified considerable variation in centromere repeat content  between domesticated maize and its wild relatives Z.
-mays ssp.
-parviglumis,), Z.
-mays ssp.
-mexicana,, and Z.
-luxurians (hereafter parviglumis, mexicana, and luxurians) (Albert et al 2010).
-
 % It is likely that our empirical value is lower than our simulated value because we assume that all CentC copies have existed since the divergence event, while it is much more likely that age is high variable across copies.
@@
 
+%!CENT4 -- where to put you, cent4?
+%We also sought to further characterize the chromosome specific repeat cent4 \citep{Page2001}.
+Previous work showed that Cent4 probes lagged behind CentC probes in cell division \citep{Jiang2002}, suggesting that the repeat is not involved in the active kinetochore and rather located in the pericentromere.
+A lack of enrichment of cent4 in ChIP data agrees with a pericentromeric location.
+Using the published cent4 sequences, we identify the cent4 repeat as the LTR retrotransposon RLX\_sela in the transposable element database \citep{Baucom2009}.
+Characterization of the TE consensus sequence shows that it contains repetitive motifs with homology to both knob and telomere repeats.
+It lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.
+Previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}.
+However, the RLX\_sela sequence also appears to be missing some of the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a tandem repeat, and further sequencing of the region will be required to uncover the role of cent4.
+Given that the centromere and pericentromere on chromosome 4 show signs of strong selection during domestication \citep{Hufford2012}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
+
+%!SUBGENOMES
 Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.
@@
 Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
-Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be very common across plants (\emph{A.
-thaliana} \cite{Cheng2012} and \emph{Gossypium} Kapp and Wendel unpublished).
-\pb{Two comments about this: first, its a terrible sentence so help make this better, because i think the point is really nice.
-second, the best citations i found for cotton was an unpublished ref that is still not published from what i can gather.
-advice?
-e.g.
-\url{http://www.plosone.org/article/info\%3Adoi\%2F10.1371\%2Fjournal.pone.0036442} also check out papers that cite Schabs: \url{http://scholar.google.com/scholar?cites=5313714612108479532&as_sdt=2005&sciodt=0,5&hl=en} } \jri{see journal policies, but citing unpublished usually kinda lame.} There are several possible explanations for the differential contribution of subgenome 1 to the diversity of CentC in the extant maize  genome.
+Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be common across plants \citep{Cheng2012}.
+There are several possible explanations for the differential contribution of subgenome 1 to the diversity of CentC in the extant maize  genome.
 We put forth several hypotheses that may explain the strong bias for CentC to be located within a block from subgenome 1.
@@
 
+%!RELATEDNESS
 We also explored how genetic relatedness between CentC’s correlated with location along the physical map.
@@
 We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.
-\pb{This part is ugly, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
+\pb{This part is ugly LIKE YOUR GIT SKILLS, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
 Furthermore, the role of the repeat arrays on chromosomes 2 and 5 appear very different, as the largest array on chromosome 2 interacts with the kinetochore, while no array on chromosome 5 does \citep{Wolfgruber2009}.
@@
 
-When studying CentC changes through domestication, our findings show that the repeat has experienced a decrease in copy number over time without a noticeable change in the structure of the repeat arrays.
-Using PacBio long read sequencing, we confirm genome-wide the observation we see on chromosomes 2 and 5, that most copies of CentC exist in large tandem arrays in all \emph{Zea} taxa.
-From short read sequencing data, we show that modern maize has reduced genomic abundance of CentC when compared to the teosintes, a finding confirmed through FISH (Figure \ref{abundance}).
-Both teosintes within the species \emph{mays} and its sister species \emph{luxurians} have elevated levels of CentC.
-Lower levels of centromere repeats in inbred maize contrasts previous studies that characterized the changing abundance of other common repeats.
-For example, the abundance of most transposable element families increased after domestication \citep{Chia2012}.
-Knowing that TE abundance increases in domesticated maize, we might have expected CentC content to increase as well if centromere size had to expand alongside most other repetitive content.
-Alternatively, due to their structural role in kinetichore formation, we might have also expected centromere repeats to be largely excluded from the genome wide fluctuations in repetitive content assuming that selection exists to maintain centromere size.
-Instead, the decrease in CentC content may indicate an active process of removing CentC.
-We hypothesize that the removal of CentC content may correlate with genome size, as the sharp decrease in knob content through domestication actually led to an overall smaller genome size in inbred maize.
-The correlation between  functional  centromere size and genome size has been observed in grass species \citep{Zhang2012}.
-Further investigation would be required to discover whether only those copies of CentC outside of the active kinetochore are being deleted.
-
-We also sought to further characterize the chromosome specific repeat cent4 \citep{Page2001}.
-Previous work showed that Cent4 probes lagged behind CentC probes in cell division \citep{Jiang2002}, suggesting that the repeat is not involved in the active kinetochore and rather located in the pericentromere.
-A lack of enrichment of cent4 in ChIP data agrees with a pericentromeric location.
-Using the published cent4 sequences, we identify the cent4 repeat as the LTR retrotransposon RLX\_sela in the transposable element database \citep{Baucom2009}.
-Characterization of the TE consensus sequence shows that it contains repetitive motifs with homology to both knob and telomere repeats.
-It lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.
-Previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}.
-However, the RLX\_sela sequence also appears to be missing some of the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a tandem repeat, and further sequencing of the region will be required to uncover the role of cent4.
-Given that the centromere and pericentromere on chromosome 4 show signs of strong selection during domestication \citep{Hufford2012}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
+%!DOMESTICATION
+\jri{insert CITE, italics, probably refs to tables/figs as appropriate}  Previous cytogenetic work  identified  differences in centromere repeat content  between domesticated maize and its wild relatives Z.
+mays ssp.
+parviglumis,), Z.
+mays ssp.
+mexicana,, and Z.
+luxurians (hereafter parviglumis, mexicana, and luxurians) \citep{Albert2010} but was unable to quantify these differences.
+Our resequencing results show that while there is little difference in the distribution of CentC in tandem arrays, the total abundance of CentC has  decreased  \jri{can we say absolute abundance?
+we have genome size estimates, right?} during domestication, and we verify this with FISH in two maize-teosinte hybrid individuals (Figure \ref{abundance}).
+This result contrasts with the apparent increase in transposable element abundance  \citep{Chia2012}, suggesting that changes in CentC are not due to causes common to all repetitive sequence.
+\jri{ are we sure TEs have increase in absolute number?
+didn't chia et al.
+look at RPKM?
+wouldn't that increase if overall genome size decreases?
+if we control for genome size, does TE content show and absolute increase?} The maize genome is smaller than its teosinte counterpart, largely due to differences in the abundance of heterochromatic knobs (Poggio et al 1998).
+\citet{Zhang2012} have postulated an adaptive relationship between centromere size and genome size based on an observed correlation between centromere size and genome size across a number of grass species.
+This model would propose that the decrease in CentC abundance seen post-domestication is due to selection for smaller active centromeres to complement the smaller overall genome size.
+While our current data are insufficient to evaluate this conclusion, future work investigating differences in CentC content among maize landraces that vary in genome size (CITE) may provide an opportunity to further test this hypothesis.
 
@@
 We wish to thank Pacific Biosciences for sequencing resources.
-We thank Siddharth Bhadra-Lobo, Vince Buffalo, Anne Lorant, Gernot Presting, Lauren Sagara, Michelle Stitzer, and NSF summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista for advice and helpful discussion.
-This project was funded by NSF grant IOS-0922703.
+We thank Siddharth Bhadra-Lobo, Vince Buffalo, Anne Lorant, Gernot Presting, Lauren Sagara, Michelle Stitzer, and National Science Foundation summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista for advice and helpful discussion.
+This project was funded by National Science Foundation grant IOS-0922703.
 \end{acknowledgements}
@@
 \begin{table}[]
-\caption{ChIP Reads mapping to Unassembled from different Oat-Maize Addition (OMA) Lines}
-\label{supp.gaby}       % Give a unique label
+\caption{ChIP Reads mapping to Unassembled from different Oat-Maize Addition (OMA) Lines.
+Reads were mapped using Bowtie2 \citep{langmead2012} with the paramater -very sensitive-local.}\label{supp.gaby}       % Give a unique label
+
 % For LaTeX tables use

### 2014-05-02 centc.tex
conflicts

@@
 %!RELATEDNESS
+This paragraph blows.
 We also explored how genetic relatedness between CentC’s correlated with location along the physical map.
@@
 We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.
-\pb{This part is ugly LIKE YOUR GIT SKILLS, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
+\pb{This part is ugly, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
 Furthermore, the role of the repeat arrays on chromosomes 2 and 5 appear very different, as the largest array on chromosome 2 interacts with the kinetochore, while no array on chromosome 5 does \citep{Wolfgruber2009}.

### 2014-05-02 centc.tex
conflicts again

@@
 %!RELATEDNESS
-This paragraph blows.
 We also explored how genetic relatedness between CentC’s correlated with location along the physical map.
@@
 We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.
-\pb{This part is ugly, and needs help} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
+\pb{This part is ugly, and needs help LIKE YOUR GIT SKILLZ I SAID} From the reference genome, we chose to investigate chromosomes 2 and 5, since they have been sequenced from end to end, allowing for inferences about the relationship between diversity and physical proximity \citep{Wolfgruber2009}.
 Furthermore, the role of the repeat arrays on chromosomes 2 and 5 appear very different, as the largest array on chromosome 2 interacts with the kinetochore, while no array on chromosome 5 does \citep{Wolfgruber2009}.

### 2014-05-03 centc.tex
more edits, plowing ahead in discussion

--- 2014-05-02 centc.tex
@@
 
+%
+ \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
+% etc.
+%
+% please place your own definitions here and don't use \def but
+\newcommand{\jri}[1]{{\small \textcolor{blue}{\bf #1}}}
+\newcommand{\pb}[1]{{\small \textcolor{Tan}{\bf #1}}}
 % supplement numbering here i come!
@@
 %
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-\newcommand{\jri}[1]{{\small \textcolor{blue}{\bf #1}}}
-\newcommand{\pb}[1]{{\small \textcolor{Tan}{\bf #1}}}
-
-%
 % Insert the name of "your journal" with
-% \journalname{myjournal}
+ \journalname{Chromosoma}
 %
@@
 
-%\titlerunning{Short form of title}        % if too long for running head
+\titlerunning{Centromere repeats in the maize genome}        % if too long for running head
 
@@
 In contrast, there is no in-depth characterization of the genetic diversity of centromere repeats in the maize genome.
-%  studies \citep{Wang2012} have only examined the flanking sequences to CentC islands despite their known association with functional centromeres.  To date,
 
@@
 \label{results}
-%i imagine results go 1) repeats in the genome. 2) subgenome stuff 3) realtedness (PcoA and spagedi and mutations 4) abundance.
 
@@
 Of these, 8,259 were unique over their full length.
-No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Supplemental Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.
+No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.
 Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.
@@
 Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.
-For example, CENH3-ChIP sequence from an oat-maize addition line including only one maize chromosome \citep{kynast2001} nonetheless has many reads that map to the unassembled BACs (Supplemental table \ref{supp.gaby}).
+For example, CENH3-ChIP sequence from an oat-maize addition line including only one maize chromosome \citep{kynast2001} nonetheless has many reads that map to the unassembled BACs (Table \ref{supp.gaby}).
 In particular, chromosome 6 had many more reads aligning to the unassembled BACs than it did to its own centromere, suggesting a particularly incomplete assembly.
@@
 In addition to CentC, previous studies have described Cent4, a second high copy number centromere repeat specific to maize chromosome 4 \citep{Page2001}. 
-BLAST analyses of Cent4 sequences, however, revealed that all 15 had high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
-Whole-genome cenH3 chromatin immuno-precipitation data from  \citet{Wolfgruber2009} shows no significant over-representation of Cent4 compared to five known non-centromeric TE’s.
-Together, these data suggesting that Cent4 is actually a pericentromeric retrotransposon that is not involved in kinetochore formation on chromosome 4.
+Available evidence does not point to an important centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, and Cent4 probes lag behind CentC probes in cell division \citep{Jiang2002}.
+BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
+Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.  
+But while previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}, RLX\_sela also appears to be missing the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a TE-derived tandem repeat unique to the pericentromere of chromosome 4.
 
@@
 We performed principle coordinate analyses (PCoA) from a genetic distance matrix estimated from our alignment, and assigned individual repeats to  genetic clusters following the approach of \citet{Patterson2006}.
-We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Figure \ref{pcoa}; Supplemental Table \ref{long}).
+We found 58 significant clusters, but observed no pattern of groupings that revealed chromosome specificity of CentC’s or the impact of historical tetraploidy (Figure \ref{pcoa}; Table \ref{long}).
 
@@
 
-% It is likely that our empirical value is lower than our simulated value because we assume that all CentC copies have existed since the divergence event, while it is much more likely that age is high variable across copies.
-Provided with such a large percentage of copies sharing mutations through homoplasy, we posit that it is possible for homoplasy, in the form of repeated mutations in different CentC copies, to drive our PCoA groupings.
+\jri{check tense throughout. i think some is past and some is present} 
+Our analysis of centromere repeat diversity across the maize genome identifies thousands of copies exhibiting tremendous diversity. 
+Nonetheless, while we can cluster the repeats into groups of related sequences, these groups seem to have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
+We see no evidence of chromosome specific repeats such as that seen in X and Y \citep{Kawabe2005, Pontes2004}, with the currently available evidence suggesting that the chromosome 4 repeats previously called Cent4 is in fact a poorly characterized retrotransposon.
 
-%!CENT4 -- where to put you, cent4?
-%We also sought to further characterize the chromosome specific repeat cent4 \citep{Page2001}.
-Previous work showed that Cent4 probes lagged behind CentC probes in cell division \citep{Jiang2002}, suggesting that the repeat is not involved in the active kinetochore and rather located in the pericentromere.
-A lack of enrichment of cent4 in ChIP data agrees with a pericentromeric location.
-Using the published cent4 sequences, we identify the cent4 repeat as the LTR retrotransposon RLX\_sela in the transposable element database \citep{Baucom2009}.
-Characterization of the TE consensus sequence shows that it contains repetitive motifs with homology to both knob and telomere repeats.
-It lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.
-Previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}.
-However, the RLX\_sela sequence also appears to be missing some of the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a tandem repeat, and further sequencing of the region will be required to uncover the role of cent4.
-Given that the centromere and pericentromere on chromosome 4 show signs of strong selection during domestication \citep{Hufford2012}, it is plausible that the increase in abundance of Cent4 seen in domesticated maize is due to the effects of linked selection on a locus important for domestication rather than novel insertions or duplications.
+Instead of chromosome-specific groups, we find that virtually all the large arrays of CentC in the maize  genome derive from the maize1 subgenome, which is known to have lower gene loss and higher average expression than the maize subgenome 2 \cite{Schnable2011}.
+\jri{is it maize1 or maize subgenome 1?
+you use both here}
+One possible explanation for this clear bias is that differential expression between subgenomes affects centromere function.
+Regulatory microRNA’s corresponding to centromeric repeats have been observed\citep{ReinhartBartel2002}, and perhaps the silenced centromere became more susceptible to deletion just like other parts of subgenome2.
+We see no evidence of a more rapid decay of CentC sequences in subgenome2, but we likely have poor ability to detect deletion of entire tandem arrays.
+An alternative explanation is simply that the subgenome1 parent had more or larger CentC arrays, leading to an unequal representation in the ancient polyploid ancestor of maize.
 
-%!SUBGENOMES
-Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.
-Most interestingly, we show that most large arrays of CentC in the maize reference genome derive from maize1 subgenome, which is known to have lower gene loss and higher average expression than maize subgenome 2 \cite{Schnable2011}.
-Regulatory microRNA’s are known to correspond to centromeric repeats \citep{ReinhartBartel2002}.
-Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be common across plants \citep{Cheng2012}.
+%Therefore, it may be worthwhile to further investigate the potential correlation between centromere retention and gene expression as the phenomenon of higher expression of genes from one ancestor may be common across plants \citep{Cheng2012}.
 There are several possible explanations for the differential contribution of subgenome 1 to the diversity of CentC in the extant maize  genome.
-We put forth several hypotheses that may explain the strong bias for CentC to be located within a block from subgenome 1.
 Subgenome 2 may have had severely reduced quantities of CentC and therefore never contributed equally.
@@
 
+% It is likely that our empirical value is lower than our simulated value because we assume that all CentC copies have existed since the divergence event, while it is much more likely that age is high variable across copies.
+Provided with such a large percentage of copies sharing mutations through homoplasy, we posit that it is possible for homoplasy, in the form of repeated mutations in different CentC copies, to drive our PCoA groupings.
+
+%Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.
+Most interestingly, we show that
+
 %!RELATEDNESS
-This paragraph blows.
 We also explored how genetic relatedness between CentC’s correlated with location along the physical map.
@@
 We find that repeats on chromosomes 2 and 5 are most highly genetically related to neighboring copies (Figure \ref{heatmap}).
-This relationship was recapitulated in analyses using SpaGeDi, where CentC’s within approximately 10-50KB of one another were more genetically similar (Supplemental Data \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
+This relationship was recapitulated in analyses using SpaGeDi, where CentC’s within approximately 10-50KB of one another were more genetically similar (Tables \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
 Though we did not investigate the relationship between CentC location and genetic similarity on the other chromosomes due to incomplete sequencing of centromeres, we observe that many CentC’s within an array fall into the same  significant grouping in our Tracy-Widom clustering analysis, suggesting that high local similarity of CentC’s is a genome-wide phenomenon.
@@
 \end{document}
-% end of file template.tex

### 2014-05-03 centc.tex
sleepy finished work on discussion. will revisit tomorrow with more brain power

@@
 Nonetheless, while we can cluster the repeats into groups of related sequences, these groups seem to have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
-We see no evidence of chromosome specific repeats such as that seen in X and Y \citep{Kawabe2005, Pontes2004}, with the currently available evidence suggesting that the chromosome 4 repeats previously called Cent4 is in fact a poorly characterized retrotransposon.
+We see no evidence of chromosome specific repeats such as that seen in X and Y \citep{Kawabe2005, Pontes2004}, suggesting the presence of mechanisms homogenizing repeats across centromeres on different chromosomes.
+We further show that Cent4, while specific to chromosome 4, appears to be a poorly characterized tandem repeat or retroelement, but is not associated with the centromere.
 
@@
 We see no evidence of a more rapid decay of CentC sequences in subgenome2, but we likely have poor ability to detect deletion of entire tandem arrays.  
-An alternative explanation is simply that the subgenome1 parent had more or larger CentC arrays, leading to an unequal representation in the ancient polyploid ancestor of maize.
+An alternative explanation is may simply be that the subgenome1 parent had more or larger CentC arrays, leading to an unequal representation in the ancient polyploid ancestor of maize.
 
@@
 %!RELATEDNESS
-We also explored how genetic relatedness between CentC’s correlated with location along the physical map.
-According to our PCoA analyses, CentC copies across the genome do not form distinct genetic groups correlating with past origin (Figure \ref{pcoa}).
-The lack of subgenome or chromosome grouping in modern CentC suggests a little differentiation between the repeats since the ancestral divergence.
-Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.
+Our sequence comparison of CentCs also enabled us to explore the relationship between genetic and physical distance between repeats.  
+Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation among repeats, and we observe genome-wide that many CentC’s within an array fall into the same genetic cluster.
+These observations are consistent with the simple idea that most repeats arise due to tandem duplication, and long distance transposition of CentC, while necessary to homogenize repeats across chromosomes \citep{Shi2010}, is relatively uncommon.
+
+One unusual result from our sequence comparison was the finding that pairs of CentC on different chromosomes may be identical or nearly so.
+Our simulations suggest that this is plausible due to homoplasmy alone under realistic assumptions about mutation rate and divergence time.
+We speculate that the vast majority of the CentC’s exist as a result of old duplications and that mutations have had a long time to accumulate.
+Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication.
+However, only 14\% of closest genetic pairs are found within 10KB of each other, suggesting that their tandem duplication is old and that most CentC’s have persisted within the genome for a long time.
+
+%pairs investigating why a pair of chromosomes on two different chromosomes are each other closest genetic relative, we revealed that homoplasy in mutations is common across CentC variants.
+We suggest that copies of CentC are sufficiently old within the genome that homoplasious mutations cause physically distant CentC’s to be highly genetically related.
+Importantly, we also do not see PCoA group capturing full clusters across chromosomes in a way that would be consistent with retrotransposition.
+
+ % Given the lack of a genome wide pattern, we also wanted to investigate genetic relatedness on individual chromosomes.
 We hypothesized that most copies of CentC arose from local duplication rather than transposition, and therefore the genetic distance between CentC’s would be lowest across CentC’s within a large tandem array.
@@
 
-When investigating why a pair of chromosomes on two different chromosomes are each other closest genetic relative, we revealed that homoplasy in mutations is common across CentC variants.
-We suggest that copies of CentC are sufficiently old within the genome that homoplasious mutations cause physically distant CentC’s to be highly genetically related.
-Importantly, we also do not see PCoA group capturing full clusters across chromosomes in a way that would be consistent with retrotransposition.
-We speculate that a vast majority of the CentC’s exist as a result of very old duplications meaning that mutations have had a long time to accumulate.
-Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, an observation we would expect if CentC’s on a chromosome share ancestry.
-However, only 14\% of closest genetic pairs are found within 10KB of each other, suggesting that their tandem duplication is old and that most CentC’s have persisted within the genome for a long time.
-\pb{This is an important sentence, and I can’t seem to get it to communicate my point properly.
-perhaps because im making too many points}
-
-%!DOMESTICATION
 \jri{insert CITE, italics, probably refs to tables/figs as appropriate}  Previous cytogenetic work  identified  differences in centromere repeat content  between domesticated maize and its wild relatives Z.
@@
 
-In conclusion, our study pairs bioinformatics and cytology to validate observations of decreasing centromere repeat content through maize domestication.
-We show that patterns of CentC similarity are consistent with tandem duplications and maize’s allopolyploid history has left no significant differentiation between CentC’s from the different maize subgenomes.
-We also identify the chromosome 4 specific repeat, cent4, as the TE RLX\_sela and show that it has increased through domestication, unlike a majority of the other TEs \citep{Chia2012}.
+In conclusion, our detailed study of centromere repeats in the maize genome has highlighted differential contribution of subgenome, spatial autocorrelation along a chromosomes, and changes in abundance over the short time scale of domestication.
+Further work evaluating CentC evolution across multiple populations and multiple related species may shed additional light on the timing and causes of these changes.
 

### 2014-05-04 centc.tex
done with discussion.

@@
 
-\jri{check tense throughout. i think some is past and some is present} 
+\jri{check tense throughout. i think some is past and some is present. also fill in species names X and Y and add CITEs} 
 Our analysis of centromere repeat diversity across the maize genome identifies thousands of copies exhibiting tremendous diversity. 
-Nonetheless, while we can cluster the repeats into groups of related sequences, these groups seem to have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
-We see no evidence of chromosome specific repeats such as that seen in X and Y \citep{Kawabe2005, Pontes2004}, suggesting the presence of mechanisms homogenizing repeats across centromeres on different chromosomes.
-We further show that Cent4, while specific to chromosome 4, appears to be a poorly characterized tandem repeat or retroelement, but is not associated with the centromere.
+But while we can cluster the repeats into groups of related sequences, these groups seem to have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
+We see no evidence of chromosome specific repeats such as that seen in X and Y \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanisms that homogenizes repeats across centromeres on different chromosomes.
+We further show that Cent4, once thought to be a chromosome-specific centromere repeat (CITE), appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere.
+Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.
 
-Instead of chromosome-specific groups, we find that virtually all the large arrays of CentC in the maize  genome derive from the maize1 subgenome, which is known to have lower gene loss and higher average expression than the maize subgenome 2 \cite{Schnable2011}.
-\jri{is it maize1 or maize subgenome 1?
-you use both here}
-One possible explanation for this clear bias is that differential expression between subgenomes affects centromere function.  
-Regulatory microRNA’s corresponding to centromeric repeats have been observed\citep{ReinhartBartel2002}, and perhaps the silenced centromere became more susceptible to deletion just like other parts of subgenome2.
-We see no evidence of a more rapid decay of CentC sequences in subgenome2, but we likely have poor ability to detect deletion of entire tandem arrays.  
-An alternative explanation is may simply be that the subgenome1 parent had more or larger CentC arrays, leading to an unequal representation in the ancient polyploid ancestor of maize.
+Instead of chromosome-specific groups, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).
+One possible explanation for this result is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
+Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
+Large CentC clusters are clearly biased in favor of subgenome 1, which is known to have lower gene loss and higher average expression than maize subgenome 2 \cite{Schnable2011}. 
+Consistent with this idea, \citet{ReinhartBartel2002} find evidence of regulatory microRNA’s corresponding to centromeric repeats.   
+Of course, an alternative explanation is simply that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.
 
@@
 %!RELATEDNESS
-Our sequence comparison of CentCs also enabled us to explore the relationship between genetic and physical distance between repeats.  
-Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation among repeats, and we observe genome-wide that many CentC’s within an array fall into the same genetic cluster.
+Our sequence comparison of CentCs also enabled us to explore the relationship between genetic and physical distance among repeats.  
+Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation of relatedness among repeats, and we observe genome-wide that many CentC’s within an array fall into the same genetic cluster.
 These observations are consistent with the simple idea that most repeats arise due to tandem duplication, and long distance transposition of CentC, while necessary to homogenize repeats across chromosomes \citep{Shi2010}, is relatively uncommon.
 
-One unusual result from our sequence comparison was the finding that pairs of CentC on different chromosomes may be identical or nearly so.
-Our simulations suggest that this is plausible due to homoplasmy alone under realistic assumptions about mutation rate and divergence time.
-We speculate that the vast majority of the CentC’s exist as a result of old duplications and that mutations have had a long time to accumulate.
-Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication.
-However, only 14\% of closest genetic pairs are found within 10KB of each other, suggesting that their tandem duplication is old and that most CentC’s have persisted within the genome for a long time.
+One unusual result from our sequence comparison was the finding that pairs of CentC on different chromosomes share very high sequence similarity.  
+Our simulations suggest that, under realistic assumptions about mutation rate and divergence time, such a pattern is possible due to homoplasious mutation alone.  
+Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication, but only 14\% of closest genetic pairs are found within 10KB of each other.
+We speculate that the vast majority of  CentC’s in the genome are thus a result of relatively old tandem duplications, and that sufficient time has occurred since duplication for rearrangements and mutations to break up patterns of identical tandem repeats.
 

### 2014-05-05 centc.tex
Figure names changed
Jose affil added, Addressed major JRI comments in disc

@@
 
-\institute{Paul Bilinski : Kevin Distor : Jose Gutierez-Lopez : Gabriela Mendoza Mendoza : Jeffrey Ross-Ibarra\at
+\institute{Paul Bilinski : Kevin Distor : Jose Gutierrez-Lopez : Gabriela Mendoza Mendoza : Jeffrey Ross-Ibarra\at
               Department of Plant Sciences, UC Davis\\
@@
               \email{rossibarra@ucdavis.edu}           %  \\
+              \and
+              Jose Gutierrez-Lopez\at
+              Department of Natural Resources and the Environment
+              University of New Hampshire\\
+              Durham, New Hampshire, 03820
 %             \emph{Present address:} of F. Author  %  if needed
@@
 We find that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
-In spite of this, we see little evidence that CentC variants form distinct genetic groups, instead finding that homoplasious mutations have likely homogenized CentC diversity.
+In spite of this, we see little evidence that CentC variants form distinct genetic groups, but instead find that homoplasious mutations may have homogenized CentC diversity.
 Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that a vast majority of all CentC repeats derive from only one of the parental genomes.
-Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC, but not the pericentromeric repeat Cent4, has decreased through domestication.
+Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC has decreased through domestication while the pericentromeric repeat Cent4 has drastically increased.
 
@@
 \label{intro}
-In spite of the rapid growth of sequenced plant genomes, plant centromeres remain poorly understood and relatively cryptic, due largely to their highly repetitive content.
+In spite of the rapid growth of sequenced plant genomes, plant centromeres remain poorly understood and relatively cryptic due to their highly repetitive content.
 Centromere repeats are highly diverse across taxa and their turnover appears to be very rapid \citep{Melters2012}.
@@
 Our simulation (code available at: \url{https://github.com/paulbilinski/CentC_Analyses/tree/master/Diversity_sims}) assumed that CentC has been evolving for  1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
-We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ \citep{clark2005}, and one generation per year.
+We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ \citep{clark2005estimating}, and one generation per year.
 
@@
 %  \includegraphics[width=0.75\textwidth]{circos.png}
-\includegraphics{Fig1_circos.png}
+\includegraphics{Fig1_circos}
 \caption{CentC repeat location in relation to the maize subgenomes.
@@
 \caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.
-On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetichores.
+On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetochores.
 Scale bar is in MB.
@@
 
-\jri{check tense throughout. i think some is past and some is present. also fill in species names X and Y and add CITEs} 
 Our analysis of centromere repeat diversity across the maize genome identifies thousands of copies exhibiting tremendous diversity. 
 But while we can cluster the repeats into groups of related sequences, these groups seem to have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
-We see no evidence of chromosome specific repeats such as that seen in X and Y \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanisms that homogenizes repeats across centromeres on different chromosomes.
-We further show that Cent4, once thought to be a chromosome-specific centromere repeat (CITE), appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere.
+We see no evidence of chromosome specific repeats such as that seen in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanisms that homogenizes repeats across centromeres on different chromosomes.
+We further show that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere.
 Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.
@@
 Instead of chromosome-specific groups, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).
-One possible explanation for this result is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
+One possible explanation for the biased inheritance pattern is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
 Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
-Large CentC clusters are clearly biased in favor of subgenome 1, which is known to have lower gene loss and higher average expression than maize subgenome 2 \cite{Schnable2011}. 
-Consistent with this idea, \citet{ReinhartBartel2002} find evidence of regulatory microRNA’s corresponding to centromeric repeats.   
-Of course, an alternative explanation is simply that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.
+Large CentC clusters are clearly biased in favor of subgenome 1, which is known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
+Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} find evidence of regulatory microRNA’s corresponding to centromeric repeats.   
+An alternative explanation is simply that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.
 
@@
 Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation of relatedness among repeats, and we observe genome-wide that many CentC’s within an array fall into the same genetic cluster.
-These observations are consistent with the simple idea that most repeats arise due to tandem duplication, and long distance transposition of CentC, while necessary to homogenize repeats across chromosomes \citep{Shi2010}, is relatively uncommon.
+Our observations are consistent with the simple idea that most repeats arise due to tandem duplication, and long distance transposition of CentC, while necessary to homogenize repeats across chromosomes \citep{Shi2010}, is relatively uncommon.
 
@@
 
-\jri{insert CITE, italics, probably refs to tables/figs as appropriate}  Previous cytogenetic work  identified  differences in centromere repeat content  between domesticated maize and its wild relatives Z.
-mays ssp.
-parviglumis,), Z.
-mays ssp.
-mexicana,, and Z.
-luxurians (hereafter parviglumis, mexicana, and luxurians) \citep{Albert2010} but was unable to quantify these differences.
-Our resequencing results show that while there is little difference in the distribution of CentC in tandem arrays, the total abundance of CentC has  decreased  \jri{can we say absolute abundance?
-we have genome size estimates, right?} during domestication, and we verify this with FISH in two maize-teosinte hybrid individuals (Figure \ref{abundance}).
-This result contrasts with the apparent increase in transposable element abundance  \citep{Chia2012}, suggesting that changes in CentC are not due to causes common to all repetitive sequence.
-\jri{ are we sure TEs have increase in absolute number?
-didn't chia et al.
-look at RPKM?
-wouldn't that increase if overall genome size decreases?
-if we control for genome size, does TE content show and absolute increase?} The maize genome is smaller than its teosinte counterpart, largely due to differences in the abundance of heterochromatic knobs (Poggio et al 1998).
+Previous cytogenetic work  identified  differences in centromere repeat content  between domesticated maize and its wild relatives \emph{Z.
+mays} ssp.
+\emph{parviglumis}, \emph{Z.
+mays} ssp.
+\emph{mexicana}, and \emph{Z.
+luxurians} (hereafter \emph{parviglumis}, \emph{mexicana,} and \emph{luxurians}) \citep{Albert2010} but was unable to quantify  differences.
+Our resequencing results show that while there is little difference in the distribution of CentC in tandem arrays, the absolute abundance of CentC has  decreased during domestication, and we verify this with FISH in two maize-teosinte hybrid individuals (Figure \ref{abundance}).
+Decreasing CentC content during domestication contrasts with the apparent lack of change in transposable element abundance after controlling for genome size \citep{Chia2012}, suggesting that changes in CentC are not due to causes common to all repetitive sequence.
+The maize genome is smaller than its teosinte counterpart, largely due to differences in the abundance of heterochromatic knobs \citep{poggio1998}.
 \citet{Zhang2012} have postulated an adaptive relationship between centromere size and genome size based on an observed correlation between centromere size and genome size across a number of grass species.
-This model would propose that the decrease in CentC abundance seen post-domestication is due to selection for smaller active centromeres to complement the smaller overall genome size.
-While our current data are insufficient to evaluate this conclusion, future work investigating differences in CentC content among maize landraces that vary in genome size (CITE) may provide an opportunity to further test this hypothesis.
+A model correlating centromere size to total genome size would propose that the decrease in CentC abundance seen post-domestication is due to selection for smaller active centromeres to complement the smaller overall genome size.
+While our current data are insufficient to evaluate this conclusion, future work investigating differences in CentC content among maize landraces that vary in genome size \citep{poggio1998} may provide an opportunity to further test our hypothesis.
 
@@
 \caption{ChIP Reads mapping to Unassembled from different Oat-Maize Addition (OMA) Lines.
-Reads were mapped using Bowtie2 \citep{langmead2012} with the paramater -very sensitive-local.}\label{supp.gaby}       % Give a unique label
+Reads were mapped using Bowtie2 \citep{langmead2012fast} with the paramater -very sensitive-local.}\label{supp.gaby}       % Give a unique label
 

### 2014-05-06 centc.tex
In process of incorporating mitch's edits

@@
 
-\author{Paul Bilinski \and Kevin Distor \and Jose Gutierez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and Kelly Dawe \and Jeffrey Ross-Ibarra}
+\author{Paul Bilinski \and Kevin Distor \and Jose Gutierrez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and Kelly Dawe \and Jeffrey Ross-Ibarra}
 
@@
 Centromere repeats are found in most eukaryotes and play a critical role in kinetochore formation.
-Though they exhibit considerable diversity both within and among species, little is understood about the mechanisms that drive centromere repeat evolution.
+Though CentC repeats exhibit considerable diversity both within and among species, little is understood about the mechanisms that drive centromere repeat evolution.
 Here, we use maize as a model to investigate how a complex history involving polyploidy, fractionation, and recent domestication has impacted the diversity of the maize CentC repeat.
@@
 We find that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
-In spite of this, we see little evidence that CentC variants form distinct genetic groups, but instead find that homoplasious mutations may have homogenized CentC diversity.
-Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that a vast majority of all CentC repeats derive from only one of the parental genomes.
+However, we do not observe distinct genetic groups within tandem arrays, but instead find that homoplasious mutations may have homogenized CentC diversity.
+Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that a vast majority of all CentC repeats derive from  one of the parental genomes.
 Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC has decreased through domestication while the pericentromeric repeat Cent4 has drastically increased.
@@
 \label{intro}
-In spite of the rapid growth of sequenced plant genomes, plant centromeres remain poorly understood and relatively cryptic due to their highly repetitive content.
+In spite of the rapid growth in the number of sequenced genomes,  centromeres remain poorly understood and relatively cryptic due to their highly repetitive content.
 Centromere repeats are highly diverse across taxa and their turnover appears to be very rapid \citep{Melters2012}. 
-However, little is known about the genetic mechanisms that produce centromere repeat diversity.
+However, little is known about the genetic mechanisms that produce centromere repeat diversity, though we can begin investigating potential mechanisms with improved assemblies of centromeric regions.
 Domesticated maize (\emph{Zea mays} ssp.
-\emph{mays}) has a high quality genome assembly \citep{Schnable2009} including complete sequence of two centromeres \citep{Wolfgruber2009}, and the breadth of research into maize centromeres makes it one of the best systems in which to investigate the processes governing centromere repeat evolution.
+\emph{mays}) has a high quality genome assembly \citep{Schnable2009} including complete sequence of two centromeres \citep{Wolfgruber2009}, and the breadth of research into maize centromeres makes it one of the best systems to investigate the processes governing centromere repeat evolution.
 
@@
 Both repeats interact with kinetochore proteins such as CENH3 \citep{Wolfgruber2009, Zhong2002} and show variation in abundance across taxa \citep{Albert2010}.
-But while there is considerable effort in investigating the molecular function of maize centromere repeats \citep{Ananiev1998B, Nagaki2003, Wolfgruber2009}, we know comparatively little about the evolution of the sequences themselves.
+While considerable effort has gone to investigating the molecular function of maize centromere repeats \citep{Ananiev1998B, Nagaki2003, Wolfgruber2009}, we know comparatively little about the evolution responsible for producing the current sequences.
 CRM elements are better understood, including the age and insertion preferences of different CRM families \citep{Wolfgruber2009, Sharma2008}.
-In contrast, there is no in-depth characterization of the genetic diversity of centromere repeats in the maize genome.
+In contrast, no in-depth characterization of the genetic diversity of centromere repeats in the maize genome exists.
 
 In this paper, we describe the patterns of diversity of centromere repeats across the maize genome.  
-We investigate the genetic diversity of these repeats, including whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation similar to that seen in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual repeats varies spatially across the genome.
-We then compare centromere abundance across a number of samples of maize and its wild relative teosinte.
-We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, but instead that most of the large blocks of CentC appear to have been inherited from only one of the ancestral subgenomes of maize.
+We investigate whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation of CentC similar to that seen in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual CentC repeats varies spatially across the genome.
+We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, despite most large arrays of CentC originating from  one of the ancestral subgenomes of maize.
 We show higher genetic similarity of CentC repeats within clusters, indicating the predominance of tandem duplications in the formation of new CentC copies.
-Lastly, we use low coverage sequencing and cytological analyses data to show that domesticated maize has less CentC than its wild relatives.
+Lastly, we use low coverage sequencing and cytological data to show that domesticated maize has less CentC than its wild relatives.
 
@@
 We downloaded 218 previously annotated CentC sequences \citep{Ananiev1998B, Nagaki2003} from Genbank.  
-We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference.
-We kept hits with a length of over 140bp and a minimum bit score of 100.
+We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference, keeping hits with a length of over 140bp and a minimum bit score of 100.
 After meeting the bit score threshold, the longest hit was retained.
@@
 Our resequencing results show that while there is little difference in the distribution of CentC in tandem arrays, the absolute abundance of CentC has  decreased during domestication, and we verify this with FISH in two maize-teosinte hybrid individuals (Figure \ref{abundance}).
-Decreasing CentC content during domestication contrasts with the apparent lack of change in transposable element abundance after controlling for genome size \citep{Chia2012}, suggesting that changes in CentC are not due to causes common to all repetitive sequence.
+%Decreasing CentC content during domestication contrasts with the apparent lack of change in transposable element abundance after controlling for genome size \citep{Chia2012}, suggesting that changes in CentC are not due to causes common to all repetitive sequence.
+Variability in observed abundance of transposable elements \citep{Chia2012} suggests that the decrease seen in CentC is not due to causes common to all repetitive sequences.
 The maize genome is smaller than its teosinte counterpart, largely due to differences in the abundance of heterochromatic knobs \citep{poggio1998}.

### 2014-05-07 centc.tex
all of mitch's edits in

@@
 
-We found a total of 12,162 CentC variants in the maize reference genome and unassembled BACs.
+We found a total of 12,162 CentC copies in the maize reference genome and unassembled BACs.  
 Of these, 8,259 were unique over their full length. 
@@
 Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.  
-Excluding the 1,523 copies that occur on the unassembled BACs, genome-wide CentC locations also show that nearly all of the 10,639 CentC copies are found in one of the 248 clusters identified on chromosomes 1-10; only 14 occurred as solo copies.
+%Excluding the 1,523 copies that occur on the unassembled BACs, 
+Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies on chromosomes 1-10 are found in clusters; only 14 occurred as solo copies.  
 Clusters varied in width from single CentC copies to 84KB with a mean of $\sim$7KB ($\sim$45 CentC copies). 
 Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.   
-For example, CENH3-ChIP sequence from an oat-maize addition line including only one maize chromosome \citep{kynast2001} nonetheless has many reads that map to the unassembled BACs (Table \ref{supp.gaby}).
-In particular, chromosome 6 had many more reads aligning to the unassembled BACs than it did to its own centromere, suggesting a particularly incomplete assembly.
+For example, CENH3-ChIP sequence from an oat-maize addition line with one maize chromosome \citep{kynast2001} has many reads that map to the unassembled BACs (Table \ref{supp.gaby}).  
+In particular, chromosome 6 has many more reads aligning to the unassembled BACs than it did to its own centromere repeats, suggesting a particularly incomplete assembly.  
 Examining total repeat number, chromosome 7 had the most CentC, with 3,200 copies, while chromosome 6 had the fewest with 32 copies.
@@
 
-In addition to CentC, previous studies have described Cent4, a second high copy number centromere repeat specific to maize chromosome 4 \citep{Page2001}. 
-Available evidence does not point to an important centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, and Cent4 probes lag behind CentC probes in cell division \citep{Jiang2002}.
-BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela, which was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
+Previous studies have also described the Cent4 repeat, a centromeric repeat that occurs only on chromosome 4 \citep{Page2001}. 
+Available evidence does not point to an important centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002}.
+BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela, that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
 Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.  
@@
 The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  
-This predicts that local clusters of CentC should be more closely related than CentC from different clusters.
-Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 indeed shows this pattern (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
+Tandem duplication predicts that  clusters of CentC should be more closely related than CentC from different clusters.  
+Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 shows that average genetic similarity is highest within clusters (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
 
@@
 The decreased genetic distance among CentCs in local clusters on chromosome 2 and 5 suggest that many of the genetic groupings discovered in our genome-wide analysis should correspond to local clusters of repeats.
-We see little evidence of this, however, as repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
-A comparison of all pairs of CentC sequences reveals a likely explanation: of the $\sim 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, likely grouping in genetic clusters despite their physical distance.
-Looking at several triplets at random from our alignment confirms that two sequences in one PCoA assignment sometimes share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.
+However, repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).  
+A comparison of shared mutations across all pairs of CentC sequences reveals a potential explanation.  
+Of the $\sim 74$ million possible pairs,  approximately 6 million share $\geq 2$ mutations different from the genome-wide consensus, causing CentC copies to group with sequences that share mutations irrespective of their physical distance.
+Comparing several triplets at random from our alignment confirms that two sequences in one PCoA assignment share greater pairwise identity than two sequences adjacent to one another in different PCoA groups.
 A simple forward simulation (see Methods) suggests this pattern could be due entirely to homoplasy rather than long-distance movement of CentC repeats.  
@@
 
-\subsection*{Variation in CentC abundance in \emph{Zea}}
+\subsection*{Variation of CentC abundance in \emph{Zea}}
 
 Shotgun sequence data from the maize HapMap v2 \citep{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize (p$<0.01$; Figure \ref{abundance}).   
-Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z.
-luxurians} \citep{Tenaillon2011}.
-Analysis of these data find nearly twice as much CentC in \emph{Z.
-luxurians} as the maize inbred B73.
+Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z. luxurians} \citep{Tenaillon2011}.  
+Analysis of these data find nearly twice as much CentC in \emph{Z. luxurians} as the maize inbred B73.  
 To corroborate these results, we performed fluorescent in-situ hybridization of F1 crosses between inbred maize and teosinte to determine if cytological observations agreed with our sequencing findings.
@@
 Our analysis of centromere repeat diversity across the maize genome identifies thousands of copies exhibiting tremendous diversity. 
-But while we can cluster the repeats into groups of related sequences, these groups seem to have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
-We see no evidence of chromosome specific repeats such as that seen in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanisms that homogenizes repeats across centromeres on different chromosomes.
+But while we can cluster the repeats into groups of related sequences, these groups have little relation to current or ancient maize chromosomes (Figure \ref{pcoa}).  
+We find no evidence of chromosome specific repeats as observed in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanism that homogenizes repeats across centromeres on different chromosomes.
 We further show that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere.
@@
 
-Instead of chromosome-specific groups, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).
+Additionally, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).  
 One possible explanation for the biased inheritance pattern is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
 Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
-Large CentC clusters are clearly biased in favor of subgenome 1, which is known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
-Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} find evidence of regulatory microRNA’s corresponding to centromeric repeats.   
-An alternative explanation is simply that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.
+Large CentC clusters are biased in favor of subgenome 1, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
+Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} found evidence of regulatory microRNA’s corresponding to centromeric repeats.   
+An alternative explanation is that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.
 

### 2014-05-11 centc.tex
working on kelly's edits
kelly edits JRI

@@
 We first validate the existence of long tandem arrays of repeats in maize and other taxa in the genus \emph{Zea}.  
-We find that genetic similarity among CentC copies is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
-However, we do not observe distinct genetic groups within tandem arrays, but instead find that homoplasious mutations may have homogenized CentC diversity.
-Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that a vast majority of all CentC repeats derive from  one of the parental genomes.
+Although we find considerable sequence diversity among CentC copies genome-wide, genetic similarity among repeats is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
+Genetic clustering analyses identify similar sequences among distant repeats, and simulations suggest that this pattern may be due to homoplasious mutation.  
+Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the vast majority of all CentC repeats derive from one of the parental genomes.
 Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC has decreased through domestication while the pericentromeric repeat Cent4 has drastically increased.
@@
 
-%We wanted to better understand the diversity present in all CentC copies across the genome.
-Looking at our alignment, we found that the sequences had a pairwise percent identify of 65.8\% with only 2.2\% of sites identical, suggesting large sequence diversity.
-However, these statistics do not inform us whether the amount of diversity observed would be expected due to cumulative mutation over time.
 In order to better understand patterns of diversity at CentC, we performed simulations to test the likelihood of homoplasious mutations (i.e.
@@
 
-Previous studies have also described the Cent4 repeat, a centromeric repeat that occurs only on chromosome 4 \citep{Page2001}. 
-Available evidence does not point to an important centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002}.
-BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela, that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
+Previous studies have also described the Cent4 repeat, a tandem pericentromeric repeat that occurs only primarily \jri{correct? it does blast elsewhere?} on chromosome 4 \citep{Page2001}. 
+Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002, jin2004maize}.
+BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
 Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.  
@@
 
-%Our study traces the changes in centromere repeat genetic relatedness across maize chromosomes to show how ancient tetraploidy and subsequent evolution has impacted centromeres.
-Most interestingly, we show that
-
 %!RELATEDNESS
@@
 We wish to thank Pacific Biosciences for sequencing resources.
+We thank Patrice Albert and James Birchler for providing the high-quality FISH image of the \emph{Z.
+luxurians} X B73 hybrid.
 We thank Siddharth Bhadra-Lobo, Vince Buffalo, Anne Lorant, Gernot Presting, Lauren Sagara, Michelle Stitzer, and National Science Foundation summer exchange program interns Cesar Alvarez Mejia, Aurelio Hernandez Bautista for advice and helpful discussion.

### 2014-05-12 centc.tex
Push b4 submit, noted one error in supplementary stuf I could not fix, but still going in.

@@
 
-\author{Paul Bilinski \and Kevin Distor \and Jose Gutierrez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and Kelly Dawe \and Jeffrey Ross-Ibarra}
+\author{Paul Bilinski \and Kevin Distor \and Jose Gutierrez-Lopez \and Gabriela Mendoza Mendoza \and Jinghua Shi \and R. Kelly Dawe \and Jeffrey Ross-Ibarra}
 
@@
 	  Athens, Georgia, 30602
-           \and
-           R. Kelly Dawe\at
-           Department of Genetics, University of Georgia\\
-           Athens, Georgia, 30602\\
 }
@@
 
-Fluorescent in situ hybridization (FISH) was used to compare abundance of CentC repeats in chromosomes of a maize x \emph{parviglumis} F1 hybrid and a maize x \emph{Z.
-luxurians} F1 hybrid.
-\jri{Kellyto fill in}
+Fluorescent in situ hybridization (FISH) was carried out as described in \citet{kato2004paint} (\emph{Z. luxurians} hybrid) and \citet{Shi2010} (\emph{Z. parviglumis} hybrid).
 
@@
 The outer ring depicts chromosomal assignment to the two subgenomes, with higher confidence regions are colored with darker colors.
+Green corresponds to subgenome 1, brown to subgenome 2.
 Breakpoints between the subgenomes remain uncolored to indicate uncertainty.
@@
 
-Previous studies have also described the Cent4 repeat, a tandem pericentromeric repeat that occurs only primarily \jri{correct? it does blast elsewhere?} on chromosome 4 \citep{Page2001}. 
+Previous studies have also described the Cent4 repeat, a tandem pericentromeric repeat that occurs primarily on chromosome 4 \citep{Page2001}. 
 Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002, jin2004maize}.
@@
 \centering
-\includegraphics[width=1\textwidth]{Figure4_noedge}
-\caption{(a) Mb of CentC in genomic libraries of maize and teosinte.
+\includegraphics[width=1\textwidth]{Fig4_kelly2}
+\caption{(a) FISH analysis of a single individual heterozygous for B73 and \emph{Zea luxurians} (GRIN accession PI422162).
+Chromosomes, ordered from 1 (left) to 10 (right), were hybridized with the Birchler probe cocktail \citep{kato2004paint}. 
+Green shows CentC and a 4-12-1 subtelomere repeat, blue the 180bp knob repeat, red the abundant TAG microsatellite and another subtelomeric repeat, white the TR1 knob repeat, orange the Cent4 repeat, yellow the 5S rDNA repeat, and aqua shows the Nucleolus Organizer Region.
+The green signals at the primary constrictions (arrows) are CentC.
+Note that \emph{Z. luxurians} has far brighter CentC signals.
+This image was graciously provided by Patrice Albert.
+(b) FISH analysis of a single individual heterozygous for B73 and \emph{Z. parviglumis}(GRIN accession PI566687).
+Chromosomes were hybridized with the \citet{Shi2010} probe cocktail showing CentC in red and CRM2 in green. 
+Each separate probe is shown separately below the two-color image, highlighting that CentC is more abundant in \emph{Z. parviglumis} and CRM2 is more abundant in maize. 
+(c) Mb of CentC in genomic libraries of maize and teosinte. 
 Box plots show data from \citet{Chia2012}. 
@@
 For comparison, the data point of maize inbred B73 in \citet{Chia2012} is shown with a tick mark on the box plot.
-(b) FISH images from an F1 hybrids between maize and the teosinte \emph{parviglumis}, and maize and the teosinte \emph{Z.
-luxurians}.
-Upper row shows the karyotype for the teosinte and the lower row shows the maize karyotype.
-Red=CentC, Green=CRM.
 }
@@
 Further work evaluating CentC evolution across multiple populations and multiple related species may shed additional light on the timing and causes of these changes.
-
-\jri{quick scan of the references looks OK, but do double check format etc.}
 

### 2014-05-20 centc.tex
fixed supplemental
added citation to include after revision

@@
 We investigate whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation of CentC similar to that seen in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual CentC repeats varies spatially across the genome.
+(Add cite of http://bioinformatics.oxfordjournals.org/content/26/17/2101.short)
 We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, despite most large arrays of CentC originating from  one of the ancestral subgenomes of maize.
@@
 
-\begin{figure}[]
+\begin{figure}[h!]
 % For two-column wide figures use
@@
 \includegraphics[width=1\textwidth]{Suppl_ACT_Seln.jpeg}
-\setcounter{figure}{0}
 \caption{Parameter selection for Alignment Candidate Threshold (ACT) for Mosaik.
@@
 
-\begin{table}[]
+\clearpage
+
+\begin{table}[h!]
 \caption{CentC Occurence Count In Maize RefGenv2}
@@
 % For tables use
-\begin{table}[]
+\begin{table}[h!]
 \caption{PacBio Read Counts and Tandem CentC}
@@
 % For tables use
-\begin{table}[]
+\begin{table}[h!]
 \caption{ChIP Reads mapping to Unassembled from different Oat-Maize Addition (OMA) Lines.

### 2014-06-09 response.tex
response to reviewers, initial draft

\documentclass[]{article}
\newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
\usepackage{color}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,       
  citecolor=blue,   
  filecolor=blue,  
  urlcolor=blue  
}
\begin{document}

\noindent Dr. Ingo Schubert and Anonymous Reviewers, \\

We very much appreciate your commentary on our manuscript and thank you for the opportunity to revise the manuscript to make it suitable for publication.  
We hope to address each of your comments in the following letter and revised paper. Comments are in black, and our response in blue.\\

\noindent Sincerely,\\

\noindent Jeffrey Ross-Ibarra and Paul Bilinski

\section*{Editor:}

Personally I wonder whether genome size differences between maize landraces will correlate with differences in CentC content (p. 11), because it is known from old cytogenetic studies that that they may vary in their number of non-centromeric heterochromatic knobs. \\

\res{We agree with the editor that the correlation between CentC abundance and knob number or genome size is of great interest. 
 Our current panel of maize lines available from the hapmap project is limited in its number of knobs, as many are inbred lines with relatively uniform knob abundance.  
However, we are investigating the correlation in another project using a broad panel of approximately 300 maize landraces across an altitudinal gradient.  
Though preliminary and unpublished, we find that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.  
We are currently in the process of further investigating this question, and hope to save our findings for another publication.}

\section*{Reviewer 1:}

Of particular note is the finding that "virtually all the large arrays of CentC in the maize genome derive from one of the two ancestral genomes present in modern day Zea”.  This is an interesting observation.  However, there are several potential pitfalls to this analysis that have not been addressed.  First, the authors state that "because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length."  In fact, large tandem repeat clusters are much more likely to be misassembled than shorter counterparts. \\

\res{We agree with the reviewer that large clusters are more likely to experience misassembly within the clusters.  
However, we had intended to refer to a lower likelihood of misassembly in the context of genomic location.  
Though large clusters are more likely to be incorrect in terms of copy number count or exact sequence, large clusters are more likely to be properly located on the maize iMap, a resource utilized in genome assembly that blended the physical, genetic, and optical maps of the maize genome.  
If the reviewer agrees with this perspective, we have altered the text in the manuscript to properly orient our point.}

\res{This has led to a large number of incorrect "joins" involving repeats.  
For the same reason perfect tandem repeats would have been collapsed into a single sequence in the final maize genome assembly.  
Therefore, the large CentC repeats are almost certainly misassembled and, if correct, would represent not the largest arrays but rather the most degenerate arrays of CentC repeats.}

\res{The reviewer makes an excellent point that the most degenerate arrays are also the most likely to not be collapsed into single CentC repeats during automated assembly.  Indeed, automated assembly may explain the apparent lack of identical tandem duplicates across the genome and, more importantly, directly impact our ability to determine copy number in extremely large clusters.  However, we would expect that any bias in assembly would impact clusters from the different subgenomes similarly and therefore should not impact our overall findings.  Only if the CentC copies originating in subgenome 2 regions have higher identity to one another than copies from subgenome 1 might we expect that the size of these particular arrays are being systematically underrepresented through automated assembly. We find no reason that mutation would impact CentC arrays differently and therefore do not believe that our ability to judge array size in subgenome 1 compared to subgenome 2 should differ greatly. }

\res{Furthermore, our ability to judge subgenome of origin is not likely to be impacted through assembly issues. In most cases, the origin of CentC clusters were determined with adjacent single copy markers specific to one of the subgenome as outlined in Schnable et al 2011.  The number of CentC's in the cluster would not change the origin call.}

\res{Regarding diversity of CentC copies from chromosomes with automated assembly versus primer walking, we also do not see striking differences.  Centromeres 2 and 5, where sequencing of CentC repeats was performed with primer walking, show similar levels of diversity across CentC repeats compared to other chromosomes.  We believe that the relatively uniform level of diversity observed across CentC is likely due to mutation over long periods of time.  Though assembly errors likely impact our ability to detect many of the tandem duplication instances of CentC in the maize genome, we believe that identical tandem duplicates account for a relatively small quantity of overall CentC copies as evidenced by primer walking on centromeres 2 and 5. }

In the abstract, the authors are even bolder with the following statement: "Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the vast majority of all CentC repeats derive from one of the parental genomes.”\\

\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors and any bias impacting copy number of CentC’s from each subgenome is likely alter both.}

Wild maize relatives (other subspecies) have the same five centromeres derived from each of the two maize ancestral sub genomes as the reference inbred genome, yet the authors themselves show by FISH that ALL centromeres of these wild species have similarly high amounts of CentC.\\

\res{We do acknowledge that FISH signal does indicate higher copy numbers of CentC in wild Zea across all chromosomes, but examination of the signals does not suggest uniformity across chromosomes.  For example, in figure 4B, the fluorescent signal on parviglumis chromosome 2 is much stronger than chromosome 3.  The centromeres on these chromosomes originate from different subgenomes.  Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content.}

I do agree with the author's finding that proximal CentC copies are more similar than distal ones, and that "The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies." Similar conclusions were found by Ma and Jackson (\url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC1361721/}).\\

\res{We acknowledge the finding in rice, have incorporated the citation, and are glad that our finding in maize that uses an different method is consistent with standing hypotheses.}

However, the authors observe that pairs of  CentC on different chromosomes share very high sequence similarity.  This again is a very interesting observation if one could be sure that the physical map is correct, but at least the early versions of the B73 map were fraught with major errors (entire BAC contigs and their associated centromeres being anchored to the wrong chromosome - see Schnable genome paper).  How sure are the authors that the specific repeats they find on multiple chromosomes are truly present, as opposed to having been misassigned to the wrong chromosome? This point should probably be addressed in the manuscript.\\

\res{We acknowledge that large assembly errors in regards to the exact location on a chromosome are likely common in the maize genome.  
Our goal in the investigation of identical CentC’s on different chromosomes or clusters is robust to such assembly errors.  
Our finding only requires that CentC copies not belong to the exact same cluster, which is highly likely given that most identical pairs derive from CentC copies from different BACs.  
%(do we need to tone down anything? include a sentence based on assembly pitfalls?)
}

\section*{Reviewer 2:}

1) Authors thought tandem duplication is the most important mechanism for the evolution of the repeat. However, as authors pointed out, tandem duplication generates similar sequence in near-by copies. I think gene conversion and unequal crossing over should play important roles for repeat evolution (in some part). For example, in Arabidopsis species (eg, Kawabe and Charlesworth 2007 JME), chimeric or mosaic structure repeats were found which could be originated by gene conversion or unequal crossing over. I think it is necessary to discuss about gene conversion and unequal crossing over in this manuscript.\\

\res{While gene conversion and unequal crossing over likely both play a large role CentC copy number change, our findings suggest that these events are likely ancient in the maize genome.  We do not find instances where entire clusters or segments of clusters are repeated, which we would observe if 2 clusters existed as a result of a gene conversion event.  Instead, we find many clusters with unique copies of CentC, suggesting that mutations have accumulated and made once identical clusters unique.}

2) Although it is not the main part in this manuscript, I wonder which arrays of CentC had decreased (or disappeared) from domesticated maize genome. Because not all CentC arrays form kinetochores, there is different importance of arrays. If it is possible to distinguish kinetochore forming and other arrays, please show level of decrements between them.\\

\res{Active kinetichores for each of the chromosomes are an ongoing project, with current boundaries published in Wolfgruber et al 2009.}

3) Authors conclude importance of tandem duplication because of the close location of highest similarity pairs and homogenization within arrays. However authors pointed out some distances between closest repeat and suggest that tandem duplication could be RELATIVELY old event. It is possible to estimate divergence between such pairs to show rough estimate of divergence times.\\

\res{The short nature of individual repeats provides for a small mutational target, causing estimates of age to have extremely high variance.  
Furthermore, the high levels of accumulated mutation make it difficult to discern ancestry of CentC copies, making age estimation a task outside the scope of this study.}

\subsection*{Minor Comments:}

\begin{itemize}
\item For the reference about chromosome specific repeat families, Hall SE et al. 2005 Genetics is more suitable than Pontes et al. 2004. Pontes et al. 2004 reported different centromere repeat from different species in allopolyploid species.

\res{Citation adjusted.}

\item I wonder why authors used Jukes-Cantor distance. I agree that the simplest method is sometimes the best, however mutation should not be uniform in centromeric repeat and cause wrong estimate values.

\res{We chose Jukes-Cantor distance as we have no better justification for using a more complex model.  }

\item page 4, materials and method simulation part, please show mutation rate per what. Per year or per generation?

\res{Per generation, assuming one generation.  Clarified in text.}

\item page 6, line 6 from bottom, a typo for kinetochore.

\res{Spelling corrected.}

\item fig3, there are lacks of labelling (a) and (b).

\res{Changes made.}

\end{itemize}

\end{document}

### 2014-06-10 response.tex
reviewer 2 done

@@
 \newcommand{\res}[1]{\noindent \textcolor{blue}{{#1}} \\}
+\newcommand{\jri}[1]{\textcolor{red}{{\bf #1}} }
+
 \usepackage{color}
@@
 Personally I wonder whether genome size differences between maize landraces will correlate with differences in CentC content (p.
-11), because it is known from old cytogenetic studies that that they may vary in their number of non-centromeric heterochromatic knobs.
+11), because it is known from old cytogenetic studies that they may vary in their number of non-centromeric heterochromatic knobs.
 \\
@@
 However, we are investigating the correlation in another project using a broad panel of approximately 300 maize landraces across an altitudinal gradient.  
-Though preliminary and unpublished, we find that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.  
-We are currently in the process of further investigating this question, and hope to save our findings for another publication.}
+Though preliminary, our as yet punpublished findings are that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.
+We are currently in the process of further investigating this question, and hope to save our findings for another publication.} \jri{is this still true?
+i can't remember where we left off in terms of B repeats and whether we were correcting for genome size appropriately}
 
@@
 
-\res{While gene conversion and unequal crossing over likely both play a large role CentC copy number change, our findings suggest that these events are likely ancient in the maize genome.
-We do not find instances where entire clusters or segments of clusters are repeated, which we would observe if 2 clusters existed as a result of a gene conversion event.
-Instead, we find many clusters with unique copies of CentC, suggesting that mutations have accumulated and made once identical clusters unique.}
+\res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text.
+Gene conversion likely serves to homogenize repeats across clusters and chromosomes, and may help explain why realted sequences are not always physically close.
+Nonetheless, most of our clusters contain many unique CentC, suggesting either that gene conversion is infreqeunt or point mutations are common.
+We have added text to the discussion mentioning gene conversion} \jri{we need to clarify text here about "tandem duplication" and add something to the discussion along the lines of the above.
+reviewer doesn't mean large-scale gene conversion (several kb) but instead small tracts the size of individual repeats.
+likely happens, but as you pointed out likely not rampant.
+don't think we can say ancient though.\\}
 
@@
 
-\res{Active kinetichores for each of the chromosomes are an ongoing project, with current boundaries published in Wolfgruber et al 2009.}
+\res{We are only able to analyze active/inactive kinetochores with much confidence in the fully sequenced centromeres 2 and 5.
+Blah blah...
+} \jri{ we should be able to say something here about 2 and 5.
+if i recall we didn't see any real differences?
+if we can say something, we can add one line/phrase to the text, which will make reviewer \#2 extra happy.\\ }
 
@@
 
-\res{The short nature of individual repeats provides for a small mutational target, causing estimates of age to have extremely high variance.  
-Furthermore, the high levels of accumulated mutation make it difficult to discern ancestry of CentC copies, making age estimation a task outside the scope of this study.}
+\res{The short nature of individual repeats provides for a small mutational target, causing standard estimates of age to have high variance.  
+Standard estimation of age also assumes fixation, while many CentC variants likely exist as copy number variants in the population.
+More sophisticated analysis (e.g.
+Thornton 2007 Genetics) require knowledge of the frequency of the copy number variant, which we do not have.
+Furthermore, we show in the text that homoplasious mutation is a problem, making it difficult to discern ancestry of CentC copies without error.
+As such, we felt that age estimation would not be complicated and unlikely to be very accurate.}
 
@@
 
-\res{We chose Jukes-Cantor distance as we have no better justification for using a more complex model.  }
+\res{We agree that mutation rates are unlikely to be uniform across the CentC repeat, but given the complexities of the alignment due to indels we felt using a simple method was the best approach. }
 

### 2014-06-10 response.tex
reviewer 1 started

@@
 Though assembly errors likely impact our ability to detect many of the tandem duplication instances of CentC in the maize genome, we believe that identical tandem duplicates account for a relatively small quantity of overall CentC copies as evidenced by primer walking on centromeres 2 and 5.
-}
+}\jri{jri edits needed still}
 
@@
 
-\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors and any bias impacting copy number of CentC’s from each subgenome is likely alter both.}
+\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors and any bias impacting copy number of CentC’s from each subgenome is likely alter both.}\jri{jri edits needed still}
 
@@
 The centromeres on these chromosomes originate from different subgenomes.
-Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content.}
+Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content.} \jri{jri edits needed still}
 
@@
 
-\res{We acknowledge the finding in rice, have incorporated the citation, and are glad that our finding in maize that uses an different method is consistent with standing hypotheses.}
+\res{We acknowledge the finding in rice, have incorporated the citation, and are glad that our finding in maize that uses a different method is consistent with standing hypotheses.}
 
@@
 
-\res{We acknowledge that large assembly errors in regards to the exact location on a chromosome are likely common in the maize genome.  
-Our goal in the investigation of identical CentC’s on different chromosomes or clusters is robust to such assembly errors.  
-Our finding only requires that CentC copies not belong to the exact same cluster, which is highly likely given that most identical pairs derive from CentC copies from different BACs.  
-%(do we need to tone down anything? include a sentence based on assembly pitfalls?)
-}
+\res{We acknowledge that large assembly errors in regards to the exact location on a chromosome likely exist in the maize reference genome, but suspect our result is robust to these errors.
+Our main finding is that CentCs are often most closely related to CentCs from different clusters.
+Because most clusters are on different BACs this result is relatively robust.
+While some of the highly-related pairs that we observe on different chromosomes may be due to assembly error (i.e.
+the BAC is placed on the wrong chromosome) we suspect it is unlikely that all such pairs can be explained by miss-assembly.
+It is also worth noting that centromeric BAC assemblies have improved since v1 of the reference genome, and LD-mapping with GBS data (unpublished) shows that while many BACs are in the wrong order, few of the BACs physically placed in a centromere are on the wrong chromosome.
+We are aware, however, of CentC-containing BACs that are placed on the wrong chromosome, so the reviewer is correct that this may explain part of the result.
+We have added text to this effect in the discussion.
+} \jri{ are there cases of centc on 2 that are most closely related to centcs on 5?
+if so, we should point that out here!
+also need to add some text to the discussion to say "while we know there is BAC error, blah blah." }
 

### 2014-06-18 response.tex
added cluster adjusting script, and the dataframe and test file for the script.  currently messes up on tallying total bp, but i think we only care about # of clusters in each subgenome

@@
 
-\res{This has led to a large number of incorrect "joins" involving repeats.  
+This has led to a large number of incorrect "joins" involving repeats.  
 For the same reason perfect tandem repeats would have been collapsed into a single sequence in the final maize genome assembly.  
-Therefore, the large CentC repeats are almost certainly misassembled and, if correct, would represent not the largest arrays but rather the most degenerate arrays of CentC repeats.}
+Therefore, the large CentC repeats are almost certainly misassembled and, if correct, would represent not the largest arrays but rather the most degenerate arrays of CentC repeats.\\
 

### 2014-06-18 AdjustCluster.pl

@@ -1,0 +1,112 @@
+#!/usr/bin/perl
+#Script to read in data frame with subgenome and cluster size and combine based on criteria
+use strict; #use warnings;
+
+die "usage: AdjustCluster.pl <Dist Cutoff> <clust size>" unless @ARGV==2;
+
+my $distcut = $ARGV[0];
+my $clustsize = $ARGV[1];
+
+open(orig, "<ClusterNeighbors_forscript.csv");
+#open(out, ">out_$distcut.$clustsize");
+
+my @bpcentc;
+my @distnext;
+my @subgenome;
+my @cluster;
+
+while(<orig>){
+	if ($_ =~ m/Cluster/){
+		@cluster = split(/,/, $_);
+	}
+#	if ($_ =~ m/Chr/){
+#		my @chr = split(/,/, $_);
+#	}
+#	if ($_ =~ m/Start/){
+#		my @Start = split(/,/, $_);
+#	}
+#	if ($_ =~ m/Stop/){
+#		my @Stop = split(/,/, $_);
+#	}
+#	if ($_ =~ m/Average/){
+#		my @Average = split(/,/, $_);
+#	}
+	if ($_ =~ m/BPCentC/){
+		@bpcentc = split(/,/, $_);
+	}	
+	if ($_ =~ m/DistancetoNext/){
+		@distnext = split(/,/, $_);
+	}	
+	if ($_ =~ m/Subgenome/){
+		@subgenome = split(/,/, $_);
+	}	
+}
+
+my $sub0 = 0;
+my $sub1 = 0;
+my $sub2 = 0;
+my $bpsub0 = 0;
+my $bpsub1 = 0;
+my $bpsub2 = 0;
+my $sumbp = 0;
+my $pos = 0;
+my $crap = 0;
+
+
+foreach my $ID (@distnext) {
+	if ($distnext[$pos] =~ m/Distance/){#skip first column
+		$sumbp=0;
+		$pos++;
+	} elsif ($distnext[$pos] =~ m/NA/){ #NA is end of chr, add up current tally to subgenome
+		if ($subgenome[$pos] == 0) { #add bp to its total subgenome bp and its count to subgenome count
+					$bpsub0 = $bpsub0 + $sumbp;
+					$sub0++;
+				}
+				if ($subgenome[$pos] == 1) {
+					$bpsub1 = $bpsub1 + $sumbp;
+					$sub1++;
+				}
[…]

### 2014-06-19 AdjustCluster.pl
Cluster Counter Works for counting all 10KB clusters, and also for each cluster individually.

@@ -58,18 +58,23 @@ foreach my $ID (@distnext) {
 		$sumbp=0;
 		$pos++;
 	} elsif ($distnext[$pos] =~ m/NA/){ #NA is end of chr, add up current tally to subgenome
-		if ($subgenome[$pos] == 0) { #add bp to its total subgenome bp and its count to subgenome count
-					$bpsub0 = $bpsub0 + $sumbp;
-					$sub0++;
-				}
-				if ($subgenome[$pos] == 1) {
-					$bpsub1 = $bpsub1 + $sumbp;
-					$sub1++;
-				}
-				if ($subgenome[$pos] == 2) {
-					$bpsub2 = $bpsub2 + $sumbp;
-					$sub2++;
-				}
+		my $tempbp = $bpcentc[$pos];
+		$sumbp = $sumbp + $tempbp;  #add its value to the running total
+		if ($sumbp >= $clustsize) {
+			if ($subgenome[$pos] == 0) { #add bp to its total subgenome bp and its count to subgenome count
+				$bpsub0 = $bpsub0 + $sumbp;
+				$sub0++;
+			}
+			if ($subgenome[$pos] == 1) {
+				$bpsub1 = $bpsub1 + $sumbp;
+				$sub1++;
+			}
+			if ($subgenome[$pos] == 2) {
+				$bpsub2 = $bpsub2 + $sumbp;
+				$sub2++;
+				print "$cluster[$pos] added to subgenome 2 WITH NA\n";
+			}
+		}
 		$sumbp=0; #end of chromosome, reset total bp
 		$pos++; #move to next cluster
 	} else { 
@@ -88,12 +93,19 @@ foreach my $ID (@distnext) {
 				if ($subgenome[$pos] == 2) {
 					$bpsub2 = $bpsub2 + $sumbp;
 					$sub2++;
+					print "$cluster[$pos] added to subgenome 2\n";
 				}
-			$sumbp = 0; #reset sumbp so you can move to next cluster
+				$sumbp = 0; #reset sumbp so you can move to next cluster
 			}
+			$sumbp = 0; #reset sumbp so you can move to next cluster
 		}		
-		elsif($distnext[$pos] <= $distcut) { #if distance is less than our cutoff, its tandem, and total up
+		elsif($distnext[$pos] < $distcut) { #if distance is less than our cutoff, its tandem, and total up
[…]

### 2014-06-19 AdjustCluster.pl
rewrite of perl

@@ -1,124 +1,45 @@
 #!/usr/bin/perl
 #Script to read in data frame with subgenome and cluster size and combine based on criteria
-use strict; #use warnings;
+use strict; 
+use warnings;
 
 die "usage: AdjustCluster.pl <Dist Cutoff> <clust size>" unless @ARGV==2;
 
 my $distcut = $ARGV[0];
 my $clustsize = $ARGV[1];
 
-open(orig, "<ClusterNeighbors_forscript.csv");
+open(ORIG, "<ClusterNeighbors_forscript.csv");
 #open(out, ">out_$distcut.$clustsize");
 
-my @bpcentc;
-my @distnext;
-my @subgenome;
-my @cluster;
+my @file=<ORIG>;
+my @cluster = split(/,/, $file[0]);
+my @bpcentc = split(/,/, $file[5]);
+my @distnext = split(/,/, $file[6]);
+my @subgenome = split(/,/, $file[7]);
+close ORIG;
 
-while(<orig>){
-	if ($_ =~ m/Cluster/){
-		@cluster = split(/,/, $_);
-	}
-#	if ($_ =~ m/Chr/){
-#		my @chr = split(/,/, $_);
-#	}
-#	if ($_ =~ m/Start/){
-#		my @Start = split(/,/, $_);
-#	}
-#	if ($_ =~ m/Stop/){
-#		my @Stop = split(/,/, $_);
-#	}
-#	if ($_ =~ m/Average/){
-#		my @Average = split(/,/, $_);
-#	}
-	if ($_ =~ m/BPCentC/){
-		@bpcentc = split(/,/, $_);
-	}	
-	if ($_ =~ m/DistancetoNext/){
-		@distnext = split(/,/, $_);
-	}	
-	if ($_ =~ m/Subgenome/){
-		@subgenome = split(/,/, $_);
-	}	
-}
-
-my $sub0 = 0;
-my $sub1 = 0;
-my $sub2 = 0;
-my $bpsub0 = 0;
-my $bpsub1 = 0;
-my $bpsub2 = 0;
+my @clustersub = (); # array for clusters in each subgenome
+my @bpsub = (); # array for bp in each subgenome
 my $sumbp = 0;
 my $pos = 0;
-my $crap = 0;
 
-
-foreach my $ID (@distnext) {
-	if ($distnext[$pos] =~ m/Distance/){#skip first column
-		$sumbp=0;
-		$pos++;
[…]

### 2014-06-19 centc.tex
scripts editted, final table S5 added to manuscript.  Im starting on text.

@@
 \usepackage{float}
+\usepackage{rotating}
 
@@
 
+\clearpage
+\begin{table}[h!]
+\caption{Number of clusters in each subgenome under varying criteria for distances apart.
+Minimum cluster size is 10,000bp, though trends remain the same under different mininum cluster sizes.}\label{supp.adjustclust}
+% For LaTeX tables use
+\begin{tabular}{ccccccc}
+\hline\noalign{\smallskip}
+Distance (bp) &  Unassigned BP &  Clusters Unassigned &  Subgenome1 BP &  Clusters Subgenome1 & Subgenome2 BP &  Clusters Subgenome2 \\\noalign{\smallskip}\hline\noalign{\smallskip}				
+1000 & 96504 & 6 & 829419 & 39 & 92126 & 7\\
+2000 & 109482 & 7 & 925028 & 43 & 115538 & 7\\
+3000 & 128492 & 8 & 969892 & 44 & 117094 & 7\\
+4000 & 128492 & 8 & 994104 & 41 & 124020 & 7\\
+5000 & 128492 & 8 & 1026845 & 40 & 136025 & 6\\
+6000 & 134203 & 8 & 1054805 & 39 & 140679 & 6\\
+7000 & 134203 & 8 & 1069516 & 38 & 140679 & 6\\
+8000 & 134203 & 7 & 1139270 & 36 & 140679 & 4\\
+9000 & 142786 & 7 & 1162918 & 37 & 147864 & 4\\
+10000 & 142786 & 7 & 1179529 & 36 & 149125 & 4\\
+11000 & 144185 & 7 & 1198451 & 36 & 149125 & 4\\
+12000 & 144185 & 6 & 1203008 & 34 & 149125 & 4\\
+13000 & 145273 & 6 & 1210187 & 32 & 149125 & 4\\
+14000 & 145273 & 6 & 1210329 & 31 & 149125 & 4\\
+15000 & 145273 & 6 & 1215760 & 30 & 149125 & 4\\
+16000 & 145273 & 5 & 1224106 & 27 & 152385 & 4\\
+17000 & 145273 & 5 & 1224106 & 25 & 152385 & 4\\
+18000 & 146060 & 5 & 1224106 & 25 & 157135 & 4\\
+19000 & 152703 & 5 & 1229229 & 25 & 157135 & 4\\
+20000 & 163555 & 6 & 1234836 & 25 & 157135 & 3\\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
 \end{document}

### 2014-06-19 AdjustCluster.pl

@@ -23,23 +23,19 @@ my @bpsub = (); # array for bp in each subgenome
 my $sumbp = 0;
 my $pos = 0;
 
-foreach my $i (1..$#distnext) {
-	#if the current cluster is far away from next one or on diff subgenome
-	if ( ($distnext[$i]=~m/NA/ ) || ($distnext[$i] >= $distcut) || ( $subgenome[$i] != $subgenome[$i-1]) ) { 		
-		#if the current total is bigger than our cluster size of interest
-		if ($sumbp >= $clustsize) { 
-			$bpsub[$subgenome[$i]]+=$sumbp;
+foreach my $i (1..$#distnext) { #walk through distance array, start from 1 not 0
+	$sumbp += $bpcentc[$i];
+	if ( ($distnext[$i]=~m/NA/ ) || ($distnext[$i] >= $distcut) || ( $subgenome[$i] != $subgenome[$i-1]) ) {  #lots of ors, needs to meet these criteria		
+		if ($sumbp >= $clustsize) { #is it bigger than current cluster size
+			$bpsub[$subgenome[$i]]+=$sumbp; #add to either 0 first or second part
 			$clustersub[$subgenome[$i]]++;
-			print "$cluster[$i] added to subgenome $subgenome[$i]\n";
+			#print "$cluster[$i]\t$sumbp\n"; # added to subgenome $subgenome[$i]\n";
 		}
-		$sumbp = $bpcentc[$i]; #reset sumbp so you can move to next cluster
+		$sumbp = 0; #reset sumbp so you can move to next cluster
 	}		
-	#if distance is less than our cutoff, its tandem, and total up
-	else { 
-		$sumbp += $bpcentc[$i];
-	}
 }
 
 for my $j (0..2){
-	print "$j\t$bpsub[$j]\t$clustersub[$j]\n";
+	print "$bpsub[$j]\t$clustersub[$j]\t";
 }
+print "\n";

### 2014-06-20 response.tex
Response done, editting paper

@@
 \res{We agree with the editor that the correlation between CentC abundance and knob number or genome size is of great interest. 
- Our current panel of maize lines available from the hapmap project is limited in its number of knobs, as many are inbred lines with relatively uniform knob abundance.  
+ Our current panel of maize lines available from the hapmap project is limited in its number of knobs, as many are inbred lines with relatively uniform knob abundance when compared to the variation seen in landraces, and therefore beleive it is not a suitable panel to investigate this question.
+(cut here if we don't want to go into it further, though I would think the editor would like this more in depth answer)
 However, we are investigating the correlation in another project using a broad panel of approximately 300 maize landraces across an altitudinal gradient.  
 Though preliminary, our as yet punpublished findings are that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.
-We are currently in the process of further investigating this question, and hope to save our findings for another publication.} \jri{is this still true?
-i can't remember where we left off in terms of B repeats and whether we were correcting for genome size appropriately}
+We are currently in the process of further investigating this question, and hope to save our findings for another publication.}
 
@@
 Though large clusters are more likely to be incorrect in terms of copy number count or exact sequence, large clusters are more likely to be properly located on the maize iMap, a resource utilized in genome assembly that blended the physical, genetic, and optical maps of the maize genome.
-If the reviewer agrees with this perspective, we have altered the text in the manuscript to properly orient our point.}
+If the reviewer agrees with this perspective, we have altered the text in the manuscript to properly orient our point. **PAPER**}
 
@@
 Only if the CentC copies originating in subgenome 2 regions have higher identity to one another than copies from subgenome 1 might we expect that the size of these particular arrays are being systematically underrepresented through automated assembly.
-We find no reason that mutation would impact CentC arrays differently and therefore do not believe that our ability to judge array size in subgenome 1 compared to subgenome 2 should differ greatly.
-}
+We find no reason that mutation would impact CentC arrays differently and therefore do not believe that our ability to judge array size in subgenome 1 compared to subgenome 2 should differ greatly.}
 
@@
 
-\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors and any bias impacting copy number of CentC’s from each subgenome is likely alter both.}\jri{jri edits needed still}
+\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors and any bias impacting copy number of CentC’s from each subgenome is likely alter both.
+We therefore believe our finding of absolute counts of CentC in each subgenome is robust to sequences errors.  
+However, from these comments, we thought it prudent to investigate whether our finding of cluster bias in subgenome 1 that was based on a set maximum distance between clusters, held under varying maximum cluster distances.
+Therefore, we wrote a script to reanalyze cluster and total bp membership in each subgenome dependent on incrementally increasing cluster sizes.
+Our findings show that cluster membership is also robust to varying tandem distance cut offs.
+The signal is likely driven by the strong bias of CentC from subgenome, showing that regardless of our definition of tandem clusters, CentC membership is biased to subgenome 1.
+We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.  **PAPER**}\jri{jri edits needed still}
 
@@
 \res{We acknowledge that large assembly errors in regards to the exact location on a chromosome likely exist in the maize reference genome, but suspect our result is robust to these errors.
-Our main finding is that CentCs are often most closely related to CentCs from different clusters.
-Because most clusters are on different BACs this result is relatively robust.
+Our main finding is that CentCs can be most closely related to CentCs from different clusters.
+Because most clusters are on different BACs in the genome assembly, our result would still provide evidence of homoplasy even if BACs were improperly assigned to a chromosome. 
 While some of the highly-related pairs that we observe on different chromosomes may be due to assembly error (i.e.
 the BAC is placed on the wrong chromosome) we suspect it is unlikely that all such pairs can be explained by miss-assembly.
+Furthermore, we investigated whether closest genetic distance pairs existed between chromosomes two and five.
+We find that several pairs exist, and have included text in the manuscript to reflect this more robust finding. (I have the sequence of these in a fasta, do I include as well?) **PAPER**
 It is also worth noting that centromeric BAC assemblies have improved since v1 of the reference genome, and LD-mapping with GBS data (unpublished) shows that while many BACs are in the wrong order, few of the BACs physically placed in a centromere are on the wrong chromosome.
 We are aware, however, of CentC-containing BACs that are placed on the wrong chromosome, so the reviewer is correct that this may explain part of the result.  
-We have added text to this effect in the discussion.
-} \jri{ are there cases of centc on 2 that are most closely related to centcs on 5?
-if so, we should point that out here!
-also need to add some text to the discussion to say "while we know there is BAC error, blah blah." }
+We have added text to this effect in the discussion. }
 
@@
 
-\res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text.
+\res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text. **PAPER**
 Gene conversion likely serves to homogenize repeats across clusters and chromosomes, and may help explain why realted sequences are not always physically close.
 Nonetheless, most of our clusters contain many unique CentC, suggesting either that gene conversion is infreqeunt or point mutations are common.
-We have added text to the discussion mentioning gene conversion} \jri{we need to clarify text here about "tandem duplication" and add something to the discussion along the lines of the above.
+We have added text to the discussion mentioning gene conversion}
+\jri{we need to clarify text here about "tandem duplication" and add something to the discussion along the lines of the above.
 reviewer doesn't mean large-scale gene conversion (several kb) but instead small tracts the size of individual repeats.
@@
 
-\res{We are only able to analyze active/inactive kinetochores with much confidence in the fully sequenced centromeres 2 and 5.
-Blah blah...
-} \jri{ we should be able to say something here about 2 and 5.
-if i recall we didn't see any real differences?
-if we can say something, we can add one line/phrase to the text, which will make reviewer \#2 extra happy.\\ }
+\res{We are only able to analyze active/inactive kinetochores with confidence in the fully sequenced centromeres 2 and 5.  
+On these chromosomes, we find no difference in genetic diversity of arrays inside and outside the active kinetochores.  
+The largest array on chromosome 2 is found in the kinetochore, but it does not show higher uniformity when compared to the arrays on chromosome 5, none of which are found in the kinetochore.  
+We have added text in the manuscript to emphasize these findings. *PAPER**}
 
@@
 Standard estimation of age also assumes fixation, while many CentC variants likely exist as copy number variants in the population. 
-More sophisticated analysis (e.g.
-Thornton 2007 Genetics) require knowledge of the frequency of the copy number variant, which we do not have.
+More sophisticated analysis (e.g. Thornton 2007 Genetics) require knowledge of the frequency of the copy number variant, which we do not have.
 Furthermore, we show in the text that homoplasious mutation is a problem, making it difficult to discern ancestry of CentC copies without error.  
@@
 
-\res{Citation adjusted.}
+\res{Citation adjusted. **Paper**}
 
@@
 
-\res{Per generation, assuming one generation.  Clarified in text.}
+\res{Per generation, assuming one generation.  Clarified in text. **PAPER**}
 
@@
 
-\res{Spelling corrected.}
+\res{Spelling corrected. **paper**}
 

### 2014-06-20 AdjustCluster.pl

@@ -1,5 +1,7 @@
 #!/usr/bin/perl
 #Script to read in data frame with subgenome and cluster size and combine based on criteria
+#for i in {1..20}; do BP=$(( $i*1000 )); perl AdjustCluster.pl $BP 10000;  done
+#code to execute it for a lot of diff distances
 use strict; 
 use warnings;
 

### 2014-06-20 centc.tex
Commit before leaving, made some changes from the responses

@@
 In this paper, we describe the patterns of diversity of centromere repeats across the maize genome.  
-We investigate whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation of CentC similar to that seen in other species \citep{Kawabe2005, Pontes2004} and how genetic relatedness among individual CentC repeats varies spatially across the genome.
-(Add cite of http://bioinformatics.oxfordjournals.org/content/26/17/2101.short)
+We investigate whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation of CentC similar to that seen in other species \citep{Kawabe2005, hall2005differential, macas2010global} and how genetic relatedness among individual CentC repeats varies spatially across the genome.
 We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, despite most large arrays of CentC originating from  one of the ancestral subgenomes of maize.
@@
 Our simulation (code available at: \url{https://github.com/paulbilinski/CentC_Analyses/tree/master/Diversity_sims}) assumed that CentC has been evolving for 1 million years since the divergence of maize and \emph{Tripsacum} \citep{Ross-Ibarra2009}, a closely related genus whose centromere repeat shares a large amount of homology \citep{Melters2012}.
-We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ \citep{clark2005estimating}, and one generation per year.
+We assumed a constant copy number, a mutation rate of $3 x 10^{-8}$ per generation \citep{clark2005estimating}, and one generation per year.
 
@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.  
-Because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
+Because large clusters are less likely to be misplaced within the genome, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
 We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.
@@
 Previous studies have also described the Cent4 repeat, a tandem pericentromeric repeat that occurs primarily on chromosome 4 \citep{Page2001}. 
-Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetichore \citep{Jiang2002, jin2004maize}.
+Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetochore \citep{Jiang2002, jin2004maize}.
 BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.

### 2014-06-25 response.tex
Commit before leaving, made some changes from the responses
Response and manuscript edits done by PB, time for JRI to look them over please

@@
 First, the authors state that "because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length."  In fact, large tandem repeat clusters are much more likely to be misassembled than shorter counterparts.
+Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).
 \\
@@
 However, we had intended to refer to a lower likelihood of misassembly in the context of genomic location.  
-Though large clusters are more likely to be incorrect in terms of copy number count or exact sequence, large clusters are more likely to be properly located on the maize iMap, a resource utilized in genome assembly that blended the physical, genetic, and optical maps of the maize genome.
-If the reviewer agrees with this perspective, we have altered the text in the manuscript to properly orient our point. **PAPER**}
+Though large clusters are more likely to be incorrect in terms of copy number count or exact sequence, large clusters are more likely to be properly located on the maize iMap (now cited in the manuscript), a resource utilized in genome assembly that blended the physical, genetic, and optical maps of the maize genome.
+If the reviewer agrees with this perspective, we have altered the text in the manuscript to properly orient our point.
+The reviewers point addressing novel insertion breaking up large clusters is very insightful.
+To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters to determine membership.
+The results are included in a new supplementary, and its consequences to further analyses such as subgenome counts are discussed later in our response.}
 
@@
 The signal is likely driven by the strong bias of CentC from subgenome, showing that regardless of our definition of tandem clusters, CentC membership is biased to subgenome 1.
-We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.  **PAPER**}\jri{jri edits needed still}
+We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.}\jri{jri edits needed still}
 
@@
 Furthermore, we investigated whether closest genetic distance pairs existed between chromosomes two and five.
-We find that several pairs exist, and have included text in the manuscript to reflect this more robust finding. (I have the sequence of these in a fasta, do I include as well?) **PAPER**
+We find that several pairs exist, and have included text in the manuscript to reflect this more robust finding. (I have the sequence of these in a fasta, do I include as well?)
 It is also worth noting that centromeric BAC assemblies have improved since v1 of the reference genome, and LD-mapping with GBS data (unpublished) shows that while many BACs are in the wrong order, few of the BACs physically placed in a centromere are on the wrong chromosome.
 We are aware, however, of CentC-containing BACs that are placed on the wrong chromosome, so the reviewer is correct that this may explain part of the result.  
-We have added text to this effect in the discussion. }
+We have added text to this effect in the discussion.}
 
@@
 
-\res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text. **PAPER**
+\res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text.
 Gene conversion likely serves to homogenize repeats across clusters and chromosomes, and may help explain why realted sequences are not always physically close.
 Nonetheless, most of our clusters contain many unique CentC, suggesting either that gene conversion is infreqeunt or point mutations are common. 
-We have added text to the discussion mentioning gene conversion}
+We have added text to the discussion mentioning gene conversion.} 
 \jri{we need to clarify text here about "tandem duplication" and add something to the discussion along the lines of the above.
@@
 
-\res{We are only able to analyze active/inactive kinetochores with confidence in the fully sequenced centromeres 2 and 5.  
+\res{We are only able to analyze genetic diversity of CentC copies within active/inactive kinetochores with confidence in the fully sequenced centromeres 2 and 5.  
 On these chromosomes, we find no difference in genetic diversity of arrays inside and outside the active kinetochores.  
 The largest array on chromosome 2 is found in the kinetochore, but it does not show higher uniformity when compared to the arrays on chromosome 5, none of which are found in the kinetochore.  
-We have added text in the manuscript to emphasize these findings. *PAPER**}
+We have added text in the manuscript to emphasize these findings.
+Unfortunately, it is not possible to determine whether active or inactive arrays are responsible for the change in abundance observed during domestication.}
 
@@
 
-\res{Citation adjusted. **Paper**}
+\res{Citation adjusted.}
 
@@
 
-\res{Per generation, assuming one generation.  Clarified in text. **PAPER**}
+\res{Per generation, assuming one generation.  Clarified in text.}
 
@@
 
-\res{Spelling corrected. **paper**}
+\res{Spelling corrected.}
 

### 2014-06-25 centc.tex
Response and manuscript edits done by PB, time for JRI to look them over please

@@
 
-Maize centromeres are comprised primarily of the 156bp satellite repeat CentC and the CRM family of retrotransposons.
+Maize centromeres are comprised primarily of the 156 bp satellite repeat CentC and the CRM family of retrotransposons.
 Both repeats interact with kinetochore proteins such as CENH3 \citep{Wolfgruber2009, Zhong2002} and show variation in abundance across taxa \citep{Albert2010}.
@@
 We downloaded 218 previously annotated CentC sequences \citep{Ananiev1998B, Nagaki2003} from Genbank.  
-We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference, keeping hits with a length of over 140bp and a minimum bit score of 100.
+We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference, keeping hits with a length of over 140 bp and a minimum bit score of 100.
 After meeting the bit score threshold, the longest hit was retained.
-We defined CentC’s as being in tandem if the CentC’s start location was within 1000bp of the start location of another CentC.
+In this analysis, we defined CentC’s as being in tandem if the CentC’s start location was within 1000 bp of the start location of another CentC.
 
@@
 We selected an optimum combination of parameters just below the parameters where we observed a large increase in the total number of reads aligning (Figure \ref{Supp_MPS}).
-Our final set of parameters for tandem repeats used an initial hash size of 8, an alignment candidate threshold of 15 bases, 20\% percent of mismatching bases, a minimum of 30\% overlap to the reference, and stored the top 100 hits for alignment.
+Our final set of parameters for tandem repeats used an initial hash size of 8, an alignment candidate threshold of 15 bases, 20\% of mismatching bases, a minimum of 30\% overlap to the reference, and stored the top 100 hits for alignment.
 After reads were mapped, we calculated the percentage of total reads hitting the given reference and multiplied this value by the relative genome size of each accession as reported in \citet{Chia2012} and \citet{Tenaillon2011}.
@@
 luxurians} with Pacific Biosciences (Pacific Biosciences, Menlo Park, CA) technology.
-Approximately 200Mbp of reads were produced from each cell, and reads with length greater than 600bp were retained for analysis of tandem CentC content using BLAST (Table \ref{supp.pacbio}).  
-CentC copies were considered tandem if the read had 4 CentC copies within 300bp of each other.
+Approximately 200 Mb of reads were produced from each cell, and reads with length greater than 600 bp were retained for analysis of tandem CentC content using BLAST (Table \ref{supp.pacbio}).  
+CentC copies were considered tandem if the read had 4 CentC copies each within 300 bp of each other.
 
@@
 Of these, 8,259 were unique over their full length.
+While centromeres 2 and 5 are the only chromosomes with high confidence sequencing of CentC copies, the levels of diversity observed on these two chromosomes is comparable to the rest of the genome, suggesting that current assemblies of CentC sequence may estimate sequence variation with some accuracy.
 No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.  
@@
 Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies on chromosomes 1-10 are found in clusters; only 14 occurred as solo copies.  
-Clusters varied in width from single CentC copies to 84KB with a mean of $\sim$7KB ($\sim$45 CentC copies). 
+Clusters varied in width from single CentC copies to 84 Kb with a mean of $\sim$7 Kb ($\sim$45 CentC copies). 
 Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.   
@@
 In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Table \ref{supp.pacbio}).  
-In one 6.7KB read from the maize reference line B73, for example, we identified approximately 40 independent CentC copies in tandem, and similar arrays were seen in all four \emph{Zea} species analyzed.
+In one 6.7 Kb read from the maize reference line B73, for example, we identified approximately 40 independent CentC copies in tandem, and similar arrays were seen in all four \emph{Zea} species analyzed.
 These results show that overall structure of the repeats has been maintained for the approximately 140,000 years since the \emph{luxurians}-\emph{mays} divergence \citep{Hanson1996,  Ross-Ibarra2009} and that a majority of CentC is found in tandem arrays (Table \ref{supp.pacbio}).
@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.  
-Because large clusters are less likely to be misplaced within the genome, we focused our analyses on the 52 clusters $>$10KB in length (Supplementary Fig \ref{Supp_clusterhist}).
+Due to the maize iMap resource \citep{zhou2009single}, large clusters are less likely to be misplaced within the genome.
+Therefore, we focused our analyses on the 52 clusters $>$10 Kb in length (Supplementary Fig \ref{Supp_clusterhist}).
 We assign clusters to a subgenome if they are flanked by two regions identified as originating from the same  subgenome.  
 Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
-If we restrict the analysis to clusters  $>$20KB with clear assignment to one subgenome, all 16 clusters were found in subgenome 1 regions.
-Even correcting for the genome-wide overrepresentation of subgenome 1 (62.5\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10KB and 20KB clusters).
-One cluster $>$20KB falls within an unassigned region on chromosome 3.
+If we restrict the analysis to clusters  $>$20 Kb with clear assignment to one subgenome, all 16 clusters were found in subgenome 1 regions.  
+Even correcting for the genome-wide overrepresentation of subgenome 1 (64.7\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10 Kb and 20 Kb clusters).
+One cluster $>$20 Kb falls within an unassigned region on chromosome 3.
+Our results are robust to varying definitions of clusters, as the bias is maintained when allowing for varying distances to define clusters \ref{supp.vartandem}.
 
@@
 The middle ring, shaded in blue, displays the locations of all CentCs across the genome.
-The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20KB in length}
+The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20 Kb in length}
 \label{circos}    
@@
 \includegraphics[width=1\textwidth]{Fig2_TWgroups}
-\caption{Presence of CentC in each of the heirarchical groups.
+\caption{Presence of CentC in each of the hierarchical groups.
 The 58 clusters found to be statistically significant in forming genetic groups are represented on the x-axis and chromosome of origin on the y-axis.
@@
 The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  
+Tandem duplications can occur through a variety of means, including slippage of the DNA polymerase \citep{jeeeeeeff.} or recombination that could lead to unequal crossing over or gene conversion \citep{edlund1981recombination}.
 Tandem duplication predicts that  clusters of CentC should be more closely related than CentC from different clusters.  
-Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 shows that average genetic similarity is highest within clusters (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50KB (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
+Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 shows that average genetic similarity is highest within clusters (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50 Kb (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
+Interestingly, we did not observe differences in genetic diversity when considering whether an array was found within or outside of the active kinetochore.
 
@@
 However, repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
+Several of these best pairs on different chromosomes are found between chromosomes 2 and 5, suggesting the phenomenon is not an artifact of errors in assembly.
 A comparison of shared mutations across all pairs of CentC sequences reveals a potential explanation.  
@@
 Chromosomes, ordered from 1 (left) to 10 (right), were hybridized with the Birchler probe cocktail \citep{kato2004paint}. 
-Green shows CentC and a 4-12-1 subtelomere repeat, blue the 180bp knob repeat, red the abundant TAG microsatellite and another subtelomeric repeat, white the TR1 knob repeat, orange the Cent4 repeat, yellow the 5S rDNA repeat, and aqua shows the Nucleolus Organizer Region.
+Green shows CentC and a 4-12-1 subtelomere repeat, blue the 180 bp knob repeat, red the abundant TAG microsatellite and another subtelomeric repeat, white the TR1 knob repeat, orange the Cent4 repeat, yellow the 5S rDNA repeat, and aqua shows the Nucleolus Organizer Region.
 The green signals at the primary constrictions (arrows) are CentC.
@@
 To corroborate these results, we performed fluorescent in-situ hybridization of F1 crosses between inbred maize and teosinte to determine if cytological observations agreed with our sequencing findings.
-FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z.
-luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
+FISH data supports our observation that the teosintes \emph{parviglumis} and \emph{Z. luxurians} have more CentC than inbred maize (Figure \ref{abundance}).
 Using whole genome shotgun PacBio long reads, we further investigated the overall structure of repeats across the different \emph{Zea} species.
@@
 We find no evidence of chromosome specific repeats as observed in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanism that homogenizes repeats across centromeres on different chromosomes.
-We further show that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere.
+%jeff, this needs to go here, but damnit its terrible.  i can't argue this well since i feel like its such a dumb point.
+Assembly errors, such as misplaced BACs known to occur in the maize genome or problems with assembling capturing biologically real CentC diversity, may also influence our ability to discover chromosomal patterns, and will only be improved over time with better sequencing technologies.
+We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere, though it is present in much of modern maize and absent in the teosintes \citep{albert2010}.
 Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.
@@
 Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
-Large CentC clusters are biased in favor of subgenome 1, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
+%POTENTIAL INSERT: We do not observe differences in genetic distances between CentC clusters on chromosomes 2 and 5 that are inside or outside the active kinetochore, which may suggest that purges occur on a large scale versus sustained mutation over an extended time.
+Large CentC clusters are biased in favor of subgenome 1 \ref{supp.vartandem}, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
 Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} found evidence of regulatory microRNA’s corresponding to centromeric repeats.
@@
 Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation of relatedness among repeats, and we observe genome-wide that many CentC’s within an array fall into the same genetic cluster.
-Our observations are consistent with the simple idea that most repeats arise due to tandem duplication, and long distance transposition of CentC, while necessary to homogenize repeats across chromosomes \citep{Shi2010}, is relatively uncommon.
+We observed no differences in genetic similarity when comparing clusters inside against cluster outside the active kinetochore.
+Our observations are consistent with the simple idea that most repeats arise due to tandem duplication or related processes with similar outcomes such as small scale gene conversion or unequal crossing over.
+Long distance transposition of CentC, while necessary to homogenize repeats across chromosomes \citep{Shi2010}, appears relatively uncommon.
 
@@
 Our simulations suggest that, under realistic assumptions about mutation rate and divergence time, such a pattern is possible due to homoplasious mutation alone.  
-Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication, but only 14\% of closest genetic pairs are found within 10KB of each other.
+Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication, but only 14\% of closest genetic pairs are found within 10 Kb of each other.
+Though assembly errors may explain a portion of these relationships, we find several closest distances pairs between chromosomes 2 and 5, suggesting our observation is biologically real.
 We speculate that the vast majority of  CentC’s in the genome are thus a result of relatively old tandem duplications, and that sufficient time has occurred since duplication for rearrangements and mutations to break up patterns of identical tandem repeats.
@@
 \caption{Graphs showing our ability to capture changes in CentC repeat abundance under constant genome size.
-We simulated 10MB of DNA with varying CentC content and  simulated Illumina reads from the DNA.
-Reads were mapped with our Mosaik pipeiline, and several simulations at each percentage of genomic content were performed.}
+We simulated 10 Mb of DNA with varying CentC content and  simulated Illumina reads from the DNA.
+Reads were mapped with our Mosaik pipeline, and several simulations at each percentage of genomic content were performed.}
 \label{Supp_Accuracy}    
@@
 \hline\noalign{\smallskip}
-Maize Line & Reads over 600bp & Reads with $\geq 4$ CentC & \% Reads Showing Tandem CentC \\
+Maize Line & Reads over 600 bp & Reads with $\geq 4$ CentC & \% Reads Showing Tandem CentC \\
 \noalign{\smallskip}\hline\noalign{\smallskip}				
@@
  \begin{longtable}[c]{|c|c|c|c|c|c|c|c|c|c|c|}
- \caption{Heirarchical clustering group assignment for copies of CentC, sorted by chromosome.  The number of CentC's from each chromosome is represented in the table.\label{long}}\\ 
+ \caption{Hierarchical clustering group assignment for copies of CentC, sorted by chromosome.  The number of CentC's from each chromosome is represented in the table.\label{long}}\\ 
 \hline
@@
 \begin{table}[h!]
+\label{supp.vartandem} 
 \caption{Number of clusters in each subgenome under varying criteria for distances apart.
-Minimum cluster size is 10,000bp, though trends remain the same under different mininum cluster sizes.}\label{supp.adjustclust}
+Minimum cluster size is 10,000 bp, though trends remain the same under different mininum cluster sizes.}\label{supp.adjustclust}
 % For LaTeX tables use
@@
 \hline\noalign{\smallskip}
-Distance (bp) &  Unassigned BP &  Clusters Unassigned &  Subgenome1 BP &  Clusters Subgenome1 & Subgenome2 BP &  Clusters Subgenome2 \\\noalign{\smallskip}\hline\noalign{\smallskip}				
+Distance ( bp) &  Unassigned  bp &  Clusters Unassigned &  Subgenome1  bp &  Clusters Subgenome1 & Subgenome2  bp &  Clusters Subgenome2 \\\noalign{\smallskip}\hline\noalign{\smallskip}				
 1000 & 96504 & 6 & 829419 & 39 & 92126 & 7\\

### 2014-06-25 DistancesHist.R

@@ -1,0 +1,14 @@
+distances <- c(6209604,1355654,10708,1398843,10659,8523,85021,4685,8593,10184,69276,29808,5571,5423,7838,12180,2502636,1863,7161,15167,2198,3180,12130,4461,34737,9714,8423,1793,10025,5742,3588,14802,4992,25081,22143,5565,29686,9496,18803,7707,4586,23386178,3942,8666,18312,NA,9232,43544,25054,1250,15628,5006,5952427,93418331,NA,42706,19258,2148803,67200,7112,2035,11317,5921,2058,15974,1450,10544,18988,1224,14221728,14257784,9476798,NA,14225472,54416488,9727128,17461,5724,8535,2173,19691,1050,3297,15789,83142020,NA,16407762,38053,3628126,4148409,3011,7174,24562,4597,9571,5091,41561,9056,10042,1365,16021,2228,15334,13111,7970,3253,NA,1307,NA,970606,2090,4572,1097,16446,1422,13895,9233,1247,2070,11418,9288,5185,2045,4509,2602,1792,6506,14062,1908,4987,11531,5437,3961,7950,3530,1430,4158,7806,52940,1430,1509,3072,3721,2853,19497426,12058013,1495,7097,5545,7339,4658,7766,8769,7848,2418622,3101335,1346,12247,7098,9640,4509,7239,106415,484025,1181926,4083,1848,3707,NA,1961,9063,1741,1201,7052,5934,4016,2163,4518,4329,7046,1472,3226,4283,20624242,30694655,NA,7367,27045536,1566203,15473134,1689,6828,2704,4195,12767,20564,29121,2414,15532,9909,6613,7136,23087,26624,2713,2717,1597,84715,11548,21309248,7715783,NA,2911,12157,8437,17999,59172,24331,2585,22597,20353,6130,7323,1793,7150,39074735,3210,14374,8636053,34106,19004,4529,382064,1028,25119,2269,266134,19498,2576344,7042,2530,15584,11103,34074,3560,146208,96416353,NA)
+plot(density(distances, na.rm=TRUE))
+hist(distances, breaks=100000, xlim=c(0,15000))
+notfar <- subset(distances, distances<50000)
+
+hist(notfar,breaks=50, na.rm=TRUE)
+
+hor <- c(1,2,3,4)
+vert <- c(.4, .6)
+barplot(vert)
+?barplot
+all<-c(1,3,2,2,2,2,2,2,1,3)
+hist(all)
+

### 2014-06-26 response.tex
response ready to go

@@
  Our current panel of maize lines available from the hapmap project is limited in its number of knobs, as many are inbred lines with relatively uniform knob abundance when compared to the variation seen in landraces, and therefore beleive it is not a suitable panel to investigate this question.
-(cut here if we don't want to go into it further, though I would think the editor would like this more in depth answer)
 However, we are investigating the correlation in another project using a broad panel of approximately 300 maize landraces across an altitudinal gradient.  
 Though preliminary, our as yet punpublished findings are that CentC exhibits an opposite trend to that observed in knobs; CentC increases when knob count and genome size decreases along an altitudinal gradient.
-We are currently in the process of further investigating this question, and hope to save our findings for another publication.}
+We are currently in the process of further investigating this question, and hope to save these findings for another publication.}
 
@@
 First, the authors state that "because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length."  In fact, large tandem repeat clusters are much more likely to be misassembled than shorter counterparts.
-Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).
 \\
@@
 \res{We agree with the reviewer that large clusters are more likely to experience misassembly within the clusters.  
-However, we had intended to refer to a lower likelihood of misassembly in the context of genomic location.  
+We had intended to refer to a lower likelihood of being in the wrong genomic location with regard to subgenome.  
 Though large clusters are more likely to be incorrect in terms of copy number count or exact sequence, large clusters are more likely to be properly located on the maize iMap (now cited in the manuscript), a resource utilized in genome assembly that blended the physical, genetic, and optical maps of the maize genome.
-If the reviewer agrees with this perspective, we have altered the text in the manuscript to properly orient our point.
-The reviewers point addressing novel insertion breaking up large clusters is very insightful.
-To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters to determine membership.
+We have altered the text in the manuscript to properly orient our point.}
+
+Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).
+
+\res{The reviewer's point addressing novel insertions breaking up large clusters is very insightful.
+To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters and variable cluster size to determine membership.
 The results are included in a new supplementary, and its consequences to further analyses such as subgenome counts are discussed later in our response.}
@@
 Indeed, automated assembly may explain the apparent lack of identical tandem duplicates across the genome and, more importantly, directly impact our ability to determine copy number in extremely large clusters.
-However, we would expect that any bias in assembly would impact clusters from the different subgenomes similarly and therefore should not impact our overall findings.  
-Only if the CentC copies originating in subgenome 2 regions have higher identity to one another than copies from subgenome 1 might we expect that the size of these particular arrays are being systematically underrepresented through automated assembly.
-We find no reason that mutation would impact CentC arrays differently and therefore do not believe that our ability to judge array size in subgenome 1 compared to subgenome 2 should differ greatly.}
+However, we would expect that any bias in assembly would impact clusters from the different subgenomes similarly and therefore should not impact our overall findings regarding subgenome.  }
 
@@
 
-\res{Regarding diversity of CentC copies from chromosomes with automated assembly versus primer walking, we also do not see striking differences.  
-Centromeres 2 and 5, where sequencing of CentC repeats was performed with primer walking, show similar levels of diversity across CentC repeats compared to other chromosomes.  
+\res{Regarding diversity of CentC copies from chromosomes with automated assembly, we also do not see striking differences.  
+Centromeres 2 and 5 were carefully assembled BAC-by-BAC as part of the reference genome (see Wolfgruber et al.
+2009).
+CentC in these centromeres  show similar levels of diversity across CentC repeats compared to other chromosomes.
 We believe that the relatively uniform level of diversity observed across CentC is likely due to mutation over long periods of time.  
-Though assembly errors likely impact our ability to detect many of the tandem duplication instances of CentC in the maize genome, we believe that identical tandem duplicates account for a relatively small quantity of overall CentC copies as evidenced by primer walking on centromeres 2 and 5.
-}\jri{jri edits needed still}
+Though assembly errors likely impact our ability to detect many of the tandem duplication instances of CentC in the maize genome, we believe that identical tandem duplicates account for a relatively small quantity of overall CentC copies as evidenced by centromeres 2 and 5.
+}
 
@@
 
-\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors and any bias impacting copy number of CentC’s from each subgenome is likely alter both.
-We therefore believe our finding of absolute counts of CentC in each subgenome is robust to sequences errors.  
-However, from these comments, we thought it prudent to investigate whether our finding of cluster bias in subgenome 1 that was based on a set maximum distance between clusters, held under varying maximum cluster distances.
-Therefore, we wrote a script to reanalyze cluster and total bp membership in each subgenome dependent on incrementally increasing cluster sizes.
-Our findings show that cluster membership is also robust to varying tandem distance cut offs.
+\res{As we explain above, we believe our ability to determine subgenome of origin is not comprised through assembly errors, and our reanalysis considering clusters of varying size and distance supports our original idea.
 The signal is likely driven by the strong bias of CentC from subgenome, showing that regardless of our definition of tandem clusters, CentC membership is biased to subgenome 1.
-We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.}\jri{jri edits needed still}
+We have included a new supplementary table in the text to report these findings, and have altered the text to include the reanalysis.}
 
@@
 The centromeres on these chromosomes originate from different subgenomes.  
-Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content.} \jri{jri edits needed still}
+Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content and are hesitant to try to compare CentC content quantitatively using FISH alone. }
 
@@
 
-\res{We acknowledge the finding in rice, have incorporated the citation, and are glad that our finding in maize that uses a different method is consistent with standing hypotheses.}
+\res{We thank the reviewer for pointing out this paper to us and have incorporated the citation.
+We're glad that our finding in maize that uses a different method is consistent with standing hypotheses.}
 
@@
 
-\res{We acknowledge that large assembly errors in regards to the exact location on a chromosome likely exist in the maize reference genome, but suspect our result is robust to these errors.
+\res{We acknowledge that large assembly errors in regards to the exact location on a chromosome likely exist in the maize reference genome (for example a cluster of CentC on the tip of the short arm of chromosome10 that belongs on chromosome 7), and have added text to this effect in the discussion.
+Nonetheless, we suspect our result is robust to these errors.
 Our main finding is that CentCs can be most closely related to CentCs from different clusters.
@@
 the BAC is placed on the wrong chromosome) we suspect it is unlikely that all such pairs can be explained by miss-assembly.
-Furthermore, we investigated whether closest genetic distance pairs existed between chromosomes two and five.
-We find that several pairs exist, and have included text in the manuscript to reflect this more robust finding. (I have the sequence of these in a fasta, do I include as well?)
-It is also worth noting that centromeric BAC assemblies have improved since v1 of the reference genome, and LD-mapping with GBS data (unpublished) shows that while many BACs are in the wrong order, few of the BACs physically placed in a centromere are on the wrong chromosome.
-We are aware, however, of CentC-containing BACs that are placed on the wrong chromosome, so the reviewer is correct that this may explain part of the result.  
-We have added text to this effect in the discussion.}
+Furthermore, looking only at the fully sequenced centromeres 2 and 5, we find several pairs of CentC on different chromosomes that are reciprocally closest genetically compared to all other CentC in the genome.
+We have included text in the manuscript to reflect this more robust finding. 
+It is also worth noting that centromeric BAC assemblies have improved since v1 of the reference genome, and LD-mapping with GBS data (unpublished) shows that while many BACs are in the wrong order, few of the BACs physically placed in a centromeres are on the wrong chromosome.
+}
 
@@
 \res{We consider unequal crossing over as a mechanism of tandem duplication; this is now clarified in the text.
-Gene conversion likely serves to homogenize repeats across clusters and chromosomes, and may help explain why realted sequences are not always physically close.
+Gene conversion likely serves to homogenize repeats across clusters and chromosomes, and may help explain why related sequences are not always physically close.
 Nonetheless, most of our clusters contain many unique CentC, suggesting either that gene conversion is infreqeunt or point mutations are common. 
 We have added text to the discussion mentioning gene conversion.}
-\jri{we need to clarify text here about "tandem duplication" and add something to the discussion along the lines of the above.
-reviewer doesn't mean large-scale gene conversion (several kb) but instead small tracts the size of individual repeats.
-likely happens, but as you pointed out likely not rampant.
-don't think we can say ancient though.\\}
 
@@
 Furthermore, we show in the text that homoplasious mutation is a problem, making it difficult to discern ancestry of CentC copies without error.  
-As such, we felt that age estimation would not be complicated and unlikely to be very accurate.}
+As such, we felt that age estimation would be complicated and unlikely to be very accurate.}
 

### 2014-06-26 centc.tex
edits done, needs a final read through

@@
 Of these, 8,259 were unique over their full length.
-While centromeres 2 and 5 are the only chromosomes with high confidence sequencing of CentC copies, the levels of diversity observed on these two chromosomes is comparable to the rest of the genome, suggesting that current assemblies of CentC sequence may estimate sequence variation with some accuracy.
+While centromeres 2 and 5 are the only chromosomes with high confidence sequencing of CentC copies, the levels of diversity observed on these two chromosomes is comparable to the rest of the genome (data not shown), suggesting that current assemblies of CentC sequence may estimate sequence variation with some accuracy.
 No CentC sequence occurred more than 10 times in the genome, and the vast majority ($>$75\%, Table \ref{supp.occurence}) of non-unique CentC variants occurred only twice.  
@@
 One cluster $>$20 Kb falls within an unassigned region on chromosome 3.
-Our results are robust to varying definitions of clusters, as the bias is maintained when allowing for varying distances to define clusters \ref{supp.vartandem}.
+This difference between the subgenomes is robust to different criteria of cluster size and distance (Table \ref{tab:suppvartandem}).
 
@@
 Comparisons of genetic and physical distance among CentC repeats on chromosomes 2 and 5 shows that average genetic similarity is highest within clusters (Figure \ref{heatmap}), revealing significant spatial autocorrelation of CentC variants over distances up to 10-50 Kb (Figure \ref{Supp_SpagediChr02} and \ref{Supp_SpagediChr05}).
-Interestingly, we did not observe differences in genetic diversity when considering whether an array was found within or outside of the active kinetochore.
 
@@
 However, repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
-Several of these best pairs on different chromosomes are found between chromosomes 2 and 5, suggesting the phenomenon is not an artifact of errors in assembly.
+We observed multiple pairs of CentC which occur in the same genetic cluster in spite of being on separated on the completely sequenced centromeres 2 and 5, suggesting that our result is not simply an artifact of errors in assembly.
 A comparison of shared mutations across all pairs of CentC sequences reveals a potential explanation.  
@@
 Shotgun sequence data from the maize HapMap v2 \citep{Chia2012}, reveals a significantly greater abundance of CentC in teosinte than in domesticated maize (p$<0.01$; Figure \ref{abundance}).   
-Further support for differences between maize and its wild relatives comes from shotgun sequence data from \emph{Z. luxurians} \citep{Tenaillon2011}.  
+Further support for differences between maize and its wild relatives comes from additional  sequence  from \emph{Z. luxurians} \citep{Tenaillon2011}.  
 Analysis of these data find nearly twice as much CentC in \emph{Z. luxurians} as the maize inbred B73.  
@@
 We find no evidence of chromosome specific repeats as observed in \emph{Arabidopsis} species \citep{Kawabe2005, Pontes2004}, suggesting the presence of a mechanism that homogenizes repeats across centromeres on different chromosomes.
-%jeff, this needs to go here, but damnit its terrible.  i can't argue this well since i feel like its such a dumb point.
-Assembly errors, such as misplaced BACs known to occur in the maize genome or problems with assembling capturing biologically real CentC diversity, may also influence our ability to discover chromosomal patterns, and will only be improved over time with better sequencing technologies.
-We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere, though it is present in much of modern maize and absent in the teosintes \citep{albert2010}.
+Although we believe our results relatively robust to assembly errors, misplaced BACs and collapsed tandem copies almost certainly occur in the maize  reference genome and may influence our assessment of chromsome-wide patterns.
+We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere, though it is present in much of modern maize and mostly absent in the teosintes \citep{albert2010}.
 Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.
@@
 Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
-%POTENTIAL INSERT: We do not observe differences in genetic distances between CentC clusters on chromosomes 2 and 5 that are inside or outside the active kinetochore, which may suggest that purges occur on a large scale versus sustained mutation over an extended time.
-Large CentC clusters are biased in favor of subgenome 1 \ref{supp.vartandem}, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
+We do not observe differences in genetic distances between CentC clusters on chromosomes 2 and 5 that are inside or outside the active kinetochore, which may suggest that purges occur on a large scale versus sustained mutation over an extended time.
+Large CentC clusters are biased in favor of subgenome 1 \ref{tab:suppvartandem}, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
 Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} found evidence of regulatory microRNA’s corresponding to centromeric repeats.
@@
 Roughly 80\% of the CentC repeats have their closest genetic relative on the same chromosome, as expected under a model of tandem duplication, but only 14\% of closest genetic pairs are found within 10 Kb of each other.
-Though assembly errors may explain a portion of these relationships, we find several closest distances pairs between chromosomes 2 and 5, suggesting our observation is biologically real.
+Though assembly errors may explain a portion of these relationships, we find several closest genetic pairs separated in the fully sequenced centromeres of chromosomes 2 and 5, suggesting our observation is biologically real.
 We speculate that the vast majority of  CentC’s in the genome are thus a result of relatively old tandem duplications, and that sufficient time has occurred since duplication for rearrangements and mutations to break up patterns of identical tandem repeats.
@@
 \begin{table}[h!]
-\label{supp.vartandem} 
 \caption{Number of clusters in each subgenome under varying criteria for distances apart.
@@
 \noalign{\smallskip}\hline
+\label{tab:suppvartandem} 
 \end{tabular}

### 2014-06-26 centc.tex
minor edits to text

@@
 Centromere repeats are found in most eukaryotes and play a critical role in kinetochore formation.  
-Though CentC repeats exhibit considerable diversity both within and among species, little is understood about the mechanisms that drive centromere repeat evolution.  
-Here, we use maize as a model to investigate how a complex history involving polyploidy, fractionation, and recent domestication has impacted the diversity of the maize CentC repeat.  
+Though centromere repeats exhibit considerable diversity both within and among species, little is understood about the mechanisms that drive centromere repeat evolution.  
+Here, we use maize as a model to investigate how a complex history involving polyploidy, fractionation, and recent domestication has impacted the diversity of the maize centromeric repeat CentC.  
 We first validate the existence of long tandem arrays of repeats in maize and other taxa in the genus \emph{Zea}.  
 Although we find considerable sequence diversity among CentC copies genome-wide, genetic similarity among repeats is highest within these arrays, suggesting that tandem duplications are the primary mechanism for the generation of new copies.
-Genetic clustering analyses identify similar sequences among distant repeats, and simulations suggest that this pattern may be due to homoplasious mutation.  
-Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the vast majority of all CentC repeats derive from one of the parental genomes.
-Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC has decreased through domestication while the pericentromeric repeat Cent4 has drastically increased.
+Nonetheless, clustering analyses identify similar sequences among distant repeats, and simulations suggest that this pattern may be due to homoplasious mutation.  
+Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the  majority of all CentC repeats derive from one of the parental genomes.
+Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC  likely decreased after domestication, while the pericentromeric repeat Cent4 has drastically increased.
 
@@
 Centromere repeats are highly diverse across taxa and their turnover appears to be very rapid \citep{Melters2012}. 
-However, little is known about the genetic mechanisms that produce centromere repeat diversity, though we can begin investigating potential mechanisms with improved assemblies of centromeric regions.
+However, little is known about the genetic mechanisms that produce centromere repeat diversity.
 Domesticated maize (\emph{Zea mays} ssp.

### 2014-06-27 centc.tex
more minor changes to text

@@
 While considerable effort has gone to investigating the molecular function of maize centromere repeats \citep{Ananiev1998B, Nagaki2003, Wolfgruber2009}, we know comparatively little about the evolution responsible for producing the current sequences.
-CRM elements are better understood, including the age and insertion preferences of different CRM families \citep{Wolfgruber2009, Sharma2008}.
+CRM elements are better understood, including the age and insertion preferences of different CRM families \citep{Wolfgruber2009, Sharma2008, Sharma09052014}.
 In contrast, no in-depth characterization of the genetic diversity of centromere repeats in the maize genome exists.
@@
 We investigate whether the differential ancestry of maize centromeres \citep{Wang2012} has led to chromosome-specific variation of CentC similar to that seen in other species \citep{Kawabe2005, hall2005differential, macas2010global} and how genetic relatedness among individual CentC repeats varies spatially across the genome.
-We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, despite most large arrays of CentC originating from  one of the ancestral subgenomes of maize.
+We find that CentC copies do not form genetic groups consistent with ancient whole genome duplications or chromosome specificity, despite most of the large arrays of CentC originating from only one of the ancestral subgenomes of maize.
 We show higher genetic similarity of CentC repeats within clusters, indicating the predominance of tandem duplications in the formation of new CentC copies.
@@
 We downloaded 218 previously annotated CentC sequences \citep{Ananiev1998B, Nagaki2003} from Genbank.  
-We then searched the maize genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference, keeping hits with a length of over 140 bp and a minimum bit score of 100.
-After meeting the bit score threshold, the longest hit was retained.
-In this analysis, we defined CentC’s as being in tandem if the CentC’s start location was within 1000 bp of the start location of another CentC.
+We then searched the B73 maize reference genome (5b60, \url{www.maizesequence.org}) with megaBLAST \citep{McGinnis2004} using the 218 annotated CentCs as a reference, keeping the longest hit with a length of over 140 bp and a minimum bit score of 100.
+In this analysis, we defined CentC copies as tandem if their start locations were within 1000 bp.
 
@@
 We mapped Illumina reads from a broad panel of \emph{Zea} species \citep{Chia2012,Tenaillon2011}  to a reference consisting of the full complement of 12,162 CentC variants identified in the B73 genome.
-We also used previously published whole genome chromatin immunoprecipitation (ChIP) \citep{Wolfgruber2009, Wang2013} with CenH3.
-Reads were mapped using  Mosaik v1.0 (\url{https://code.google.com/p/mosaik-aligner/}).
+We also used previously published sequence from whole genome chromatin immunoprecipitation (ChIP) \citep{Wolfgruber2009, Wang2013} using CenH3.
+Reads were mapped with  Mosaik v1.0 (\url{https://code.google.com/p/mosaik-aligner/}).
 We first optimized mapping parameters by relaxing mapping stringency and evaluating the number of successfully mapped reads with each combination.  
@@
 
-Library preparation and sequencing was performed according to the methods cited in \citep{Melters2012}.  
+Library preparation and sequencing was performed according to the methods described in \citep{Melters2012}.  
 Using those protocols, we sequenced one individual from \emph{mays}, \emph{mexicana}, \emph{parviglumis}, and \emph{Z.
@@
 
-Fluorescent in situ hybridization (FISH) was carried out as described in \citet{kato2004paint} (\emph{Z. luxurians} hybrid) and \citet{Shi2010} (\emph{Z. parviglumis} hybrid).
+Fluorescent in situ hybridization (FISH) was carried out as described in \citet{kato2004paint} for a B73 by \emph{Z. luxurians} hybrid) and \citet{Shi2010} for a B73 by \emph{Z. parviglumis} hybrid.
 
@@
 Of the 2,266 non-unique CentC sequences, only 3 were tandem, identical duplicates.  
-%Excluding the 1,523 copies that occur on the unassembled BACs, 
-Genome-wide CentC locations also show that nearly all of the 10,639 CentC copies on chromosomes 1-10 are found in clusters; only 14 occurred as solo copies.  
+Nearly all of the 10,639 CentC copies on chromosomes 1-10 are found in clusters; only 14 occurred as solo copies.  
 Clusters varied in width from single CentC copies to 84 Kb with a mean of $\sim$7 Kb ($\sim$45 CentC copies). 
-Chromosomes varied greatly in CentC copy number, though we know that centromere assemblies for all of the chromosomes are not complete.   
+Chromosomes varied greatly in CentC copy number, although centromere assemblies for all of the chromosomes are not complete.   
 For example, CENH3-ChIP sequence from an oat-maize addition line with one maize chromosome \citep{kynast2001} has many reads that map to the unassembled BACs (Table \ref{supp.gaby}).  

### 2014-06-27 centc.tex
working through final read

@@
 We used long-read Pacific Biosciences sequencing to verify that most CentC is in tandem arrays. 
-We sequenced whole genome ($\sim 0.1$X) libraries from 4 \emph{Zea} species.  
+We sequenced whole genome ($\sim 0.1$X) libraries from 4 \emph{Zea} taxa.  
 In spite of the low coverage, we recovered reads containing CentC sequence from all four taxa (Table \ref{supp.pacbio}).  
@@
 Thirty-eight of these clusters could be assigned to subgenome 1 (out of 43 assignable).
-If we restrict the analysis to clusters  $>$20 Kb with clear assignment to one subgenome, all 16 clusters were found in subgenome 1 regions.  
+If we restrict the analysis to clusters  $>$20 Kb with clear assignment to one subgenome, all 16 clusters were found in subgenome 1.  
 Even correcting for the genome-wide overrepresentation of subgenome 1 (64.7\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10 Kb and 20 Kb clusters).
@@
 \caption{CentC repeat location in relation to the maize subgenomes.
-The outer ring depicts chromosomal assignment to the two subgenomes, with higher confidence regions are colored with darker colors.
+The outer ring depicts chromosomal assignment to the two subgenomes, with higher confidence regions  in darker colors.
 Green corresponds to subgenome 1, brown to subgenome 2.
@@
 Available evidence does not point to any centromere function for Cent4: CenH3 chromatin immunoprecipitation data \citep{Wolfgruber2009, jin2004maize} show no significant over-representation of Cent4 compared to five known non-centromeric TE’s, fiber FISH shows clear separation of Cent4 from centromeric repeats \citep{jin2004maize} and Cent4 probes lag behind CentC probes in cell division, suggesting that they are not found in the kinetochore \citep{Jiang2002, jin2004maize}.
-BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}.
-Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.  
+BLAST analyses of Cent4 sequences from Genbank revealed high homology to the poorly characterized LTR retrotransposon RLX\_sela that was previously shown to be associated with heterochromatic  knobs \citep{Tenaillon2011, Chia2012}, but Cent4 lacks any of the protein sequences necessary for autonomous transposition, such as GAG and POL complexes.
 But while previous work in rice has documented the presence of nonautonomous LTR retrotransposons  in or near the centromere \citep{Jiang2002}, RLX\_sela also appears to be missing the necessary primer binding sites that would distinguish it as a nonautonomous TE, suggesting that it may be a TE-derived tandem repeat unique to the pericentromere of chromosome 4.
@@
 \end{figure}
-
+%JRI STOPPED HERE
 The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  

### 2014-06-27 centc.tex
SUBMIT

@@
 \end{figure}
-%JRI STOPPED HERE
+
 The tandem nature of CentC suggests it increases in copy number through local duplications that produce initially identical copies.  
-Tandem duplications can occur through a variety of means, including slippage of the DNA polymerase \citep{jeeeeeeff.} or recombination that could lead to unequal crossing over or gene conversion \citep{edlund1981recombination}.
+Tandem duplications can occur through a variety of means, including slippage of the DNA polymerase or recombination that could lead to unequal crossing over or gene conversion.
 Tandem duplication predicts that  clusters of CentC should be more closely related than CentC from different clusters.  
@@
 \caption{CentC physical location and genetic relatedness for (a) chromosome 2 and (b) chromosome 5.
-On the physical map, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetochores.
+On the physical map above, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetochores.
 Scale bar is in MB.
@@
 However, repeats within individual clusters are frequently found in different genetic groups as defined by principle coordinate analysis (Figure \ref{pcoa}).
-We observed multiple pairs of CentC which occur in the same genetic cluster in spite of being on separated on the completely sequenced centromeres 2 and 5, suggesting that our result is not simply an artifact of errors in assembly.
+We observed multiple pairs of CentC which occur in the same genetic cluster in spite of being separated on the completely sequenced centromeres 2 and 5, suggesting that our result is not simply an artifact of errors in assembly.
 A comparison of shared mutations across all pairs of CentC sequences reveals a potential explanation.  
@@
 \caption{(a) FISH analysis of a single individual heterozygous for B73 and \emph{Zea luxurians} (GRIN accession PI422162).
-Chromosomes, ordered from 1 (left) to 10 (right), were hybridized with the Birchler probe cocktail \citep{kato2004paint}. 
+Chromosomes, ordered from 1 (left) to 10 (right), were hybridized with the \citet{kato2004paint} probe cocktail. 
 Green shows CentC and a 4-12-1 subtelomere repeat, blue the 180 bp knob repeat, red the abundant TAG microsatellite and another subtelomeric repeat, white the TR1 knob repeat, orange the Cent4 repeat, yellow the 5S rDNA repeat, and aqua shows the Nucleolus Organizer Region.
@@
 Using whole genome shotgun PacBio long reads, we further investigated the overall structure of repeats across the different \emph{Zea} species.
-Percentages of the libraries showing tandem repeats did not differ greatly across the taxa (Figure \ref{supp.pacbio}).
+Percentages of the libraries showing tandem repeats were also higher in PacBio sequences from three teosinte compared to B73 (Table \ref{supp.pacbio}).
 
@@
 \label{discussion}
-
-%% STUFF JEFF MAY ADD TO DISCUSSION
-%Phylogenetic analysis of centromere repeats from a large number of eukaryotes found that repeat abundance evolves more rapidly than expected under a simple model of brownian motion \ref{Melters2012}.
 
@@
 Although we believe our results relatively robust to assembly errors, misplaced BACs and collapsed tandem copies almost certainly occur in the maize  reference genome and may influence our assessment of chromsome-wide patterns.
-We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere, though it is present in much of modern maize and mostly absent in the teosintes \citep{albert2010}.
-Inter-chromosomal transposition was postulated by \citet{Shi2010} as an explanation for marker genotypes showing evidence of recombination, and such transposition could also explain the similarity of repeats across all chromosomes.
+We further verify that Cent4, once thought to be a chromosome-specific centromere repeat \citep{Page2001}, appears to be a poorly characterized tandem repeat or nonautonomous retroelement, but is not associated with the centromere.
 
-Additionally, we find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}).  
-One possible explanation for the biased inheritance pattern is that centromere inactivation mirrors the differential expression seen between the subgenomes.  
-Centromere inactivation could then lead to differential loss of CentC clusters, such that large clusters remain only in the active centromeres.
-We do not observe differences in genetic distances between CentC clusters on chromosomes 2 and 5 that are inside or outside the active kinetochore, which may suggest that purges occur on a large scale versus sustained mutation over an extended time.
-Large CentC clusters are biased in favor of subgenome 1 \ref{tab:suppvartandem}, the subgenome known to have lower gene loss and higher average expression than maize subgenome 2 \citep{Schnable2011}. 
-Consistent with a correlation between expression and CentC loss in subgenome 2, \citet{ReinhartBartel2002} found evidence of regulatory microRNA’s corresponding to centromeric repeats.
-An alternative explanation is that the subgenome 1 parent had more or larger CentC arrays, leading to an initial unequal representation in the ancient polyploid ancestor of maize.
+We find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}, Table \ref{tab:suppvartandem}).
+This biased ancestry mirrors differences in genic expression and deletion seen between the subgenomes  \citep{Schnable2011}.
+Higher deletion rates on subgenome 2 may explain our observation, but the finding of small regulatory RNA’s corresponding to centromeric repeats \citet{ReinhartBartel2002} in other taxa may suggest a more active mechanism behind the observed differences.
+
+%% STUFF JEFF MAY ADD TO DISCUSSION
+%Phylogenetic analysis of centromere repeats from a large number of eukaryotes found that repeat abundance evolves more rapidly than expected under a simple model of brownian motion \ref{Melters2012}.
 
@@
 Our sequence comparison of CentCs also enabled us to explore the relationship between genetic and physical distance among repeats.  
-Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation of relatedness among repeats, and we observe genome-wide that many CentC’s within an array fall into the same genetic cluster.
+Using the well-assembled centromeres on chromosomes 2 and 5, we found spatial autocorrelation of relatedness among repeats, but also observed genome-wide that many CentC’s within an array fall into the same genetic cluster.
 We observed no differences in genetic similarity when comparing clusters inside against cluster outside the active kinetochore.
@@
 \emph{mexicana}, and \emph{Z.
-luxurians} (hereafter \emph{parviglumis}, \emph{mexicana,} and \emph{luxurians}) \citep{Albert2010} but was unable to quantify  differences.
+luxurians}  \citep{Albert2010} but was unable to quantify  differences.
 Our resequencing results show that while there is little difference in the distribution of CentC in tandem arrays, the absolute abundance of CentC has  decreased during domestication, and we verify this with FISH in two maize-teosinte hybrid individuals (Figure \ref{abundance}).

### 2014-07-01 response.tex
edits done, needs a final read through
Final before resubmit

@@
 First, the authors state that "because large clusters are less likely to be misassembled, we focused our analyses on the 52 clusters $>$10KB in length."  In fact, large tandem repeat clusters are much more likely to be misassembled than shorter counterparts.
+The maize genome assembly was done with minimum coverage and targeted to genic regions, with gaps over repeats being ignored, meaning no attempt was made to "finish", i.e.
+close these gaps.
 \\
@@
 We have altered the text in the manuscript to properly orient our point.}
-
-Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).
-
-\res{The reviewer's point addressing novel insertions breaking up large clusters is very insightful.
-To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters and variable cluster size to determine membership.
-The results are included in a new supplementary, and its consequences to further analyses such as subgenome counts are discussed later in our response.}
 
@@
 
+Also, tandem repeats are often broken up by retrotransposon insertions, so that large uninterrupted CentC runs could be interpreted quite differently as "immune to retrotransposon insertion" (although this is certainly not what I suggest is going on).\\
+
+\res{The reviewer's point addressing novel insertions breaking up large clusters is very insightful.
+To investigate this, we reanalyzed cluster size data allowing for variable distances between clusters and variable cluster size to determine membership.
+The results are included in a new supplementary figure, and its consequences to further analyses such as subgenome counts are discussed later in our response.}
+
 In the abstract, the authors are even bolder with the following statement: "Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the vast majority of all CentC repeats derive from one of the parental genomes.”\\
@@
 
-Wild maize relatives (other subspecies) have the same five centromeres derived from each of the two maize ancestral sub genomes as the reference inbred genome, yet the authors themselves show by FISH that ALL centromeres of these wild species have similarly high amounts of CentC.\\
+Wild maize relatives (other subspecies) have the same five centromeres derived from each of the two maize ancestral sub genomes as the reference inbred genome, yet the authors themselves show by FISH that ALL centromeres of these wild species have similarly high amounts of CentC.
+This conclusion, therefore, is not justified in my mind.
+\\
 
@@
 The centromeres on these chromosomes originate from different subgenomes.  
-Furthermore, we believe a qualitative measure such as fluorescence is not a good measure for precise CentC content and are hesitant to try to compare CentC content quantitatively using FISH alone. }
+We believe a qualitative measure such as fluorescence is not a good measure for precise CentC content and are hesitant to try to compare CentC content quantitatively using FISH alone.}
+
+Also, I agree with the authors that "Centromere inactivation could then lead to differential loss of CentC clusters", but wouldn't the same centromeres be inactivated in the wild relative as in the maize inbred?
+In other words, the inactive centromeres have lost CentC long before the domesticated maize diverged from the wild ancestors.
+\\In sum, I find the observation of these larger CentC clusters originating from one ancestral genome extremely interesting, but am not sure it has been proven to be real (as opposed to an artifact of the genome assembly) or explained quite satisfactorily.\\
+
+\res{
+We have also modified the text regarding centromere inactivation.  
+Similar numbers of currently active centromeres come from subgenome 2 (4 centromeres) and subgenome 1 (6 centromeres).  
+Moreover, evidence from the fully sequenced centromere 5 suggests that large CentC arrays are often found outside of the active kinetochore.  
+The bias in CentC content is thus likely to be the result of subgenome-wide patterns of deletion rather than centromere inactivation.
+} %JEFFFF}
 
@@
 The largest array on chromosome 2 is found in the kinetochore, but it does not show higher uniformity when compared to the arrays on chromosome 5, none of which are found in the kinetochore.  
-We have added text in the manuscript to emphasize these findings.
 Unfortunately, it is not possible to determine whether active or inactive arrays are responsible for the change in abundance observed during domestication.}

### 2014-07-01 centc.tex
Final before resubmit

@@
 We designed a simulation to estimate the accuracy of our measurements of CentC content (code available at: \url{https://github.com/kddistor/dnasims}).  
-In short, our simulations altered the copy number of CentC repeats over a region of fixed length (10Mb), changing the percentage of the genome deriving from the repeat.  
+In short, our simulations altered the copy number of CentC repeats over a region of fixed length (10 Mb), changing the percentage of the genome deriving from the repeat.  
 Illumina reads were simulated from each of the DNA strings and mapped using our pipeline.
@@
 On the physical map above, red lines show locations of numbered CentC clusters and blue blocks show the location of the active kinetochores.
-Scale bar is in MB.
+Scale bar is in Mb.
 Below each physical map is shown a heatmap of genetic relatedness of each CentC to (top row) other copies within its island of tandem repeats delineated by dotted lines and (bottom row) all other copies on the chromosome.

### 2014-07-14 centc.tex
Accepted! some touch ups round 1

@@
 Nonetheless, clustering analyses identify similar sequences among distant repeats, and simulations suggest that this pattern may be due to homoplasious mutation.  
-Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the majority of all CentC repeats derive from one of the parental genomes.
+Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the majority of all CentC repeats derive from one of the parental genomes, with a even stronger bias when examining the largest assembled contiguous clusters.
 Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC likely decreased after domestication, while the pericentromeric repeat Cent4 has drastically increased.
@@
 We then identified how many large clusters of CentC were retained from each of the two parental genomes that comprise the extant maize genome, referred to here as subgenome 1 and subgenome 2 (Figure \ref{circos}).
-Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}.  
-Due to the maize iMap resource \citep{zhou2009single}, large clusters are less likely to be misplaced within the genome.
+Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}, with centromeres 1, 2, 5, 7, 9, and 10 deriving from subgenome 1 and 3, 4, 6, 8 from subgenome 2.
+Due to the maize iMap resource \citep{zhou2009single}, we believe large clusters are less likely to be misplaced within the genome.
 Therefore, we focused our analyses on the 52 clusters $>$10 Kb in length (Supplementary Fig \ref{Supp_clusterhist}).
@@
 
-In conclusion, our detailed study of centromere repeats in the maize genome has highlighted differential contribution of subgenome, spatial autocorrelation along a chromosomes, and changes in abundance over the short time scale of domestication.
-Further work evaluating CentC evolution across multiple populations and multiple related species may shed additional light on the timing and causes of these changes.
+In conclusion, our detailed study of centromere repeats in the B73 maize genome has highlighted differential contribution of subgenome, spatial autocorrelation along a chromosomes, and changes in abundance over the short time scale of domestication.
+Since maize lines show vast cytogentic variation, further work evaluating CentC evolution across multiple populations and multiple related species may shed additional light on the timing and causes of these changes.
 

### 2014-07-16 centc.tex
PB final manuscript for JRI to look over changes

@@
 Previous work identified the parental genome for individual chromosomal segments \citep{Schnable2011} and centromeres \citep{Wang2012}, with centromeres 1, 2, 5, 7, 9, and 10 deriving from subgenome 1 and 3, 4, 6, 8 from subgenome 2.
-Due to the maize iMap resource \citep{zhou2009single}, we believe large clusters are less likely to be misplaced within the genome.
+Due to the maize iMap resource \citep{zhou2009single}, large clusters should be less likely to be misplaced within the genome.
 Therefore, we focused our analyses on the 52 clusters $>$10 Kb in length (Supplementary Fig \ref{Supp_clusterhist}).
@@
 If we restrict the analysis to clusters  $>$20 Kb with clear assignment to one subgenome, all 16 clusters were found in subgenome 1.  
-Even correcting for the genome-wide overrepresentation of subgenome 1 (64.7\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for both 10 Kb and 20 Kb clusters).
+Even correcting for the genome-wide overrepresentation of subgenome 1 (64.7\% of assigned base pairs), these results suggest a strong inequality in the origin of large CentC clusters (Fisher's exact test, $p<0.005$ for total CentC, 10 Kb, and 20 Kb clusters).
 One cluster $>$20 Kb falls within an unassigned region on chromosome 3.
@@
 %  \includegraphics[width=0.75\textwidth]{circos.png}
-\includegraphics{Fig1_circos}
+\includegraphics[width=10cm,height=10cm,keepaspectratio]{Fig1_circosv3.png}
 \caption{CentC repeat location in relation to the maize subgenomes.
@@
 The middle ring, shaded in blue, displays the locations of all CentCs across the genome.
-The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20 Kb in length}
+The inner ring, shaded in yellow, displays the locations of all CentC clusters greater than 20 Kb in length.
+Asterisks indicate approximate location of the centromere}
 \label{circos}    
@@
 \centering
-\includegraphics[width=1\textwidth]{Fig4_kelly2}
+\includegraphics[width=15cm,height=15cm,keepaspectratio]{Fig4_kelly2}
 \caption{(a) FISH analysis of a single individual heterozygous for B73 and \emph{Zea luxurians} (GRIN accession PI422162).
@@
 
-We find that virtually all the large arrays of CentC in the maize  genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}, Table \ref{tab:suppvartandem}).
+We find that virtually all the large arrays of CentC in the maize reference genome derive from one of the two ancestral genomes present in modern day \emph{Zea} (Figure \ref{circos}, Table \ref{tab:suppvartandem}).
 This biased ancestry mirrors differences in genic expression and deletion seen between the subgenomes  \citep{Schnable2011}. 
@@
 In conclusion, our detailed study of centromere repeats in the B73 maize genome has highlighted differential contribution of subgenome, spatial autocorrelation along a chromosomes, and changes in abundance over the short time scale of domestication.
-Since maize lines show vast cytogentic variation, further work evaluating CentC evolution across multiple populations and multiple related species may shed additional light on the timing and causes of these changes.
+Since maize lines show vast cytogentic variation, further work evaluating CentC evolution across multiple populations and multiple taxa may shed additional light on the timing and causes of these changes.
 

### 2014-07-19 centc.tex
added useles comment
n

@@
 Nonetheless, clustering analyses identify similar sequences among distant repeats, and simulations suggest that this pattern may be due to homoplasious mutation.  
-Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the majority of all CentC repeats derive from one of the parental genomes, with a even stronger bias when examining the largest assembled contiguous clusters.
+Although the two ancestral subgenomes of maize have contributed nearly equal numbers of centromeres, our analysis shows that the majority of all CentC repeats derive from one of the parental genomes, with an even stronger bias when examining the largest assembled contiguous clusters.
 Finally, by comparing maize with its wild progenitor teosinte, we find that the abundance of CentC likely decreased after domestication, while the pericentromeric repeat Cent4 has drastically increased.
@@
 \label{methods}
-
+%this is a comment
 \subsection*{CentC Repeat Identification and Diversity}
